# Baseline characteristics by study and intervention arm

The IMiC subsamples analyzed here come from three randomized controlled trials: MISAME-III (Burkina Faso), Mumta-LW (Pakistan), and ELICIT (Tanzania). Baseline maternal, infant, and household characteristics are presented for each study by randomized intervention arm, at the earliest postnatal visit with milk sample collection in each trial.

The tables report demographic, anthropometric, socioeconomic, and household variables plausibly associated with maternal nutritional status, breastfeeding context, or infant outcomes: infant sex, maternal age, education, height (Mumta-LW and ELICIT only), BMI, MUAC, parity, household size and number of rooms, flooring, cooking space, water source, season and place of delivery, asset-based household wealth, and food security. These variables were harmonized across studies during IMiC data curation; the manuscript's Materials and methods list those used as adjustment covariates. The manuscript Methods cite these summaries as Section 1 of the online resource. These tables have no supplementary-table number.

The summaries are descriptive. They characterize the analytic population contributing milk samples in each trial (for MISAME-III, the 276 of 290 dyads with a 14–21-day sample) and show baseline balance across randomized arms. As is standard for randomized trials, between-arm differences were not tested statistically.

```{r baseline-setup, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
library(tidyverse)
library(here)
library(table1)

raw_path <- here::here("data/merged_analysis_datasets.RDS")
have_baseline <- file.exists(raw_path)
if (!have_baseline) {
  cat("*The baseline data file (`data/merged_analysis_datasets.RDS`) is not present in this build. Rerun the analysis code (see README) to generate it.*\n\n")
  raw_df <- NULL
} else {
  raw_df <- readRDS(raw_path)

label(raw_df$sex)           <- "Infant sex"
label(raw_df$mage)          <- "Maternal age (years)"
label(raw_df$meducyrs)      <- "Maternal education (years)"
label(raw_df$mbmi)          <- "Maternal BMI (kg/m^2)"
label(raw_df$mmuaccm)       <- "Maternal MUAC (cm)"
if ("mhtcm" %in% names(raw_df)) label(raw_df$mhtcm) <- "Maternal height (cm)"
label(raw_df$parity)        <- "Parity"
label(raw_df$nperson)       <- "Household size (persons)"
label(raw_df$nrooms)        <- "Number of rooms in household"
label(raw_df$imp_water_src) <- "Improved water source"
label(raw_df$impfloor)      <- "Improved flooring"
label(raw_df$cookplac)      <- "Separate cooking space"
label(raw_df$dvseason)      <- "Season of delivery"
label(raw_df$dlvloc)        <- "Place of delivery"
label(raw_df$hhwealth)      <- "Household wealth index"
label(raw_df$hhfoodsecure)  <- "Household food security"
label(raw_df$arm)           <- "Intervention arm"
}

drop_empty_levels <- function(df) df %>% dplyr::mutate(dplyr::across(where(is.factor), droplevels))
```

## MISAME-III (Burkina Faso)

**Baseline characteristics: MISAME-III (Burkina Faso).** Maternal, infant, and household characteristics by randomized arm, at the earliest postnatal milk-collection visit.

```{r baseline-misame, echo=FALSE}
if (have_baseline) table1(~ sex + mage + meducyrs + mbmi + mmuaccm + parity + nperson + nrooms +
         imp_water_src + impfloor + cookplac + dvseason + dlvloc +
         hhwealth + hhfoodsecure | arm,
       data = raw_df %>% dplyr::filter(studyid == "MISAME-3", visit == 1) %>%
         drop_empty_levels(),
       transpose = FALSE)
```

## Mumta-LW (Pakistan)

**Baseline characteristics: Mumta-LW (Pakistan).** Maternal, infant, and household characteristics by randomized arm, at the earliest postnatal milk-collection visit.

```{r baseline-mumta, echo=FALSE}
if (have_baseline) table1(~ sex + mage + meducyrs + mbmi + mmuaccm + mhtcm + parity + nperson + nrooms +
         imp_water_src + impfloor + cookplac + dvseason + dlvloc +
         hhwealth + hhfoodsecure | arm,
       data = raw_df %>% dplyr::filter(studyid == "VITAL-Lactation", visit == 40) %>%
         drop_empty_levels(),
       transpose = FALSE)
```

## ELICIT (Tanzania)

**Baseline characteristics: ELICIT (Tanzania).** Maternal, infant, and household characteristics by randomized arm, at the earliest postnatal milk-collection visit.

```{r baseline-elicit, echo=FALSE}
if (have_baseline) table1(~ sex + mage + meducyrs + mbmi + mmuaccm + mhtcm + parity + nperson + nrooms +
         imp_water_src + impfloor + cookplac + dvseason + dlvloc +
         hhwealth + hhfoodsecure | arm,
       data = raw_df %>% dplyr::filter(studyid == "ELICIT", visit == 1) %>%
         drop_empty_levels(),
       transpose = FALSE)
```
