# Baseline characteristics by study and intervention arm

The IMiC subsamples analyzed here are drawn from three randomized controlled trials included in the intervention-effects analysis: **MISAME-III** (Burkina Faso), **Mumta-LW** (Pakistan), and **ELICIT** (Tanzania). Baseline maternal, infant, and household characteristics are presented separately by study and stratified by randomized intervention arm, using the earliest available postnatal visit corresponding to milk sample collection in each trial.

Tables report demographic, anthropometric, socioeconomic, and household variables that are plausibly associated with maternal nutritional status, breastfeeding context, or infant outcomes: infant sex, maternal age, education, height, BMI, MUAC, parity, household composition, housing characteristics, water source, season and place of delivery, asset-based household wealth, and food security. These variables were harmonized across studies during the IMiC data curation process and are documented in the manuscript's Supplementary Methods. The baseline summaries here are the **Section 1** content cited from the manuscript Methods. Because the printed-manuscript Supplementary Tables (S1–S6) cover pathway-enrichment results and Table S7 the native-unit intervention effects, no manuscript table number is assigned to the baseline characteristics.

These summaries serve a descriptive purpose: to characterize the analytic populations contributing milk samples in each trial and to assess baseline balance across intervention arms following randomization. Consistent with best practices for randomized trials, no statistical tests of between-arm differences are performed. The tables provide context for interpreting the intervention-effect estimates presented in subsequent chapters.

```{r baseline-setup, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
library(tidyverse)
library(here)
library(table1)

raw_path <- here::here("data/merged_analysis_datasets.RDS")
have_baseline <- file.exists(raw_path)
if (!have_baseline) {
  cat("*Baseline table data (`data/merged_analysis_datasets.RDS`) is not present in this build. Re-run the upstream analysis pipeline (see README) to populate it.*\n\n")
  raw_df <- NULL
} else {
  raw_df <- readRDS(raw_path)

label(raw_df$sex)           <- "Infant sex"
label(raw_df$mage)          <- "Maternal age (years)"
label(raw_df$meducyrs)      <- "Maternal education (years)"
label(raw_df$mbmi)          <- "Maternal BMI (kg/m^2)"
label(raw_df$mmuaccm)       <- "Maternal MUAC (cm)"
if ("mhtcm" %in% names(raw_df)) label(raw_df$mhtcm) <- "Maternal height (cm)"
label(raw_df$parity)        <- "Parity"
label(raw_df$nperson)       <- "Household size (persons)"
label(raw_df$nrooms)        <- "Number of rooms in household"
label(raw_df$imp_water_src) <- "Improved water source"
label(raw_df$impfloor)      <- "Improved flooring"
label(raw_df$cookplac)      <- "Separate cooking space"
label(raw_df$dvseason)      <- "Season of delivery"
label(raw_df$dlvloc)        <- "Place of delivery"
label(raw_df$hhwealth)      <- "Household wealth index"
label(raw_df$hhfoodsecure)  <- "Household food security"
label(raw_df$arm)           <- "Intervention arm"
}

drop_empty_levels <- function(df) df %>% dplyr::mutate(dplyr::across(where(is.factor), droplevels))
```

## MISAME-III (Burkina Faso)

**Baseline characteristics — MISAME-III (Burkina Faso).** Maternal, infant, and household characteristics by randomized arm, at the earliest postnatal milk-collection visit. No between-arm significance tests are reported (randomized trial).

```{r baseline-misame, echo=FALSE}
if (have_baseline) table1(~ sex + mage + meducyrs + mbmi + mmuaccm + parity + nperson + nrooms +
         imp_water_src + impfloor + cookplac + dvseason + dlvloc +
         hhwealth + hhfoodsecure | arm,
       data = raw_df %>% dplyr::filter(studyid == "MISAME-3", visit == 1) %>%
         drop_empty_levels(),
       transpose = FALSE)
```

## Mumta-LW (Pakistan)

**Baseline characteristics — Mumta-LW (Pakistan).** Maternal, infant, and household characteristics by randomized arm, at the earliest postnatal milk-collection visit. No between-arm significance tests are reported (randomized trial).

```{r baseline-mumta, echo=FALSE}
if (have_baseline) table1(~ sex + mage + meducyrs + mbmi + mmuaccm + mhtcm + parity + nperson + nrooms +
         imp_water_src + impfloor + cookplac + dvseason + dlvloc +
         hhwealth + hhfoodsecure | arm,
       data = raw_df %>% dplyr::filter(studyid == "VITAL-Lactation", visit == 40) %>%
         drop_empty_levels(),
       transpose = FALSE)
```

## ELICIT (Tanzania)

**Baseline characteristics — ELICIT (Tanzania).** Maternal, infant, and household characteristics by randomized arm, at the earliest postnatal milk-collection visit. No between-arm significance tests are reported (randomized trial).

```{r baseline-elicit, echo=FALSE}
if (have_baseline) table1(~ sex + mage + meducyrs + mbmi + mmuaccm + mhtcm + parity + nperson + nrooms +
         imp_water_src + impfloor + cookplac + dvseason + dlvloc +
         hhwealth + hhfoodsecure | arm,
       data = raw_df %>% dplyr::filter(studyid == "ELICIT", visit == 1) %>%
         drop_empty_levels(),
       transpose = FALSE)
```
