# Outcomes and human milk component distributions

This section has two parts. The first is the outcomes list cited in the manuscript Methods (*"A list of all outcomes and descriptions is available in Section 2 of the online resource"*), a searchable inventory of every targeted milk component (primary, secondary, and tertiary outcomes) in the harmonized IMiC dataset, grouped by prespecified outcome class and annotated with per-study availability. The second part shows the distribution of each component, beginning with manuscript Figure 4.

```{r distributions-setup, include=FALSE}
library(tidyverse)
library(here)
library(DT)
library(plotly)
source(here::here("functions.R"))
safe_load(here::here("metadata/milk_component.Rdata"))
## Reviewer flagged static plot labels as too small. Apply a larger base font
## to all new ggplots in this chapter and bump default fig dimensions.
ggplot2::theme_set(theme_imic_web(base_size = 14))
knitr::opts_chunk$set(fig.width = 14, fig.height = 11, dpi = 110)
```

## Outcomes list

The table below lists every targeted milk component (primary, secondary, and tertiary outcomes), grouped by the prespecified outcome class used throughout the manuscript (primary: macronutrients, micronutrients, B-vitamins; secondary: HMOs, bioactive proteins; tertiary: targeted metabolites from the Biocrates MxP Quant 500 panel). Per-study columns show which of the three IMiC trials measured each component.

```{r outcomes-dictionary, echo=FALSE}
build_dict <- function() {
  if (!exists("all_milk_components")) {
    cat("*`all_milk_components` was not available; metadata file missing or out of sync.*")
    return(invisible(NULL))
  }
  groups <- list(
    "Macronutrients (primary)"                = all_milk_components$macro,
    "Micronutrients (primary)"                = all_milk_components$micro,
    "B-vitamins (primary)"                    = all_milk_components$bvit,
    "Human milk oligosaccharides (secondary)" = all_milk_components$hmo,
    "Bioactive proteins (secondary)"          = all_milk_components$protein,
    "Targeted metabolites (tertiary)"         = all_milk_components$metabolomics
  )
  per_study <- list(
    "MISAME-III" = if (exists("misame_milk_components")) misame_milk_components else NULL,
    "Mumta-LW"   = if (exists("vital_milk_components"))  vital_milk_components  else NULL,
    "ELICIT"     = if (exists("elicit_milk_components")) elicit_milk_components else NULL
  )
  pooled <- dplyr::bind_rows(lapply(names(groups), function(g) {
    tibble::tibble(
      outcome_group = g,
      component     = unlist(groups[[g]], use.names = FALSE)
    )
  }))
  for (s in names(per_study)) {
    if (is.null(per_study[[s]])) next
    in_s <- unlist(per_study[[s]], use.names = FALSE)
    pooled[[paste0("measured_in_", s)]] <- pooled$component %in% in_s
  }
  pooled
}
dict <- build_dict()
if (!is.null(dict) && nrow(dict)) {
  nice_dt(dict, caption = "All milk components analyzed in this online resource, grouped by prespecified outcome class.")
}
```

Two subsets of this inventory are repeated in the sections that the manuscript figure legends name: the primary and secondary components cited in the Fig. 3A legend (*"metadata in Section 5 of the online resource"*) in [§5.1](#primary-component-metadata), and the full targeted-metabolite metadata cited in the Fig. 5A legend (*"metadata are in Section 6 of the online resource"*) in [§6.1](#targeted-metabolite-metadata).

## Distributions

The rest of this section shows unadjusted distributions: first violin plots for the 15 primary components in manuscript Fig. 4, by study, visit, and intervention arm, then boxplots of every measured component, one panel per component, with study and visit on the x-axis and boxes filled by arm. In the violin plots, gray and red reference lines mark the 50th and 10th percentiles of the age-specific MILQ reference distribution, and asterisks mark average treatment effects significant after FDR correction (\* *Q* < 0.05, \*\* *Q* < 0.01, \*\*\* *Q* < 0.001). Y-axes are log-transformed.

### MILQ-overlay density plots (Figure 4)

```{r distributions-figure-s2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=11, results='asis'}
p_s2 <- safe_readRDS(here::here("figure-data/figure_s2_plots.RDS"),
                     label = "figure_s2_plots.RDS")
show_any(p_s2)
```

**Figure 4.** IMiC distributions of HM nutrients relative to MILQ reference standards. Each panel shows one of 15 primary components (total energy, total fat, total protein, vitamins B~1~, B~2~, B~3~, B~5~, B~6~, B~7~, and B~12~, vitamin A, α- and γ-tocopherol, selenium, and copper) as violin plots of its native-unit concentration, with the horizontal line within each violin marking the median. Distributions are stratified by study (El: ELICIT, Mi: MISAME-III, Mu: Mumta-LW) and visit in months, given together in the x-axis labels, and by intervention arm: within each study × visit pair the control arm is the left, blue violin and the pooled intervention arms the right, orange violin. Horizontal gray and red lines mark the 50th and 10th percentiles of the age-specific MILQ reference distribution, respectively; asterisks mark average treatment effects significant after FDR correction (\* *Q* < 0.05, \*\* *Q* < 0.01, \*\*\* *Q* < 0.001). Y-axes are log-transformed.

### Per-component boxplots by study, visit, and arm

These boxplots show the raw concentration of each primary and secondary component on its native assay scale, one figure per outcome class, by study, milk-collection visit, and randomized arm. Boxes show the median and IQR, whiskers extend to 1.5 × IQR, and the y-axes are log-transformed. Boxes are filled by arm, in the color each randomized contrast has throughout this online resource (control in gray), and dodged in legend order. The 581 targeted metabolites are too many to plot one by one; their native-unit arm means and intervention effects are in `targeted_metabolites_native_units.csv` (§\@ref(native-units-tertiary)).

```{r distributions-boxplots-setup, include=FALSE}
raw_path <- here::here("data/merged_analysis_datasets.RDS")
box_ok <- file.exists(raw_path) && exists("all_milk_components")
box_df <- if (box_ok) readRDS(raw_path) else NULL
## The analyzed components of each class, labeled as in the results tables.
box_class <- function(csv) {
  d <- safe_read_csv(here::here("results/subsetted results", csv))
  if (is.null(d) || !box_ok) return(NULL)
  d <- unique(d[, c("biomarker", "label_f")])
  d <- d[d$biomarker %in% names(box_df), , drop = FALSE]
  stats::setNames(canonical_label(d$biomarker, d$label_f), d$biomarker)
}
box_classes <- list(macro = box_class("primary_macro.csv"), micro = box_class("primary_micro.csv"),
                    bvit = box_class("primary_bvit.csv"), hmo = box_class("secondary_hmo.csv"),
                    protein = box_class("secondary_bioactives.csv"))
box_height <- function(k) max(3.5, 2.4 * ceiling(length(box_classes[[k]]) / 4) + 1.4)
box_plot <- function(k) {
  labs_k <- box_classes[[k]]
  if (!box_ok || !length(labs_k)) {
    cat("*The dyad-level dataset is not available in this build; the boxplots are drawn when `data/merged_analysis_datasets.RDS` is present.*")
    return(invisible(NULL))
  }
  long <- box_df %>%
    dplyr::select(studyid, visit, arm, dplyr::all_of(names(labs_k))) %>%
    tidyr::pivot_longer(dplyr::all_of(names(labs_k)), names_to = "component", values_to = "value") %>%
    dplyr::filter(!is.na(value), value > 0) %>%
    dplyr::mutate(component = factor(labs_k[component], levels = unique(labs_k)),
                  sv = study_visit_panel(study_name(studyid), visit_name(studyid, visit), order = "study"))
  ## One fill per randomized arm (its contrast colour; control gray); boxes are
  ## dodged in legend order, so position is a second cue alongside colour.
  p <- ggplot2::ggplot(long, ggplot2::aes(x = sv, y = value, fill = arm)) +
    ggplot2::geom_boxplot(outlier.size = 0.4, alpha = 0.8, linewidth = 0.3) +
    ggplot2::scale_fill_manual(values = imic_arm_cols, na.value = "grey50", name = "Arm") +
    ggplot2::scale_y_log10() +
    ggplot2::facet_wrap(~ component, scales = "free_y", ncol = 4) +
    ggplot2::labs(x = NULL, y = "Concentration (native units, log scale)") +
    theme_imic_web() +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
  show_any(p)
}
```

#### Macronutrients

```{r distributions-box-macro, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=14, fig.height=box_height("macro")}
box_plot("macro")
```

#### Micronutrients (including fat-soluble vitamins)

```{r distributions-box-micro, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=14, fig.height=box_height("micro")}
box_plot("micro")
```

#### B-vitamins

```{r distributions-box-bvit, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=14, fig.height=box_height("bvit")}
box_plot("bvit")
```

#### Human milk oligosaccharides

```{r distributions-box-hmo, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=14, fig.height=box_height("hmo")}
box_plot("hmo")
```

#### Bioactive proteins

```{r distributions-box-protein, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=14, fig.height=box_height("protein")}
box_plot("protein")
```

**Figures (online-only).** Native-unit concentrations of the primary and secondary human-milk components by study, milk-collection visit, and randomized arm.
