# Outcomes and human milk component distributions

This section serves two purposes. The first half is the **outcomes list** referenced from the manuscript Methods (*"A list of all outcomes and descriptions is available in the Online Appendix (Section 2)"*): a searchable inventory of every milk component analysed in the IMiC harmonised dataset, grouped by prespecified outcome class and annotated with per-study availability. The second half (from *MILQ-overlay density plots* onward) is the descriptive **distribution** of each component, stratified by study, visit, and randomised intervention arm, and corresponds to **Figure S2** of the printed Word supplement.

Together these two views give the reader the full per-component context — *what was measured, where, and what the raw distribution looks like* — before encountering the modelled intervention-effect estimates in the subsequent sections.

```{r distributions-setup, include=FALSE}
library(tidyverse)
library(here)
library(DT)
library(plotly)
source(here::here("functions.R"))
safe_load(here::here("metadata/milk_component.Rdata"))
## Reviewer flagged static plot labels as too small. Apply a larger base font
## to all new ggplots in this chapter and bump default fig dimensions.
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 14))
knitr::opts_chunk$set(fig.width = 14, fig.height = 11, dpi = 110)
```

## Outcomes list

The table below enumerates every milk component analysed in this supplement, grouped by the prespecified outcome class used throughout the manuscript (primary: macronutrients, micronutrients, B-vitamins; secondary: HMOs, bioactive proteins; tertiary: targeted metabolites from the Biocrates MxP Quant 500 panel). Per-study availability indicators show in which of the three IMiC trials each component was measured.

Use the column filters at the top of the table to look up a specific component, or search by name in the upper-right.

```{r outcomes-dictionary, echo=FALSE}
build_dict <- function() {
  if (!exists("all_milk_components")) {
    cat("*`all_milk_components` was not available — metadata file missing or out of sync.*")
    return(invisible(NULL))
  }
  groups <- list(
    "Macronutrients (primary)"                = all_milk_components$macro,
    "Micronutrients (primary)"                = all_milk_components$micro,
    "B-vitamins (primary)"                    = all_milk_components$bvit,
    "Human milk oligosaccharides (secondary)" = all_milk_components$hmo,
    "Bioactive proteins (secondary)"          = all_milk_components$protein,
    "Targeted metabolites (tertiary)"         = all_milk_components$metabolomics
  )
  per_study <- list(
    "MISAME-III" = if (exists("misame_milk_components")) misame_milk_components else NULL,
    "Mumta-LW"   = if (exists("vital_milk_components"))  vital_milk_components  else NULL,
    "ELICIT"     = if (exists("elicit_milk_components")) elicit_milk_components else NULL
  )
  pooled <- dplyr::bind_rows(lapply(names(groups), function(g) {
    tibble::tibble(
      outcome_group = g,
      component     = unlist(groups[[g]], use.names = FALSE)
    )
  }))
  for (s in names(per_study)) {
    if (is.null(per_study[[s]])) next
    in_s <- unlist(per_study[[s]], use.names = FALSE)
    pooled[[paste0("measured_in_", s)]] <- pooled$component %in% in_s
  }
  pooled
}
dict <- build_dict()
if (!is.null(dict) && nrow(dict)) {
  nice_dt(dict, caption = "All milk components analysed in this supplement, grouped by prespecified outcome class.")
}
```

The full **targeted-metabolite metadata** (Biocrates panel categories, with per-study availability) is reproduced as a focused subsection in [§6 Reductions in HM nutrient deficiency](#reductions-in-hm-nutrient-deficiency-relative-to-milq-reference-standards) so that the metadata called out in the manuscript Fig. 5A legend (*"metadata in Online Supplement 6"*) lives at that section's anchor.

## Distributions

The remainder of this section shows the unadjusted distribution of each measured human milk component, stratified by study, visit, and randomized intervention arm. Two views are provided: (i) density plots overlaid by intervention arm, and (ii) per-component boxplots faceted by study and visit. Density plots include grey and red reference lines at the 50th and 10th percentiles of the MILQ age-specific reference distribution where applicable. Asterisks indicate FDR-corrected statistical significance of the pooled intervention vs. control contrast for that component–visit (\\* q\\<0.05, \\*\\* q\\<0.01, \\*\\*\\* q\\<0.001). Y-axes are log-transformed where the underlying distribution is markedly right-skewed (most micronutrients and metabolites).

### MILQ-overlay density plots (Figure S2)

```{r distributions-figure-s2, echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=11}
p_s2 <- safe_readRDS(here::here("figure-data/figure_s2_plots.RDS"),
                     label = "figure_s2_plots.RDS")
show_any(p_s2)
```

### Per-component boxplots by study, visit, and arm

These boxplots summarize the raw concentration of each component on its native assay scale. Hover (in HTML output) to see the median, IQR, and sample size for each box.

```{r distributions-boxplots, echo=FALSE, message=FALSE, warning=FALSE, fig.width=14, fig.height=11}
raw_path <- here::here("data/merged_analysis_datasets.RDS")
metadata_loaded <- safe_load(here::here("metadata/milk_component.Rdata"))

if (file.exists(raw_path) && isTRUE(metadata_loaded)) {
  df <- readRDS(raw_path)
  comps <- unique(unlist(all_milk_components, use.names = FALSE))
  comps <- intersect(comps, names(df))
  if (length(comps) == 0) {
    cat("*No matching milk-component columns were found in the merged analysis dataset.*")
  } else {
    long <- df %>%
      dplyr::select(studyid, visit, arm, dplyr::all_of(comps)) %>%
      tidyr::pivot_longer(dplyr::all_of(comps),
                          names_to = "component", values_to = "value") %>%
      dplyr::filter(!is.na(value))
    p <- ggplot2::ggplot(long,
                         ggplot2::aes(x = factor(visit), y = value, fill = arm)) +
      ggplot2::geom_boxplot(outlier.size = 0.5, alpha = 0.7) +
      ggplot2::scale_y_log10() +
      ggplot2::facet_wrap(~ component, scales = "free_y") +
      ggplot2::labs(x = "Visit", y = "Concentration (native units, log scale)",
                    fill = "Arm") +
      ggplot2::theme_minimal() +
      ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1),
                     legend.position = "bottom")
    show_any(p)
  }
} else {
  cat("*Raw IMiC dyad-level dataset and/or component metadata not available in this build; boxplots will be rendered when `data/merged_analysis_datasets.RDS` and `metadata/milk_component.Rdata` are present.*")
}
```
