# Outcomes and human milk component distributions

```{r distributions-setup, include=FALSE}
library(tidyverse)
library(here)
library(DT)
library(plotly)
source(here::here("functions.R"))
safe_load(here::here("metadata/milk_component.Rdata"))
## Reviewer flagged static plot labels as too small. Apply a larger base font
## to all new ggplots in this chapter and bump default fig dimensions.
ggplot2::theme_set(theme_imic_web(base_size = 14))
knitr::opts_chunk$set(fig.width = 14, fig.height = 11, dpi = 110)
```

## Outcomes list

All primary, secondary, and tertiary milk components, by outcome class, with the trials that measured each.

```{r outcomes-dictionary, echo=FALSE}
build_dict <- function() {
  if (!exists("all_milk_components")) {
    cat("*`all_milk_components` was not available; metadata file missing or out of sync.*")
    return(invisible(NULL))
  }
  groups <- list(
    "Macronutrients (primary)"                = all_milk_components$macro,
    "Micronutrients (primary)"                = all_milk_components$micro,
    "B-vitamins (primary)"                    = all_milk_components$bvit,
    "Human milk oligosaccharides (secondary)" = all_milk_components$hmo,
    "Bioactive proteins (secondary)"          = all_milk_components$protein,
    "Targeted metabolites (tertiary)"         = all_milk_components$metabolomics
  )
  per_study <- list(
    "MISAME-III" = if (exists("misame_milk_components")) misame_milk_components else NULL,
    "Mumta-LW"   = if (exists("vital_milk_components"))  vital_milk_components  else NULL,
    "ELICIT"     = if (exists("elicit_milk_components")) elicit_milk_components else NULL
  )
  pooled <- dplyr::bind_rows(lapply(names(groups), function(g) {
    tibble::tibble(
      outcome_group = g,
      component     = unlist(groups[[g]], use.names = FALSE)
    )
  }))
  for (s in names(per_study)) {
    if (is.null(per_study[[s]])) next
    in_s <- unlist(per_study[[s]], use.names = FALSE)
    pooled[[paste0("measured_in_", s)]] <- pooled$component %in% in_s
  }
  pooled
}
dict <- build_dict()
if (!is.null(dict) && nrow(dict)) {
  nice_dt(dict, caption = "All milk components analyzed in this online resource, grouped by prespecified outcome class.")
}
```

## Distributions

Raw concentrations of each primary and secondary component in native units, by trial, visit, and randomized arm. Boxes show the median and interquartile range, whiskers extend to 1.5 times the interquartile range, and y-axes are on a log scale. Native-unit means for the targeted metabolites are in Section \@ref(native-units-tertiary).

```{r distributions-boxplots-setup, include=FALSE}
raw_path <- here::here("data/merged_analysis_datasets.RDS")
box_ok <- file.exists(raw_path) && exists("all_milk_components")
box_df <- if (box_ok) readRDS(raw_path) else NULL
## The analyzed components of each class, labeled as in the results tables.
box_class <- function(csv) {
  d <- safe_read_csv(here::here("results/subsetted results", csv))
  if (is.null(d) || !box_ok) return(NULL)
  d <- unique(d[, c("biomarker", "label_f")])
  d <- d[d$biomarker %in% names(box_df), , drop = FALSE]
  stats::setNames(canonical_label(d$biomarker, d$label_f), d$biomarker)
}
box_classes <- list(macro = box_class("primary_macro.csv"), micro = box_class("primary_micro.csv"),
                    bvit = box_class("primary_bvit.csv"), hmo = box_class("secondary_hmo.csv"),
                    protein = box_class("secondary_bioactives.csv"))
box_height <- function(k) max(3.5, 2.4 * ceiling(length(box_classes[[k]]) / 4) + 1.4)
box_plot <- function(k) {
  labs_k <- box_classes[[k]]
  if (!box_ok || !length(labs_k)) {
    cat("*The dyad-level dataset is not available in this build; the boxplots are drawn when `data/merged_analysis_datasets.RDS` is present.*")
    return(invisible(NULL))
  }
  long <- box_df %>%
    dplyr::select(studyid, visit, arm, dplyr::all_of(names(labs_k))) %>%
    tidyr::pivot_longer(dplyr::all_of(names(labs_k)), names_to = "component", values_to = "value") %>%
    dplyr::filter(!is.na(value), value > 0) %>%
    dplyr::mutate(component = factor(labs_k[component], levels = unique(labs_k)),
                  sv = study_visit_panel(study_name(studyid), visit_name(studyid, visit), order = "study"))
  ## One fill per randomized arm (its contrast colour; control gray); boxes are
  ## dodged in legend order, so position is a second cue alongside colour.
  p <- ggplot2::ggplot(long, ggplot2::aes(x = sv, y = value, fill = arm)) +
    ggplot2::geom_boxplot(outlier.size = 0.4, alpha = 0.8, linewidth = 0.3) +
    ggplot2::scale_fill_manual(values = imic_arm_cols, na.value = "grey50", name = "Arm") +
    ggplot2::scale_y_log10() +
    ggplot2::facet_wrap(~ component, scales = "free_y", ncol = 4) +
    ggplot2::labs(x = NULL, y = "Concentration (native units, log scale)") +
    theme_imic_web() +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
  show_any(p)
}
```

### Macronutrients

```{r distributions-box-macro, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=14, fig.height=box_height("macro")}
box_plot("macro")
```

### Micronutrients (including fat-soluble vitamins)

```{r distributions-box-micro, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=14, fig.height=box_height("micro")}
box_plot("micro")
```

### B-vitamins

```{r distributions-box-bvit, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=14, fig.height=box_height("bvit")}
box_plot("bvit")
```

### Human milk oligosaccharides

```{r distributions-box-hmo, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=14, fig.height=box_height("hmo")}
box_plot("hmo")
```

### Bioactive proteins

```{r distributions-box-protein, echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=14, fig.height=box_height("protein")}
box_plot("protein")
```
