# Subgroup analyses stratified by maternal BMI

This section examines whether intervention effects on human milk components depended on maternal nutritional status at baseline, using maternal body mass index (BMI) as a marker of that status, dichotomized into *underweight* (BMI ≤ 18.5 kg/m²) and *normal or overweight* (BMI > 18.5 kg/m²) strata. We estimated stratum-specific ATEs with the same TMLE framework used in the primary adjusted analyses. Interaction *P*-values come from a Wald-type test of the difference between the two stratum-specific estimates, divided by the standard error of the difference. The results shown here are the component × study × visit combinations with a nominally significant interaction (all interaction *P* < 0.01): 129 in Mumta-LW (1.5 and 2 months) and one in MISAME-III (3–4 months); six, all in Mumta-LW, remain significant after FDR correction.

The manuscript cites this section as "Section 3 of the online resource" in the Results paragraph on macronutrient and micronutrient effects, where it reports that maternal BMI did not consistently modify macronutrient or micronutrient effects.

> These subgroup analyses are exploratory and not powered to detect modest interactions, particularly in studies with a small underweight stratum. Differences in the size or direction of per-component estimates between BMI strata should be interpreted cautiously, especially when the 95% CIs are wide.

```{r bmi-setup, include=FALSE}
library(tidyverse)
library(here)
library(DT)
library(plotly)
source(here::here("functions.R"))

stratified_forest <- function(tab, title = "BMI-stratified intervention effects") {
  tab <- .harmonize_factors(tab)
  if (!nrow(tab)) { cat("*No rows in this outcome group.*"); return(invisible(NULL)) }
  tab$tooltip <- paste0(tab$label_f,
                        "<br>BMI stratum: ", tab$subgroup,
                        "<br>Study: ", tab$study, " | Visit: ", tab$visit,
                        if ("contrast" %in% names(tab)) paste0("<br>Contrast: ", tab$contrast) else "",
                        "<br>ATE: ", round(tab$est, 3),
                        " (", round(tab$cil, 3), ", ", round(tab$ciu, 3), ")",
                        if ("chi_pval_adj" %in% names(tab)) paste0("<br>Q: ", signif(tab$chi_pval_adj, 3)) else "",
                        if ("interaction_pvalue" %in% names(tab)) paste0("<br>Interaction P: ", signif(tab$interaction_pvalue, 3)) else "")
  tab$panel <- study_visit_panel(tab$study, tab$visit)
  ## BMI stratum is carried by colour, symbol shape, and CI line type.
  strata <- sort(unique(as.character(tab$subgroup)))
  p <- ggplot2::ggplot(tab,
                       ggplot2::aes(y = stats::reorder(label_f, -est), x = est,
                                    color = subgroup, shape = subgroup, text = tooltip)) +
    ggplot2::geom_point(position = ggplot2::position_dodge(width = 0.6), size = 2) +
    ggplot2::geom_errorbar(ggplot2::aes(xmin = cil, xmax = ciu, linetype = subgroup),
                          width = 0.2, orientation = "y",
                          position = ggplot2::position_dodge(width = 0.6)) +
    ggplot2::geom_vline(xintercept = 0, linetype = "dashed", color = "grey60") +
    ggplot2::scale_color_manual(values = stats::setNames(
      rep_len(unname(okabe_ito[c("blue", "vermillion", "green", "purple")]), length(strata)), strata)) +
    ggplot2::scale_shape_manual(values = stats::setNames(rep_len(c(16, 17, 15, 18), length(strata)), strata)) +
    ggplot2::scale_linetype_manual(values = stats::setNames(
      rep_len(c("solid", "dashed", "dotted", "dotdash"), length(strata)), strata)) +
    ggplot2::facet_wrap(~ panel) +
    ggplot2::labs(title = title, y = NULL, x = "Average treatment effect (SD units)",
                  color = "Maternal BMI stratum", shape = "Maternal BMI stratum",
                  linetype = "Maternal BMI stratum") +
    theme_imic_web()
  as_widget(p)
}

bmi_df <- safe_readRDS(here::here("figure-data/subgroup_results.RDS"),
                       label = "subgroup_results.RDS")
```

## Per-component stratified forest plots

**Figure (online-only).** Adjusted ATEs (SD units, with 95% CI), for component × study × visit combinations with a nominally significant interaction, within maternal-BMI strata (underweight, BMI ≤ 18.5 kg/m²; normal/overweight, BMI > 18.5 kg/m²), by study and milk-collection visit, shown separately for primary, secondary, and tertiary outcomes. BMI strata differ in color, symbol shape, and CI line type (normal/overweight: blue circles, solid lines; underweight: vermillion triangles, dashed lines). In the web version, hovering over a point shows the stratum-specific estimate, *Q*-value, and Wald-type interaction *P*-value.

```{r bmi-primary, echo=FALSE, message=FALSE, warning=FALSE, fig.height=14}
if (!is.null(bmi_df) && is.data.frame(bmi_df)) {
  primary <- bmi_df %>%
    dplyr::filter(outcome_group %in% c("primary","Primary","Macronutrients","Micronutrients","B-vitamins"))
  stratified_forest(primary,
                    title = "Primary outcomes (macro / micro / B-vitamins): BMI-stratified ATEs")
}
```

```{r bmi-secondary, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12}
if (!is.null(bmi_df) && is.data.frame(bmi_df)) {
  sec <- bmi_df %>%
    dplyr::filter(outcome_group %in% c("secondary","Secondary","HMOs","Bioactive proteins"))
  stratified_forest(sec,
                    title = "Secondary outcomes (HMOs and bioactive proteins): BMI-stratified ATEs")
}
```

```{r bmi-tertiary, echo=FALSE, message=FALSE, warning=FALSE, fig.height=14}
if (!is.null(bmi_df) && is.data.frame(bmi_df)) {
  tert <- bmi_df %>%
    dplyr::filter(outcome_group %in% c("tertiary","Tertiary","Targeted metabolites"))
  stratified_forest(tert,
                    title = "Tertiary outcomes (targeted metabolites): BMI-stratified ATEs")
}
```

## Searchable table of all BMI-stratified estimates

```{r bmi-table, echo=FALSE, message=FALSE, warning=FALSE}
if (!is.null(bmi_df) && is.data.frame(bmi_df)) {
  tbl <- bmi_df %>%
    dplyr::select(dplyr::any_of(c(
      "study","visit","subgroup","contrast","outcome_group","category",
      "label_f","est","cil","ciu","chi_pval","chi_pval_adj",
      "interaction_pvalue","interaction_pvalue_adj","sig","sigFDR"
    ))) %>%
    dplyr::mutate(dplyr::across(where(is.numeric), ~ round(.x, 3)))
  nice_dt(tbl,
          caption = "BMI-stratified ATEs and Wald-type interaction P-values for the component × study × visit combinations with a nominally significant interaction.")
}
```
