# Subgroup analyses stratified by maternal BMI

This chapter presents subgroup analyses evaluating whether the effects of maternal nutritional interventions on human milk components differed by maternal nutritional status at baseline, operationalized using **maternal body mass index (BMI)**. BMI was dichotomized a priori into *underweight* (BMI ≤ 18.5 kg/m²) and *normal or overweight* (BMI > 18.5 kg/m²) strata. Stratum-specific ATEs were estimated using the same TMLE framework as in the primary adjusted analyses, and **interaction p-values** were computed by comparing stratum-specific estimates with a Wald-type test on the difference in estimates scaled by the pooled standard error.

This section is cited as **"Online Appendix 3"** in the main manuscript (Results paragraphs on macronutrient and micronutrient effects). The manuscript reports **no consistent evidence of effect modification by maternal BMI** for primary, secondary, or tertiary outcomes. The plots and tables below allow that conclusion to be verified at the per-component level.

> **Interpretation note.** These subgroup analyses are exploratory and not powered to detect modest interactions, particularly within studies where the underweight stratum is small. Per-component point estimates that differ in magnitude or direction across BMI strata should be interpreted cautiously, especially where the corresponding 95% CIs are wide.

```{r bmi-setup, include=FALSE}
library(tidyverse)
library(here)
library(DT)
library(plotly)
source(here::here("functions.R"))

stratified_forest <- function(tab, title = "BMI-stratified intervention effects") {
  tab <- .harmonize_factors(tab)
  if (!nrow(tab)) { cat("*No rows in this outcome group.*"); return(invisible(NULL)) }
  tab$tooltip <- paste0(tab$label_f,
                        "<br>BMI stratum: ", tab$subgroup,
                        "<br>Study: ", tab$study, " | Visit: ", tab$visit,
                        if ("contrast" %in% names(tab)) paste0("<br>Contrast: ", tab$contrast) else "",
                        "<br>ATE: ", round(tab$est, 3),
                        " (", round(tab$cil, 3), ", ", round(tab$ciu, 3), ")",
                        if ("chi_pval_adj" %in% names(tab)) paste0("<br>q: ", signif(tab$chi_pval_adj, 3)) else "",
                        if ("interaction_pvalue" %in% names(tab)) paste0("<br>p_interaction: ", signif(tab$interaction_pvalue, 3)) else "")
  tab$panel <- factor(paste0(tab$study, " — ", tab$visit))
  p <- ggplot2::ggplot(tab,
                       ggplot2::aes(y = stats::reorder(label_f, est), x = est,
                                    color = subgroup, text = tooltip)) +
    ggplot2::geom_point(position = ggplot2::position_dodge(width = 0.6), size = 2) +
    ggplot2::geom_errorbar(ggplot2::aes(xmin = cil, xmax = ciu),
                          width = 0.2, orientation = "y",
                          position = ggplot2::position_dodge(width = 0.6)) +
    ggplot2::geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
    ggplot2::scale_color_manual(values = tableau10[c(1, 2, 3, 4)]) +
    ggplot2::facet_wrap(~ panel) +
    ggplot2::labs(title = title, y = "Outcome", x = "Estimate (SD units)",
                  color = "Maternal BMI stratum") +
    ggplot2::theme_minimal() + ggplot2::theme(legend.position = "bottom")
  if (knitr::is_html_output() && requireNamespace("plotly", quietly = TRUE)) {
    conv <- try(plotly::ggplotly(p, tooltip = "text"), silent = TRUE)
    if (!inherits(conv, "try-error")) return(conv)
  }
  p
}

bmi_df <- safe_readRDS(here::here("figure-data/subgroup_results.RDS"),
                       label = "subgroup_results.RDS")
```

## Per-component stratified forest plots

```{r bmi-primary, echo=FALSE, message=FALSE, warning=FALSE, fig.height=14}
if (!is.null(bmi_df) && is.data.frame(bmi_df)) {
  primary <- bmi_df %>%
    dplyr::filter(outcome_group %in% c("primary","Primary","Macronutrients","Micronutrients","B-vitamins"))
  stratified_forest(primary,
                    title = "Primary outcomes (macro / micro / B-vitamins) — BMI-stratified ATEs")
}
```

```{r bmi-secondary, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12}
if (!is.null(bmi_df) && is.data.frame(bmi_df)) {
  sec <- bmi_df %>%
    dplyr::filter(outcome_group %in% c("secondary","Secondary","HMOs","Bioactive proteins"))
  stratified_forest(sec,
                    title = "Secondary outcomes (HMOs and bioactive proteins) — BMI-stratified ATEs")
}
```

```{r bmi-tertiary, echo=FALSE, message=FALSE, warning=FALSE, fig.height=14}
if (!is.null(bmi_df) && is.data.frame(bmi_df)) {
  tert <- bmi_df %>%
    dplyr::filter(outcome_group %in% c("tertiary","Tertiary","Targeted metabolites"))
  stratified_forest(tert,
                    title = "Tertiary outcomes (targeted metabolites) — BMI-stratified ATEs")
}
```

## Searchable table of all BMI-stratified estimates

```{r bmi-table, echo=FALSE, message=FALSE, warning=FALSE}
if (!is.null(bmi_df) && is.data.frame(bmi_df)) {
  tbl <- bmi_df %>%
    dplyr::select(dplyr::any_of(c(
      "study","visit","subgroup","contrast","outcome_group","category",
      "label_f","est","cil","ciu","chi_pval","chi_pval_adj",
      "interaction_pvalue","interaction_pvalue_adj","sig","sigFDR"
    ))) %>%
    dplyr::mutate(dplyr::across(where(is.numeric), ~ round(.x, 3)))
  nice_dt(tbl,
          caption = "BMI-stratified ATEs and Wald-type interaction p-values for every analyzed milk component.")
}
```
