# Exploratory high-dimensional outcomes: microbiome, untargeted proteomics, untargeted metabolomics

This section covers human milk components that were designated as exploratory outcomes: the milk microbiome (α-diversity and individual taxa), untargeted proteomics, and untargeted metabolomics.

```{r explor-setup, include=FALSE}
library(tidyverse)
library(here)
library(DT)
library(plotly)
source(here::here("functions.R"))

keep_ate <- function(df) {
  if (is.null(df)) return(NULL)
  if ("measure" %in% names(df)) df <- dplyr::filter(df, measure == "ATE")
  df
}

## Wrapper: read, filter to ATE, attach label_f, alias chi_pval/chi_pval_adj.
prep_explor <- function(rds_path) {
  df <- safe_readRDS(here::here(rds_path), label = basename(rds_path))
  if (is.null(df)) return(NULL)
  if ("res" %in% names(df)) df <- unnest_imic_results(df)
  df <- keep_ate(df)
  if (is.null(df) || !nrow(df)) return(NULL)
  if (!"label_f" %in% names(df))       df$label_f  <- df$biomarker
  if (!"pval" %in% names(df) && "chi_pval" %in% names(df))         df$pval     <- df$chi_pval
  if (!"pval_adj" %in% names(df) && "chi_pval_adj" %in% names(df)) df$pval_adj <- df$chi_pval_adj
  df
}
```

## Microbiome

### α-diversity

Effects on z-scored observed richness and Shannon diversity, pooled and then by randomized arm.

```{r microbiome-div-combined, echo=FALSE, message=FALSE, warning=FALSE}
df <- prep_explor("results/microbiome_diversity_intervention_effects_results.RDS")
if (!is.null(df)) clean_tab(df, caption = "Pooled intervention vs. control.")
```

```{r microbiome-div-strat, echo=FALSE, message=FALSE, warning=FALSE}
df <- prep_explor("results/microbiome_diversity_intervention_effects_results_arm_strat.RDS")
if (!is.null(df)) clean_tab(df, caption = "Each randomized arm vs. control.")
```

### Individual microbial features

Per-taxon effects, pooled intervention vs. control.

```{r microbiome-vol-combined, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7, results='asis'}
df <- prep_explor("results/microbiome_intervention_effects_results.RDS")
if (!is.null(df)) {
  p <- volcano_capped(df, title = "Microbial features")
  print(htmltools::tagList(p, nice_dt(df %>% dplyr::mutate(dplyr::across(where(is.numeric), ~ signif(.x, 3))),
          caption = "Per-taxon effects, pooled intervention vs. control.")))
}
```

## Untargeted proteomics

Per-protein effects, pooled intervention vs. control, in the dyads with proteomics. ELICIT is not included because its proteome samples came only from the control and azithromycin arms.

```{r prot-vol, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7, results='asis'}
df <- prep_explor("results/adjusted_combined_arms_intervention_effects_proteomics_results_clean.RDS")
if (!is.null(df)) {
  p <- volcano_capped(df, title = "Untargeted proteome")
  print(htmltools::tagList(
    p,
    nice_dt(df %>% dplyr::select(dplyr::any_of(c("study","visit","contrast","label_f","biomarker",
                                                  "est","cil","ciu","pval","pval_adj","sig","sigFDR"))) %>%
              dplyr::mutate(dplyr::across(where(is.numeric), ~ signif(.x, 3))),
            caption = "Per-protein effects, pooled intervention vs. control.")
  ))
}
```

## Untargeted metabolomics

Per-feature effects, pooled intervention vs. control, for about 97,000 features across trials, visits, and ionization modes.

```{r met-vol, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7, results='asis'}
df <- prep_explor("results/adjusted_combined_arms_intervention_effects_untargeted_results_clean.RDS")
if (!is.null(df)) {
  p <- volcano_capped(df, n_show = 5000,
                      table_note = "the table lists the features with Q < 0.10",
                      title = "Untargeted metabolome")
  ## The full ~1M-row table is far too large to render as DT; expose the
  ## Q < 0.10 subset only.
  df_sig <- df %>% dplyr::filter(!is.na(pval_adj), pval_adj < 0.1) %>%
    dplyr::mutate(dplyr::across(where(is.numeric), ~ signif(.x, 3)))
  print(htmltools::tagList(
    p,
    nice_dt(df_sig %>% dplyr::select(dplyr::any_of(c(
      "study","visit","contrast","label_f","biomarker","est","cil","ciu",
      "pval","pval_adj","sig","sigFDR"))), page_length = 25,
      caption = sprintf("The %s features with Q < 0.10, of %s estimates.",
                        format(nrow(df_sig), big.mark = ","), format(nrow(df), big.mark = ",")))
  ))
}
```
