# Exploratory high-dimensional outcomes: microbiome, untargeted proteomics, untargeted metabolomics

The analyses collected here cover human milk components that were not designated as primary or secondary outcomes: the milk microbiome (α-diversity and individual taxa), untargeted proteomics, and untargeted metabolomics. They are meant to survey where intervention-related signals might exist across these high-dimensional modalities and to generate hypotheses for follow-up.

Every result here warrants caution. The estimates carry a heavy multiple-testing burden and variable measurement error, and in some cases rest on limited sample sizes or incomplete intervention contrasts. Formal inference and primary conclusions come from the prespecified primary and secondary outcomes in the *Adjusted intervention effects* chapter. Pathway-level interpretation of the metabolomic and proteomic signals, which gives a far more readable view of the biology, appears in the *Pathway enrichment* chapter.

The forest and volcano plots in this chapter render as **interactive plotly figures** in the HTML build: hover any point to read the compound or taxon name, study, visit, ATE, and FDR-adjusted q-value. A searchable table accompanies each plot, offering the same data with CSV/Excel download. For the high-cardinality modalities (untargeted metabolomics: ~97,000 features × studies × visits), the interactive plot shows only the top hits by FDR-adjusted p-value so the page stays responsive, while the downloadable table contains every feature.

The microbiome individual-taxa analyses in the first half of this section are cited as **"Online Appendix 4"** in the main manuscript (Microbiome subsection). The section also includes the data underlying **Figure S6 (microbiome α-diversity)** of the printed Word supplement, plus exploratory volcano plots for untargeted proteomics and untargeted metabolomics referenced from later results paragraphs.

```{r explor-setup, include=FALSE}
library(tidyverse)
library(here)
library(DT)
library(plotly)
source(here::here("functions.R"))

keep_ate <- function(df) {
  if (is.null(df)) return(NULL)
  if ("measure" %in% names(df)) df <- dplyr::filter(df, measure == "ATE")
  df
}

## Wrapper: read, filter to ATE, attach label_f, alias chi_pval/chi_pval_adj.
prep_explor <- function(rds_path) {
  df <- safe_readRDS(here::here(rds_path), label = basename(rds_path))
  if (is.null(df)) return(NULL)
  if ("res" %in% names(df) && !is.data.frame(df$res[[1]] %||% NULL)) {
    df <- unnest_imic_results(df)
  } else if ("res" %in% names(df)) {
    df <- unnest_imic_results(df)
  }
  df <- keep_ate(df)
  if (is.null(df) || !nrow(df)) return(NULL)
  if (!"label_f" %in% names(df))       df$label_f  <- df$biomarker
  if (!"pval" %in% names(df) && "chi_pval" %in% names(df))         df$pval     <- df$chi_pval
  if (!"pval_adj" %in% names(df) && "chi_pval_adj" %in% names(df)) df$pval_adj <- df$chi_pval_adj
  df
}
```

## Microbiome

### Intervention effects on microbiome α-diversity (Figure S6)

Forest plots display ATEs on z-scored α-diversity metrics (observed richness and Shannon diversity), pooled across all intervention arms and then stratified by randomized arm. Across studies and visits, **no consistent intervention-related changes in microbiome diversity were observed**, apart from a transient increase in both indices at the first Mumta-LW visit. Estimated effects varied in size and direction across timepoints and studies, and confidence intervals generally overlapped the null.

#### Combined intervention arms

```{r microbiome-div-combined, echo=FALSE, message=FALSE, warning=FALSE, fig.height=6}
df <- prep_explor("results/microbiome_diversity_intervention_effects_results.RDS")
if (!is.null(df)) htmltools::tagList(
  forest_plot(df, title = "Microbiome α-diversity — pooled"),
  htmltools::tags$p(htmltools::tags$strong("Figure S6. "),
    htmltools::tags$em("Maternal nutritional intervention effects on human-milk microbiota alpha diversity (observed richness and Shannon diversity), Z-scored for cross-study comparability of intervention effects, shown by study, visit, and intervention contrast (pooled intervention vs. control). Filled symbols are significant after FDR correction. Individual microbial taxa results are shown later in this chapter (cited from the manuscript as Online Appendix 4 / Section 4).")),
  clean_tab(df, caption = "Adjusted ATEs on milk-microbiome α-diversity metrics, pooled intervention vs. control (manuscript Fig. S6)."))
```

#### Stratified by randomized arm

```{r microbiome-div-strat, echo=FALSE, message=FALSE, warning=FALSE, fig.height=6}
df <- prep_explor("results/microbiome_diversity_intervention_effects_results_arm_strat.RDS")
if (!is.null(df)) htmltools::tagList(forest_plot(df, arm_strat = TRUE,
                                                 title = "Microbiome α-diversity — per arm"),
                                     clean_tab(df, caption = "Adjusted ATEs on milk-microbiome α-diversity metrics, by randomized arm vs. control."))
```

### Intervention effects on individual microbial features

The volcano plots below summarize per-taxon (or per-ASV) intervention effects. In every analysis, the distribution of effect estimates and p-values is consistent with a global null hypothesis: no taxa show reproducible intervention-associated shifts across studies or visits.

#### Combined intervention arms

```{r microbiome-vol-combined, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7, results='asis'}
df <- prep_explor("results/microbiome_intervention_effects_results.RDS")
if (!is.null(df)) {
  p <- volcano_capped(df, title = "Microbial features — pooled")
  print(htmltools::tagList(p, nice_dt(df %>% dplyr::mutate(dplyr::across(where(is.numeric), ~ signif(.x, 3))),
          caption = "Per-taxon adjusted intervention effects on milk-microbiome features, pooled intervention vs. control.")))
}
```

#### Stratified by randomized arm

```{r microbiome-vol-strat, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7, results='asis'}
df <- prep_explor("results/microbiome_intervention_effects_results_arm_strat.RDS")
if (!is.null(df)) {
  p <- volcano_capped(df, title = "Microbial features — per arm")
  print(htmltools::tagList(p, nice_dt(df %>% dplyr::mutate(dplyr::across(where(is.numeric), ~ signif(.x, 3))),
          caption = "Per-taxon adjusted intervention effects on milk-microbiome features, by randomized arm vs. control.")))
}
```

## Untargeted proteomics

Untargeted proteomic analyses were run in a subset of mother–infant dyads and carried important design constraints. Data were available for a limited number of samples, and in the ELICIT trial they were collected only in the control arm, which ruled out formal estimation of randomized intervention effects for ELICIT. The volcano plot below is descriptive: the share of nominally significant protein-level associations matches the false-positive rate expected under a global null distribution. **GO Biological Process enrichment** of the directional protein lists appears in the *Pathway enrichment* chapter.

```{r prot-vol, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7, results='asis'}
df <- prep_explor("results/adjusted_combined_arms_intervention_effects_proteomics_results_clean.RDS")
if (!is.null(df)) {
  p <- volcano_capped(df, title = "Untargeted proteomics — pooled intervention vs. control")
  print(htmltools::tagList(
    p,
    nice_dt(df %>% dplyr::select(dplyr::any_of(c("study","visit","contrast","label_f","biomarker",
                                                  "est","cil","ciu","pval","pval_adj","sig","sigFDR"))) %>%
              dplyr::mutate(dplyr::across(where(is.numeric), ~ signif(.x, 3))),
            caption = "Per-protein adjusted intervention effects on the untargeted milk proteome, pooled BEP vs. control (descriptive; ELICIT proteomics were control-arm only).")
  ))
}
```

## Untargeted metabolomics

The volcano below displays effect estimates and significance for the full untargeted metabolomics feature set (≈97,000 features across studies, visits, and ionization modes). Relative to the microbiome and proteomics analyses, untargeted metabolomics shows a higher density of features with nominal evidence of intervention association, though only a small fraction survives FDR correction. The pattern reads more clearly at the **pathway level** (MetaboAnalyst ORA and Mummichog) in the *Pathway enrichment* chapter and is summarized in the manuscript's Figure 6 (panels A–B).

```{r met-vol, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7, results='asis'}
df <- prep_explor("results/adjusted_combined_arms_intervention_effects_untargeted_results_clean.RDS")
if (!is.null(df)) {
  p <- volcano_capped(df, n_show = 5000,
                      title = "Untargeted metabolomics — pooled intervention vs. control")
  ## The full ~1M-row table is far too large to render as DT; expose the
  ## FDR-significant subset only.
  df_sig <- df %>% dplyr::filter(!is.na(pval_adj), pval_adj < 0.1) %>%
    dplyr::mutate(dplyr::across(where(is.numeric), ~ signif(.x, 3)))
  cat(sprintf("\n\n*Table below shows the %s metabolite-feature **ATE estimates** with q < 0.10 across all study × visit strata; %s ATE estimates total (after filtering the upstream RDS to `measure==\"ATE\"`). Counts and percentages are computed on ATE rows only — the RDS file also stores `MN` (mean) rows for the control and intervention arms, which are excluded here so the denominator matches per-arm intervention-effect estimates.*\n\n",
              format(nrow(df_sig), big.mark = ","), format(nrow(df), big.mark = ",")))
  print(htmltools::tagList(
    p,
    nice_dt(df_sig %>% dplyr::select(dplyr::any_of(c(
      "study","visit","contrast","label_f","biomarker","est","cil","ciu",
      "pval","pval_adj","sig","sigFDR"))), page_length = 25,
      caption = "Untargeted-metabolomics features with q < 0.10 (adjusted intervention effects, pooled BEP vs. control); the full feature set is in the linked RDS.")
  ))
}
```

## Pre-rendered volcano plot objects (manuscript figure panels)

For reference, we also include the volcano plot objects used to assemble the manuscript figures. These are saved ggplot objects from the upstream figure-prep pipeline; they render here as static images when plotly cannot convert their internal structure.

```{r explor-prerendered, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7}
v_combined <- safe_readRDS(here::here("figures/figure-data/volcano_plots_pooled_arms.RDS"),
                           label = "volcano_plots_pooled_arms.RDS")
show_any(v_combined)
v_strat <- safe_readRDS(here::here("figures/figure-data/volcano_plots.RDS"),
                        label = "volcano_plots.RDS")
show_any(v_strat)
```
