# Adjusted intervention effects on individual milk components

This section reports the primary covariate-adjusted estimates of maternal nutritional intervention effects on individual human milk components. Each component appears in an interactive forest plot and a searchable table, stratified by study and milk collection visit. We estimated average treatment effects (ATEs) with targeted maximum likelihood estimation (TMLE), adjusting for prespecified baseline prognostic factors, and applied Benjamini–Hochberg false discovery rate (FDR) correction within each outcome group × study × visit. Every results table also includes `pval_adj_global`, a more conservative correction pooled across all studies and visits within an outcome group, which we retain as a sensitivity analysis; the significance flag (`sigFDR`) is based on the per-visit `pval_adj`.

Each outcome group appears twice. The first view pools the intervention arms (all maternal nutritional intervention arms combined vs. the corresponding control) to maximize precision. The second stratifies by randomized arm, comparing each randomized nutritional arm with its trial's control. In the forest plots, filled triangles mark effects that survive FDR correction, filled circles ("Sig before FDR") mark nominal *P* < 0.05 that did not survive correction, and open circles mark neither; in the arm-stratified plots, each contrast also has its own color and CI line type. Hovering over a point shows the exact estimate, 95% CI, and *Q*-value, and the column filters in each table narrow the view to specific components, studies, or visits.

These estimates correspond to Figure 2 (primary outcomes) of the manuscript and Figure S3 (secondary outcomes) of the supplementary materials, and supply the per-component estimates behind the Figure 5A targeted-metabolite volcano plot. The machine-learning prediction of treatment arm and the PCA composites of manuscript Figure 1 are in §8.

> **Native assay units.** The estimates below are on the standardized (SD) scale so that components can be compared directly. The same intervention effects in native assay units (g/100 mL for macronutrients; µg/L or mg/L for micronutrients; nmol/mL for HMOs) are in §8.2 "Unscaled (natural-unit) intervention effects" of the *Sensitivity analyses and additional outputs* section, and each results table below includes an `ATE_unscaled` column with the natural-unit estimate next to the standardized one. Unadjusted estimates, for sensitivity, are in §8.1.

```{r effects-setup, include=FALSE}
library(tidyverse)
library(here)
library(DT)
library(plotly)
source(here::here("functions.R"))

## Render the forest plot and the searchable estimates table for one CSV.
## Returns an htmltools tagList so that htmlwidgets (plotly and DT) are picked
## up by knit_print at the chunk top level, since print() of widgets from inside
## a function does not trigger the htmlwidget rendering path.
show_forest_and_table <- function(csv_path, arm_strat = FALSE, title = NULL,
                                  fig_legend = NULL, tab_caption = NULL) {
  df <- safe_read_csv(here::here(csv_path))
  if (is.null(df) || !nrow(df)) return(htmltools::tagList())
  ttl <- title %||% "Forest plot of adjusted ATEs"
  legend <- fig_legend %||% paste0(
    "Each point is the adjusted average treatment effect (ATE, SD units) with 95% CI for one milk component, ",
    "faceted by study and milk-collection visit. ",
    if (arm_strat)
      "Each randomized contrast has its own color and CI line type (solid, dashed, dotted); filled triangles mark effects significant after FDR correction (Q < 0.05), filled circles nominal P < 0.05, and open circles non-significant. "
    else
      "Orange filled triangles mark effects significant after FDR correction (Q < 0.05); gray filled circles are nominally significant (P < 0.05); open circles are non-significant. ",
    "FDR correction is applied within outcome group, study, and visit. Hover over a point for the exact estimate, 95% CI, and Q-value.")
  htmltools::tagList(
    forest_plot(df, arm_strat = arm_strat, title = ttl),
    htmltools::tags$p(htmltools::tags$strong("Figure legend. "), htmltools::tags$em(legend)),
    clean_tab(df, caption = tab_caption %||% paste0("Adjusted ATEs, 95% CIs, and FDR-adjusted Q-values: ", ttl, "."))
  )
}
```

## Primary and secondary component metadata {#primary-component-metadata}

Primary outcomes are the macronutrients, micronutrients, and B-vitamins; secondary outcomes are the human milk oligosaccharides and bioactive proteins. The table below lists each component, grouped by prespecified outcome class, with indicators of whether it was measured in each of the three IMiC trials.

The manuscript's Fig. 3A legend cites this inventory as Section 5 of the online resource. The corresponding targeted-metabolite metadata, cited from Fig. 5A, are in [§6.1](#targeted-metabolite-metadata), and the complete inventory across all outcome classes, including tertiary, is in [§2 *Outcomes and human milk component distributions*](#outcomes-and-human-milk-component-distributions). The intervention-effect estimates for these components follow in the rest of this section.

```{r primary-component-metadata, echo=FALSE, message=FALSE, warning=FALSE}
metadata_loaded <- safe_load(here::here("metadata/milk_component.Rdata"))
if (isTRUE(metadata_loaded) && exists("all_milk_components")) {
  groups <- list(
    "Macronutrients (primary)"                = all_milk_components$macro,
    "Micronutrients (primary)"                = all_milk_components$micro,
    "B-vitamins (primary)"                    = all_milk_components$bvit,
    "Human milk oligosaccharides (secondary)" = all_milk_components$hmo,
    "Bioactive proteins (secondary)"          = all_milk_components$protein
  )
  groups <- groups[!vapply(groups, is.null, logical(1))]
  ps_tbl <- dplyr::bind_rows(lapply(names(groups), function(g) {
    tibble::tibble(outcome_group = g,
                   component     = unlist(groups[[g]], use.names = FALSE))
  })) %>% dplyr::distinct()
  per_study <- list(
    "MISAME-III" = if (exists("misame_milk_components")) unlist(misame_milk_components, use.names = FALSE) else NULL,
    "Mumta-LW"   = if (exists("vital_milk_components"))  unlist(vital_milk_components,  use.names = FALSE) else NULL,
    "ELICIT"     = if (exists("elicit_milk_components")) unlist(elicit_milk_components, use.names = FALSE) else NULL
  )
  for (s in names(per_study)) {
    if (is.null(per_study[[s]])) next
    ps_tbl[[paste0("measured_in_", s)]] <- ps_tbl$component %in% per_study[[s]]
  }
  nice_dt(ps_tbl,
          caption = "Primary and secondary human milk component metadata: outcome class and per-study availability. Referenced from Fig. 3A of the manuscript.")
} else {
  cat("*Component metadata is not available in this build (`metadata/milk_component.Rdata` did not load).*")
}
```

## Interactive effect explorer {#effect-explorer-app}

The app below collects every primary, secondary, and tertiary intervention effect. You can choose an outcome group, switch between standardized (SD) and native assay units, filter by study, visit, and pooled or arm-stratified contrasts, and move between volcano, forest, and table views. It runs in the browser on a WebAssembly build of R, with no server, so the first load takes a moment while the R runtime downloads. If the frame is cramped, [open the explorer full-screen in a new tab](apps/effect-explorer/).

```{r effect-explorer-embed, echo=FALSE, results='asis'}
if (knitr::is_html_output()) {
  cat('<iframe src="apps/effect-explorer/" title="IMiC Intervention-Effect Explorer" ',
      'width="100%" height="900" loading="lazy" ',
      'style="border:1px solid #ddd; border-radius:4px;"></iframe>\n', sep = "")
} else {
  cat("*The interactive effect explorer is available in the web version of this online resource.*\n")
}
```

## Macronutrients

> Total protein, fat, and digestible carbohydrate (primarily lactose), measured in g/100 mL, and total energy (kcal/L).

### Combined intervention arms

```{r effects-macro-combined, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
show_forest_and_table("results/subsetted results/primary_macro.csv",
                      title = "Macronutrients: pooled intervention vs. control")
```

### Stratified by randomized arm

```{r effects-macro-strat, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
show_forest_and_table("results/subsetted results/primary_macro_arm_strat.csv",
                      arm_strat = TRUE,
                      title = "Macronutrients: per-arm contrasts vs. control")
```

## Micronutrients (including fat-soluble vitamins)

### Combined intervention arms

```{r effects-micro-combined, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12}
show_forest_and_table("results/subsetted results/primary_micro.csv",
                      title = "Micronutrients: pooled intervention vs. control")
```

### Stratified by randomized arm

```{r effects-micro-strat, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12}
show_forest_and_table("results/subsetted results/primary_micro_arm_strat.csv",
                      arm_strat = TRUE,
                      title = "Micronutrients: per-arm contrasts vs. control")
```

## B-vitamins

### Combined intervention arms

```{r effects-bvit-combined, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12}
show_forest_and_table("results/subsetted results/primary_bvit.csv",
                      title = "B-vitamins: pooled intervention vs. control")
```

### Stratified by randomized arm

```{r effects-bvit-strat, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12}
show_forest_and_table("results/subsetted results/primary_bvit_arm_strat.csv",
                      arm_strat = TRUE,
                      title = "B-vitamins: per-arm contrasts vs. control")
```

## Human milk oligosaccharides (HMOs)

### Combined intervention arms

```{r effects-hmo-combined, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12}
show_forest_and_table("results/subsetted results/secondary_hmo.csv",
                      title = "HMOs: pooled intervention vs. control")
```

### Stratified by randomized arm

```{r effects-hmo-strat, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12}
show_forest_and_table("results/subsetted results/secondary_hmo_arm_strat.csv",
                      arm_strat = TRUE,
                      title = "HMOs: per-arm contrasts vs. control")
```

## Bioactive proteins

### Combined intervention arms

```{r effects-bio-combined, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12}
show_forest_and_table("results/subsetted results/secondary_bioactives.csv",
                      title = "Bioactive proteins: pooled intervention vs. control")
```

### Stratified by randomized arm

```{r effects-bio-strat, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12}
show_forest_and_table("results/subsetted results/secondary_bioactives_arm_strat.csv",
                      arm_strat = TRUE,
                      title = "Bioactive proteins: per-arm contrasts vs. control")
```

## Targeted metabolites (tertiary outcomes)

Summaries of the large targeted metabolomics panel (Biocrates MxP Quant 500) are elsewhere: pathway-level views in the *Pathway enrichment* section, and the corresponding volcano plot in manuscript Fig. 5A. The forest plots below show the same metabolites individually in one long plot (hover to identify points) and are most useful alongside a table filtered to a single compound class.

### Interactive explorer (linked volcano + table)

With thousands of metabolite estimates in the targeted panel, the linked view below is often the quickest place to start. Drawing a box or lasso (Plotly's *Box Select* / *Lasso*) around points in the volcano plot filters the table to those metabolites, and the dropdowns and the table's column filters in turn narrow the plotted points. Each point is one metabolite (pooled intervention vs. control), placed by effect size and −log10(*P*); FDR-significant metabolites are orange triangles, nominal (*P* < 0.05) ones blue squares, and the rest gray circles.

```{r effects-tertiary-linked, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7}
df_ter <- safe_read_csv(here::here("results/subsetted results/tertiary_targeted_metabolomics.csv"))
linked_volcano_table(df_ter,
                     title = "Targeted metabolites: pooled intervention vs. control (select points to filter the table)")
```

### Combined intervention arms

```{r effects-tertiary-combined, echo=FALSE, message=FALSE, warning=FALSE, fig.height=14}
show_forest_and_table("results/subsetted results/tertiary_targeted_metabolomics.csv",
                      title = "Targeted metabolites: pooled intervention vs. control")
```

### Stratified by randomized arm

```{r effects-tertiary-strat, echo=FALSE, message=FALSE, warning=FALSE, fig.height=14}
show_forest_and_table("results/subsetted results/tertiary_targeted_metabolomics_arm_strat.csv",
                      arm_strat = TRUE,
                      title = "Targeted metabolites: per-arm contrasts vs. control")
```
