# Adjusted intervention effects on individual milk components

Here we report the **primary, covariate-adjusted estimates** of how maternal nutritional interventions affect individual human milk components. Each component appears as an interactive forest plot alongside a searchable table, stratified by study and milk collection visit. We estimate average treatment effects (ATEs) using targeted maximum likelihood estimation (TMLE), adjusting for prespecified baseline prognostic factors, and apply the Benjamini–Hochberg false-discovery rate (FDR) correction within each outcome group × study × visit. Every result table also carries `pval_adj_global`, a more conservative correction pooled across all studies and visits within an outcome group (the pre-revision scheme), which we retain for sensitivity; significance flags (`sigFDR`) rest on the per-visit `pval_adj`.

Each modality appears twice. The first view **pools** the intervention arms (all maternal nutritional intervention arms combined vs. the corresponding control) to maximize precision. The second **stratifies by randomized arm**, contrasting each post-randomization nutritional arm against its trial's control. Within the forest plots, solid filled symbols mark effects that survive FDR correction and open symbols do not, while intermediate ("Sig before FDR") symbols flag nominal *p* < 0.05 that fell short of the multiplicity threshold. In the HTML version, hovering over a point reveals the exact estimate, 95% CI, and q-value, and the column filters in each table let you drill into specific components, studies, or visits.

These outputs map to **Figure 3** (primary-outcome forest plots) and **Figure S4** (secondary outcomes) in the manuscript, and they supply the per-component estimates behind the **Figure 5A** targeted-metabolite volcano. For the multivariate treatment-arm prediction and PCA composites of manuscript Figure 2, see §8.

> **Looking for absolute concentrations?** The estimates below sit on the **standardised (SD) scale** so that components can be compared directly. To read the same intervention effects in their **native assay units** (g/100 mL for macronutrients; µg/L or mg/L for micronutrients; relative abundance for HMOs), see **§8.2 "Unscaled (natural-unit) intervention effects"** in the *Sensitivity analyses and additional outputs* chapter. Each result table below carries an `ATE_unscaled` column giving the per-component natural-unit estimate next to the standardised one. Unadjusted estimates appear in **§8.1** for sensitivity.

```{r effects-setup, include=FALSE}
library(tidyverse)
library(here)
library(DT)
library(plotly)
source(here::here("functions.R"))

## Render the forest plot and the searchable estimates table for one CSV.
## Returns an htmltools tagList so that htmlwidgets (plotly and DT) are picked
## up by knit_print at the chunk top level — print() of widgets from inside
## a function does not trigger the htmlwidget rendering path.
show_forest_and_table <- function(csv_path, arm_strat = FALSE, title = NULL,
                                  fig_legend = NULL, tab_caption = NULL) {
  df <- safe_read_csv(here::here(csv_path))
  if (is.null(df) || !nrow(df)) return(htmltools::tagList())
  ttl <- title %||% "Forest plot of adjusted ATEs"
  legend <- fig_legend %||% paste0(
    "Each point is the adjusted average treatment effect (ATE, SD units) with 95% CI for one milk component, ",
    "faceted by study and milk-collection visit. ",
    if (arm_strat)
      "Symbols are coloured by randomised contrast; solid filled symbols mark effects significant after FDR correction (q<0.05), intermediate symbols nominal p<0.05, and open symbols non-significant. "
    else
      "Solid orange symbols mark effects significant after FDR correction (q<0.05); grey filled symbols are nominally significant (p<0.05); open symbols are non-significant. ",
    "FDR correction is applied within outcome group, study, and visit. Hover (HTML) for the exact estimate, 95% CI, and q-value.")
  htmltools::tagList(
    forest_plot(df, arm_strat = arm_strat, title = ttl),
    htmltools::tags$p(htmltools::tags$strong("Figure legend. "), htmltools::tags$em(legend)),
    clean_tab(df, caption = tab_caption %||% paste0("Adjusted ATEs, 95% CIs, and FDR-adjusted q-values — ", ttl, "."))
  )
}
```

## Interactive effect explorer {#effect-explorer-app}

The app below brings every primary, secondary, and tertiary intervention effect into one place. Choose an outcome group, switch between **standardised (SD) and native assay units**, filter by study, visit, and combined-vs-arm-stratified contrasts, and move between volcano, forest, and table views. It runs entirely in your browser through a WebAssembly build of R — no server — so the first load fetches the R runtime and takes a moment, after which it responds instantly. If the frame is cramped, [open the explorer full-screen in a new tab](apps/effect-explorer/).

```{r effect-explorer-embed, echo=FALSE, results='asis'}
if (knitr::is_html_output()) {
  cat('<iframe src="apps/effect-explorer/" title="IMiC Intervention-Effect Explorer" ',
      'width="100%" height="900" loading="lazy" ',
      'style="border:1px solid #ddd; border-radius:4px;"></iframe>\n', sep = "")
} else {
  cat("*The interactive effect explorer is available in the HTML build of this appendix.*\n")
}
```

## Macronutrients

> Total protein, fat, and digestible carbohydrate (primarily lactose), measured in g/100 mL.

### Combined intervention arms

```{r effects-macro-combined, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
show_forest_and_table("results/subsetted results/primary_macro.csv",
                      title = "Macronutrients — pooled intervention vs. control")
```

### Stratified by randomized arm

```{r effects-macro-strat, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
show_forest_and_table("results/subsetted results/primary_macro_arm_strat.csv",
                      arm_strat = TRUE,
                      title = "Macronutrients — per-arm contrasts vs. control")
```

## Micronutrients (including fat-soluble vitamins)

### Combined intervention arms

```{r effects-micro-combined, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12}
show_forest_and_table("results/subsetted results/primary_micro.csv",
                      title = "Micronutrients — pooled intervention vs. control")
```

### Stratified by randomized arm

```{r effects-micro-strat, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12}
show_forest_and_table("results/subsetted results/primary_micro_arm_strat.csv",
                      arm_strat = TRUE,
                      title = "Micronutrients — per-arm contrasts vs. control")
```

## B-vitamins

### Combined intervention arms

```{r effects-bvit-combined, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12}
show_forest_and_table("results/subsetted results/primary_bvit.csv",
                      title = "B-vitamins — pooled intervention vs. control")
```

### Stratified by randomized arm

```{r effects-bvit-strat, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12}
show_forest_and_table("results/subsetted results/primary_bvit_arm_strat.csv",
                      arm_strat = TRUE,
                      title = "B-vitamins — per-arm contrasts vs. control")
```

## Human milk oligosaccharides (HMOs)

### Combined intervention arms

```{r effects-hmo-combined, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12}
show_forest_and_table("results/subsetted results/secondary_hmo.csv",
                      title = "HMOs — pooled intervention vs. control")
```

### Stratified by randomized arm

```{r effects-hmo-strat, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12}
show_forest_and_table("results/subsetted results/secondary_hmo_arm_strat.csv",
                      arm_strat = TRUE,
                      title = "HMOs — per-arm contrasts vs. control")
```

## Bioactive proteins

### Combined intervention arms

```{r effects-bio-combined, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12}
show_forest_and_table("results/subsetted results/secondary_bioactives.csv",
                      title = "Bioactive proteins — pooled intervention vs. control")
```

### Stratified by randomized arm

```{r effects-bio-strat, echo=FALSE, message=FALSE, warning=FALSE, fig.height=12}
show_forest_and_table("results/subsetted results/secondary_bioactives_arm_strat.csv",
                      arm_strat = TRUE,
                      title = "Bioactive proteins — per-arm contrasts vs. control")
```

## Targeted metabolites (tertiary outcomes)

The targeted-metabolomics panel (Biocrates MxP Quant 500) is large, so higher-level summaries live elsewhere: pathway-level views in the *Pathway enrichment* chapter and volcano plots in the *Trajectory analyses* chapter. The forest plots below lay out the same metabolites as a long, hover-searchable forest, most useful when you filter the table to a specific compound class.

### Interactive explorer (linked volcano + table)

Because the targeted panel has thousands of metabolite estimates, the linked view below is often the quickest way in. Drag a box (Plotly's *Box Select* / *Lasso*) around points in the volcano to filter the table to just those metabolites; conversely, the dropdowns and the table's own column filters narrow the plotted points. Every point is one metabolite (pooled intervention vs. control), positioned by effect size and −log10(p); orange marks FDR-significant hits.

```{r effects-tertiary-linked, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7}
df_ter <- safe_read_csv(here::here("results/subsetted results/tertiary_targeted_metabolomics.csv"))
linked_volcano_table(df_ter,
                     title = "Targeted metabolites — pooled intervention vs. control (brush to filter the table)")
```

### Combined intervention arms

```{r effects-tertiary-combined, echo=FALSE, message=FALSE, warning=FALSE, fig.height=14}
show_forest_and_table("results/subsetted results/tertiary_targeted_metabolomics.csv",
                      title = "Targeted metabolites — pooled intervention vs. control")
```

### Stratified by randomized arm

```{r effects-tertiary-strat, echo=FALSE, message=FALSE, warning=FALSE, fig.height=14}
show_forest_and_table("results/subsetted results/tertiary_targeted_metabolomics_arm_strat.csv",
                      arm_strat = TRUE,
                      title = "Targeted metabolites — per-arm contrasts vs. control")
```
