# Adjusted intervention effects on individual milk components

Covariate-adjusted effects on each primary, secondary, and tertiary milk component, by trial and visit, for the pooled intervention arms and for each randomized arm vs. control. In the tables, `ATE` is the effect in SD units and `ATE_unscaled` in native units, both with 95% CIs; `perc_imp` is the difference between the intervention and control means as a percentage of the control mean; `pval_adj` is FDR-adjusted within outcome group, trial, and visit; `pval_adj_global` is adjusted across all trials and visits of an outcome group; and asterisks mark *P* or *Q* below 0.05, 0.01, and 0.001.

```{r effects-setup, include=FALSE}
library(tidyverse)
library(here)
library(DT)
library(plotly)
source(here::here("functions.R"))

## Searchable table of the estimates in one results CSV.
effects_table <- function(csv_path, caption) {
  df <- safe_read_csv(here::here(csv_path))
  if (is.null(df) || !nrow(df)) return(htmltools::tagList())
  clean_tab(df, caption = caption)
}
metadata_loaded <- safe_load(here::here("metadata/milk_component.Rdata"))
```

## Primary and secondary component metadata {#primary-component-metadata}

Primary and secondary components by outcome class, with the trials that measured each.

```{r primary-component-metadata, echo=FALSE, message=FALSE, warning=FALSE}
if (isTRUE(metadata_loaded) && exists("all_milk_components")) {
  groups <- list(
    "Macronutrients (primary)"                = all_milk_components$macro,
    "Micronutrients (primary)"                = all_milk_components$micro,
    "B-vitamins (primary)"                    = all_milk_components$bvit,
    "Human milk oligosaccharides (secondary)" = all_milk_components$hmo,
    "Bioactive proteins (secondary)"          = all_milk_components$protein
  )
  groups <- groups[!vapply(groups, is.null, logical(1))]
  ps_tbl <- dplyr::bind_rows(lapply(names(groups), function(g) {
    tibble::tibble(outcome_group = g,
                   component     = unlist(groups[[g]], use.names = FALSE))
  })) %>% dplyr::distinct()
  per_study <- list(
    "MISAME-III" = if (exists("misame_milk_components")) unlist(misame_milk_components, use.names = FALSE) else NULL,
    "Mumta-LW"   = if (exists("vital_milk_components"))  unlist(vital_milk_components,  use.names = FALSE) else NULL,
    "ELICIT"     = if (exists("elicit_milk_components")) unlist(elicit_milk_components, use.names = FALSE) else NULL
  )
  for (s in names(per_study)) {
    if (is.null(per_study[[s]])) next
    ps_tbl[[paste0("measured_in_", s)]] <- ps_tbl$component %in% per_study[[s]]
  }
  nice_dt(ps_tbl, caption = "Primary and secondary components.")
} else {
  cat("*Component metadata is not available in this build (`metadata/milk_component.Rdata` did not load).*")
}
```

## Targeted-metabolite metadata {#targeted-metabolite-metadata}

The targeted metabolites (Biocrates MxP Quant 500 panel) by compound class, with the trials that measured each.

```{r targeted-metabolite-metadata, echo=FALSE, message=FALSE, warning=FALSE}
if (isTRUE(metadata_loaded) && exists("all_milk_components") &&
    !is.null(all_milk_components$metabolomics)) {
  ## all_milk_components$metabolomics may be a single character vector or a
  ## named list keyed by Biocrates category. Handle both.
  metab_obj <- all_milk_components$metabolomics
  if (is.list(metab_obj) && !is.null(names(metab_obj))) {
    metab_tbl <- dplyr::bind_rows(lapply(names(metab_obj), function(cat) {
      tibble::tibble(
        biocrates_category = cat,
        metabolite         = unlist(metab_obj[[cat]], use.names = FALSE)
      )
    }))
  } else {
    metab_tbl <- tibble::tibble(
      biocrates_category = "Targeted metabolites",
      metabolite         = unlist(metab_obj, use.names = FALSE)
    )
  }
  metab_tbl <- dplyr::distinct(metab_tbl)
  pick_metab <- function(x) if (is.list(x)) unlist(x$metabolomics %||% x, use.names = FALSE)
                            else unlist(x, use.names = FALSE)
  per_study <- list(
    "MISAME-III" = if (exists("misame_milk_components")) pick_metab(misame_milk_components) else NULL,
    "Mumta-LW"   = if (exists("vital_milk_components"))  pick_metab(vital_milk_components)  else NULL,
    "ELICIT"     = if (exists("elicit_milk_components")) pick_metab(elicit_milk_components) else NULL
  )
  for (s in names(per_study)) {
    if (is.null(per_study[[s]])) next
    metab_tbl[[paste0("measured_in_", s)]] <- metab_tbl$metabolite %in% per_study[[s]]
  }
  nice_dt(metab_tbl, caption = "Targeted metabolites.")
} else {
  cat("*Targeted-metabolite metadata is not available in this build (`metadata/milk_component.Rdata` did not load).*")
}
```

## Interactive effect explorer {#effect-explorer-app}

Every primary, secondary, and tertiary effect as a volcano plot, forest plot, or table, in SD or native units, filtered by trial, visit, and contrast. The app runs R in the browser, so the first load takes a moment. [Open it in a new tab](apps/effect-explorer/).

```{r effect-explorer-embed, echo=FALSE, results='asis'}
if (knitr::is_html_output()) {
  cat('<iframe src="apps/effect-explorer/" title="IMiC Intervention-Effect Explorer" ',
      'width="100%" height="900" loading="lazy" ',
      'style="border:1px solid #ddd; border-radius:4px;"></iframe>\n', sep = "")
} else {
  cat("*The interactive effect explorer is available in the web version of this online resource.*\n")
}
```

## Macronutrients

```{r effects-macro, echo=FALSE, message=FALSE, warning=FALSE}
htmltools::tagList(
  effects_table("results/subsetted results/primary_macro.csv", "Pooled intervention vs. control."),
  effects_table("results/subsetted results/primary_macro_arm_strat.csv", "Each randomized arm vs. control."))
```

## Micronutrients (including fat-soluble vitamins)

```{r effects-micro, echo=FALSE, message=FALSE, warning=FALSE}
htmltools::tagList(
  effects_table("results/subsetted results/primary_micro.csv", "Pooled intervention vs. control."),
  effects_table("results/subsetted results/primary_micro_arm_strat.csv", "Each randomized arm vs. control."))
```

## B-vitamins

```{r effects-bvit, echo=FALSE, message=FALSE, warning=FALSE}
htmltools::tagList(
  effects_table("results/subsetted results/primary_bvit.csv", "Pooled intervention vs. control."),
  effects_table("results/subsetted results/primary_bvit_arm_strat.csv", "Each randomized arm vs. control."))
```

## Human milk oligosaccharides (HMOs)

```{r effects-hmo, echo=FALSE, message=FALSE, warning=FALSE}
htmltools::tagList(
  effects_table("results/subsetted results/secondary_hmo.csv", "Pooled intervention vs. control."),
  effects_table("results/subsetted results/secondary_hmo_arm_strat.csv", "Each randomized arm vs. control."))
```

## Bioactive proteins

```{r effects-bio, echo=FALSE, message=FALSE, warning=FALSE}
htmltools::tagList(
  effects_table("results/subsetted results/secondary_bioactives.csv", "Pooled intervention vs. control."),
  effects_table("results/subsetted results/secondary_bioactives_arm_strat.csv", "Each randomized arm vs. control."))
```

## Targeted metabolites (tertiary outcomes)

Selecting points in the volcano plot (box or lasso select) filters the table beneath it, and the menus filter both. Each point is one metabolite, pooled intervention vs. control; FDR is controlled within each trial and visit, so this pooled view has no FDR line.

```{r effects-tertiary-linked, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7}
df_ter <- safe_read_csv(here::here("results/subsetted results/tertiary_targeted_metabolomics.csv"))
linked_volcano_table(df_ter,
                     title = "Targeted metabolites")
```

```{r effects-tertiary, echo=FALSE, message=FALSE, warning=FALSE}
htmltools::tagList(
  effects_table("results/subsetted results/tertiary_targeted_metabolomics.csv", "Pooled intervention vs. control."),
  effects_table("results/subsetted results/tertiary_targeted_metabolomics_arm_strat.csv", "Each randomized arm vs. control."))
```
