# Trajectory analyses of human milk component change across lactation

Effects on the change in each milk component between consecutive visits (the later value minus the earlier one, within each mother), standardized and estimated by TMLE, by trial and visit interval.

```{r traj-setup, include=FALSE}
library(tidyverse)
library(here)
library(DT)
library(plotly)
source(here::here("functions.R"))
## Larger base font for this chapter's plots.
ggplot2::theme_set(theme_imic_web(base_size = 14))
knitr::opts_chunk$set(fig.width = 12, fig.height = 9, dpi = 110)
```

## Volcano plots of trajectory effects

```{r traj-volcanoes, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7, results='asis'}
traj_primary   <- safe_readRDS(here::here("results/adjusted_intervention_effects_traj_results_clean.RDS"),
                               label = "adjusted_intervention_effects_traj_results_clean.RDS")
traj_combined  <- safe_readRDS(here::here("results/adjusted_combined_arms_intervention_effects_traj_results_clean.RDS"),
                               label = "adjusted_combined_arms_intervention_effects_traj_results_clean.RDS")

## Restrict to ATE rows (the trajectory RDS files contain both ATE estimates and
## MN/mean-estimate rows; only ATE rows are intervention-effect estimates, so
## counts, significance, volcano points, and table rows must use ATE only).
if (is.data.frame(traj_primary)  && "measure" %in% names(traj_primary))
  traj_primary  <- traj_primary  %>% dplyr::filter(measure == "ATE")
if (is.data.frame(traj_combined) && "measure" %in% names(traj_combined))
  traj_combined <- traj_combined %>% dplyr::filter(measure == "ATE")

## Each estimate is the effect on the change from the preceding visit, so panels
## are titled by visit interval ("MISAME-III (14-21 days to 1-2 mo.)").
traj_interval <- c("Misame 1-2 mo." = "14-21 days to 1-2 mo.", "Misame 3-4 mo." = "1-2 to 3-4 mo.",
                   "Vital 2 mo." = "1.5 to 2 mo.", "Elicit 5 mo." = "1 to 5 mo.")
make_traj_volcano <- function(df, cat_filter, title) {
  if (is.null(df) || !is.data.frame(df)) return(invisible(NULL))
  key <- paste(df$study, df$visit)
  df$visit[key %in% names(traj_interval)] <- traj_interval[key[key %in% names(traj_interval)]]
  cond <- rep(FALSE, nrow(df))
  if ("category" %in% names(df))      cond <- cond | df$category %in% cat_filter
  if ("outcome_group" %in% names(df)) cond <- cond | df$outcome_group %in% cat_filter
  if ("outcome_class" %in% names(df)) cond <- cond | df$outcome_class %in% cat_filter
  sub <- df[cond, , drop = FALSE]
  if (!nrow(sub)) return(invisible(NULL))
  ## Cap interactive points for tertiary metabolites (thousands of features).
  volcano_capped(sub, n_show = 1500, title = title, ncol = 2,
                 x_lab = "Average treatment effect on the visit-to-visit change (SD units)")
}

make_traj_volcano(traj_combined,
                  cat_filter = c("primary","Primary"),
                  title = "Trajectory effects: primary outcomes")
make_traj_volcano(traj_combined,
                  cat_filter = c("secondary","Secondary"),
                  title = "Trajectory effects: secondary outcomes")
make_traj_volcano(traj_combined,
                  cat_filter = c("tertiary","Tertiary"),
                  title = "Trajectory effects: targeted metabolites")
```

## Searchable tables of trajectory effects

### Per-arm contrasts

```{r traj-table-armstrat, echo=FALSE, message=FALSE, warning=FALSE}
if (!is.null(traj_primary) && is.data.frame(traj_primary)) {
  nice_dt(traj_primary %>% dplyr::mutate(dplyr::across(where(is.numeric), ~ round(.x, 3))),
          caption = "Adjusted trajectory ATEs, per randomized arm vs. control.")
}
```

### Combined intervention arms

```{r traj-table-combined, echo=FALSE, message=FALSE, warning=FALSE}
if (!is.null(traj_combined) && is.data.frame(traj_combined)) {
  nice_dt(traj_combined %>% dplyr::mutate(dplyr::across(where(is.numeric), ~ round(.x, 3))),
          caption = "Adjusted trajectory ATEs, pooled intervention arms vs. control.")
}
```
