# Trajectory analyses of human milk component change across lactation

Did maternal nutritional interventions alter **within-individual changes** in human milk composition across successive postpartum visits? The trajectory analyses in this chapter address that question. Where the cross-sectional analyses elsewhere in the supplement compare arms at a given visit, these track visit-to-visit change within the same mother–infant dyad.

For each milk component, we transformed the raw concentration measurements into **visit-to-visit differences** (current visit minus the immediately preceding visit within each mother). Baseline visits without a prior observation were dropped. We estimated trajectory-specific intervention effects with TMLE, adjusting for the same prespecified baseline prognostic covariates used in the primary adjusted analyses, and standardized outcomes before estimation. Analyses ran separately by study and visit interval, and arms were pooled where appropriate to improve precision.

Results appear as **volcano plots** for primary, secondary, and tertiary outcomes, and as **trajectory line plots** showing mean milk component concentrations across visits by intervention arm. In the HTML build the volcano plots render as interactive plotly figures; hover any point to read the component name, study, visit interval, ATE, and FDR-adjusted q-value.

> **Interpretation note.** Trajectory analyses are exploratory. The number of repeated measurements per participant and the per-stratum sample size both limit power to detect trajectory-specific effects. Primary inference about intervention effects rests on the cross-sectional adjusted analyses. The trajectory results add context, letting readers ask whether supplementation produces changes that are consistent, delayed, or cumulative across early lactation.

```{r traj-setup, include=FALSE}
library(tidyverse)
library(here)
library(DT)
library(plotly)
source(here::here("functions.R"))
## Reviewer flagged static plot labels as too small. Apply a larger base font
## for newly built plots and bump default fig dimensions for the trajectory
## line plots (saved upstream as smaller-font ggplots).
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 14))
knitr::opts_chunk$set(fig.width = 12, fig.height = 9, dpi = 110)
```

## Volcano plots of trajectory effects

```{r traj-volcanoes, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7, results='asis'}
traj_primary   <- safe_readRDS(here::here("results/adjusted_intervention_effects_traj_results_clean.RDS"),
                               label = "adjusted_intervention_effects_traj_results_clean.RDS")
traj_combined  <- safe_readRDS(here::here("results/adjusted_combined_arms_intervention_effects_traj_results_clean.RDS"),
                               label = "adjusted_combined_arms_intervention_effects_traj_results_clean.RDS")

## Restrict to ATE rows (the trajectory RDS files contain both ATE estimates and
## MN/mean-estimate rows; only ATE rows are intervention-effect estimates, so
## counts, significance, volcano points, and table rows must use ATE only).
if (is.data.frame(traj_primary)  && "measure" %in% names(traj_primary))
  traj_primary  <- traj_primary  %>% dplyr::filter(measure == "ATE")
if (is.data.frame(traj_combined) && "measure" %in% names(traj_combined))
  traj_combined <- traj_combined %>% dplyr::filter(measure == "ATE")

make_traj_volcano <- function(df, cat_filter, title) {
  if (is.null(df) || !is.data.frame(df)) return(invisible(NULL))
  cond <- rep(FALSE, nrow(df))
  if ("category" %in% names(df))      cond <- cond | df$category %in% cat_filter
  if ("outcome_group" %in% names(df)) cond <- cond | df$outcome_group %in% cat_filter
  if ("outcome_class" %in% names(df)) cond <- cond | df$outcome_class %in% cat_filter
  sub <- df[cond, , drop = FALSE]
  if (!nrow(sub)) return(invisible(NULL))
  ## Cap interactive points for tertiary metabolites (thousands of features).
  volcano_capped(sub, n_show = 1500, title = title)
}

make_traj_volcano(traj_combined,
                  cat_filter = c("primary","Primary"),
                  title = "Trajectory effects — primary outcomes")
make_traj_volcano(traj_combined,
                  cat_filter = c("secondary","Secondary"),
                  title = "Trajectory effects — secondary outcomes")
make_traj_volcano(traj_combined,
                  cat_filter = c("tertiary","Tertiary"),
                  title = "Trajectory effects — tertiary outcomes (targeted metabolites)")
```

**Trajectory volcano plots.** Each point is the adjusted intervention effect on the visit-to-visit *change* in one milk component (primary, secondary, and tertiary outcomes shown separately), pooled intervention vs. control. The x-axis is the ATE (SD units) and the y-axis is −log~10~(*p*), with the dotted line marking *p* = 0.05. Points are coloured by significance tier. In the high-cardinality tertiary panel, the interactive plot shows the top features by raw *p*-value, and the full set is in the searchable table below. Trajectory analyses are exploratory.

## Searchable tables of trajectory effects

### Per-arm contrasts

```{r traj-table-armstrat, echo=FALSE, message=FALSE, warning=FALSE}
if (!is.null(traj_primary) && is.data.frame(traj_primary)) {
  nice_dt(traj_primary %>% dplyr::mutate(dplyr::across(where(is.numeric), ~ round(.x, 3))),
          caption = "Adjusted trajectory ATEs, per randomized arm vs. control.")
}
```

### Combined intervention arms

```{r traj-table-combined, echo=FALSE, message=FALSE, warning=FALSE}
if (!is.null(traj_combined) && is.data.frame(traj_combined)) {
  nice_dt(traj_combined %>% dplyr::mutate(dplyr::across(where(is.numeric), ~ round(.x, 3))),
          caption = "Adjusted trajectory ATEs, pooled intervention arms vs. control.")
}
```

## Trajectory line plots by intervention arm

The figures below trace human milk components over successive visits by intervention arm within each study. Line plots display mean milk component levels across visits, aggregated within biologically defined component categories (e.g. macronutrients, B-vitamins, HMOs, amino acids, other metabolite classes) and stratified by randomized intervention arm. To build the category-level summaries, we averaged component values across all available biomarkers within a category in each study. Plots appear on two scales: a standardized scale (z-scored within study × visit) for cross-category comparison, and the original measurement scale, which keeps absolute concentrations interpretable.

```{r traj-lineplots, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
p_s5 <- safe_readRDS(here::here("figure-data/figure_s4_trajectory_plots.RDS"),
                     label = "figure_s4_trajectory_plots.RDS")
show_any(p_s5)
```

**Trajectory line plots.** Mean human-milk component levels across successive milk-collection visits by randomized intervention arm within each study, averaged within biologically defined component categories (e.g. macronutrients, B-vitamins, HMOs, amino acids, other metabolite classes). Shown on a standardized (z-scored within study × visit) scale for cross-category comparison and on the original measurement scale, which preserves absolute concentrations.
