# Sensitivity analyses and additional outputs

The robustness analyses and supplementary outputs referenced from the manuscript are gathered here, grouped by analysis class. Each subsection asks a version of the same question of the primary adjusted, standardised TMLE analyses presented earlier: would the headline conclusions still hold?

- **8.1 Unadjusted intervention effects** — the same TMLE framework with covariate adjustment removed, which shows whether adjustment is doing meaningful work in the primary analyses.
- **8.2 Unscaled (natural-unit) intervention effects** — adjusted TMLE ATEs in the native assay units (e.g. g/100 mL, µg/L), letting readers interpret effect magnitudes clinically.
- **8.3 Fat-adjusted triglyceride effects (Figure S5)** — separating changes in triglyceride composition from transient changes in total milk fat.
- **8.4 SuperLearner CV-AUC treatment-arm prediction** (manuscript **Fig. 2A**) — a multivariate test of whether the *joint* distribution of HM components carries intervention signal beyond per-component testing.
- **8.5 PCA-based composite outcome scores** (manuscript **Fig. 2B**) — ATEs on the first principal component of each modality, asking whether each modality moved as a whole.
- **8.6 Pathway enrichment analyses** — now reported in full in [§9 *Reproducible pathway and enrichment analyses*](#reproducible-pathway-and-enrichment-analyses); only the network views and a pointer remain here.
- **8.7 Infant growth analyses (Figure S1)** — intervention effects on infant anthropometry within the IMiC subsamples, as a within-trial sanity check.

Primary inference about whether intervention effects are present and consistent rests on the adjusted analyses in §5. The analyses in this chapter contextualise those findings rather than replace them.

```{r sens-setup, include=FALSE}
library(tidyverse)
library(here)
library(DT)
library(plotly)
source(here::here("functions.R"))

## Apply a slightly larger base font to all static ggplots in this chapter
## (reviewer flagged static plot labels as too small to read).
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 14))

metadata_loaded <- safe_load(here::here("metadata/milk_component.Rdata"))
```

## Unadjusted intervention effects

Unadjusted estimates use the same TMLE framework but omit covariate adjustment. Where adjusted and unadjusted estimates differ, the gap may reflect confounding by baseline maternal, infant, or household characteristics, differences in sampling structure across studies and visits, or chance. Within a single randomised trial a large divergence would be surprising and would warrant a closer look at the adjustment set; these forest plots are best read alongside those in §5.

```{r sens-unadjusted, echo=FALSE, message=FALSE, warning=FALSE, fig.height=10}
res_unadj <- safe_readRDS(here::here("results/unadjusted_intervention_effects_results_clean.RDS"),
                          label = "unadjusted_intervention_effects_results_clean.RDS")
if (!is.null(res_unadj) && isTRUE(metadata_loaded)) {
  res_unadj <- res_unadj %>% dplyr::filter(measure == "ATE") %>%
    dplyr::mutate(biomarker = stringr::str_to_lower(biomarker))
  groups <- list(
    "Macronutrients"      = c(all_milk_components$macro, "total carbohydrate"),
    "Micronutrients"      = all_milk_components$micro,
    "B-vitamins"          = all_milk_components$bvit,
    "HMOs"                = all_milk_components$hmo,
    "Bioactive proteins"  = all_milk_components$protein
  )
  out <- htmltools::tagList()
  for (g in names(groups)) {
    sub <- res_unadj %>% dplyr::filter(biomarker %in% groups[[g]])
    if (!nrow(sub)) next
    out <- htmltools::tagList(
      out,
      htmltools::h3(g),
      forest_plot(sub, title = paste0("Unadjusted ATEs — ", g)),
      clean_tab(sub)
    )
  }
  out
}
```

## Unscaled (natural-unit) intervention effects

Unscaled estimates are adjusted TMLE ATEs reported on the **native assay scale** of each milk component (for example, g/100 mL for macronutrients; µg/L or mg/L for micronutrients; relative abundance for HMOs). The standardised, SD-unit estimates in §5 are built for cross-component comparison. The natural-unit estimates here restate them in biologically meaningful units, so a reader can ask "how much did B3 actually go up, in mg/L?" rather than only "by how many SD units did B3 go up?". Think of this section as the **absolute-concentration view** of the §5 results. These native-unit estimates correspond to **Table S7** of the manuscript (per-arm sample sizes, means, SDs, medians and IQRs, alongside native-unit ATEs, 95% CIs, and FDR-adjusted *q*-values for all primary and secondary outcomes).

```{r sens-unscaled, echo=FALSE, message=FALSE, warning=FALSE, fig.height=10}
res_unscaled <- safe_readRDS(here::here("results/adjusted_combined_arms_intervention_effects_unscaled_results_clean.RDS"),
                             label = "adjusted_combined_arms_intervention_effects_unscaled_results_clean.RDS")
if (!is.null(res_unscaled) && isTRUE(metadata_loaded)) {
  res_unscaled <- res_unscaled %>% dplyr::filter(measure == "ATE") %>%
    dplyr::mutate(biomarker = stringr::str_to_lower(biomarker))
  groups <- list(
    "Macronutrients (native units)"     = c(all_milk_components$macro, "total carbohydrate"),
    "Micronutrients (native units)"     = all_milk_components$micro,
    "B-vitamins (native units)"         = all_milk_components$bvit,
    "HMOs (native units)"               = all_milk_components$hmo,
    "Bioactive proteins (native units)" = all_milk_components$protein
  )
  out <- htmltools::tagList()
  for (g in names(groups)) {
    sub <- res_unscaled %>% dplyr::filter(biomarker %in% groups[[g]])
    if (!nrow(sub)) next
    out <- htmltools::tagList(
      out,
      htmltools::h3(g),
      forest_plot(sub, title = g),
      clean_tab(sub)
    )
  }
  out
}
```

### Table S7 — per-arm native-unit concentrations and descriptive statistics

The full manuscript **Table S7** is reproduced below as a searchable table. For every primary and secondary milk component, by study × visit × arm, it gives the sample size, the covariate-adjusted mean and 95% CI (`mean`, `mean_cil`, `mean_ciu`), and the raw distributional summary (`raw_mean`, `raw_sd`, `raw_median`, `raw_q25`, `raw_q75`) in native assay units. It provides the tabular backing for the concentration **distributions** in manuscript Fig. 4 and the effect estimates in Figs. 2–3.

```{r sens-table-s7, echo=FALSE, message=FALSE, warning=FALSE}
s7 <- safe_read_csv(here::here("results/tables/table_s7_primary_secondary_native_units.csv"))
if (!is.null(s7)) {
  nice_dt(s7,
          caption = "Manuscript Table S7 — native-unit (absolute) concentrations and adjusted means for all primary and secondary milk outcomes, by study, visit, and intervention arm.")
}
```

## Fat-adjusted triglyceride effects (Figure S5)

The largest tertiary-outcome signal in the manuscript is on **triglycerides (TGs)**. In MISAME-III the same TGs were *increased* at 14–21 days and 3–4 months but *decreased* at the 1–2 month visit. This temporal inversion disappears once concentrations are expressed as a fraction of total fat content, which implies that the apparent 1–2 month decrease reflects the transient drop in total fat at that visit rather than a compositional change in the TG pool. The plot below reproduces Figure S5 (treatment-specific TG means with and without fat adjustment), and the forest plot beneath it shows the fat-adjusted ATEs behind that interpretation.

> **Milk-volume caveat.** Total milk *output* was not measured in any of the three IMiC trials. The transient reduction in milk fat concentration seen at 1–2 months in MISAME-III may therefore reflect dilution from increased milk volume in the supplemented arm rather than a true reduction in fat synthesis or secretion. Concentration-based effects on macronutrients, and on any fat-dependent metabolite, should be read with this in mind: without a denominator on total daily output, a change in concentration cannot be translated cleanly into a change in infant nutrient delivery. The fat-adjusted analysis below removes one form of this confounding, proportional re-allocation across TGs, but it cannot recover the missing volume term.

```{r sens-fat-tg-figure, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
p_s5 <- safe_readRDS(here::here("figure-data/figure_s7_tri_pattern.RDS"),
                     label = "figure_s7_tri_pattern.RDS")
show_any(p_s5)
```

```{r sens-fat-tg-forest, echo=FALSE, message=FALSE, warning=FALSE, fig.height=10}
p_s6 <- safe_readRDS(here::here("figure-data/figure_sX_forest_plot_fat_adjusted.RDS"),
                     label = "figure_sX_forest_plot_fat_adjusted.RDS")
show_any(p_s6)
```

**Figure S5.** Treatment-specific means (and 95% CIs) of milk triglycerides (N = 235 TG species from the targeted-metabolomics panel) in MISAME-III, with and without adjustment for total milk fat. *Top:* treatment-specific mean TG-species concentrations; points are coloured by significance of the ATE estimate, and TGs are sorted by the mean Z-scored control-arm concentration at the first visit. The reductions seen at the second timepoint in the postnatal BEP arms (panels O and U) dissipate when concentrations are expressed as a percentage of total fat (panels P and V), revealing the same TG-species pattern as at the first and third timepoints. *Bottom:* fat-adjusted adjusted ATEs (SD units, 95% CI) for individual TG species, confirming a consistent direction of effect across visits once total milk fat is accounted for.

## SuperLearner multi-class treatment-arm prediction

The Methods describe a **SuperLearner ensemble** trained to predict multi-class randomised treatment-arm membership from the full set of measured primary, secondary, and tertiary HM outcomes within each study × visit. A cross-validated AUC (CV-AUC) above chance indicates that the joint distribution of HM components contains intervention signal beyond what any single component carries, and variable-importance (VIMP) measures identify which components carry most of that predictive signal.

This is the multivariate counterpart to the univariate forest plots in §5. It asks whether the per-component signals, taken together, predict arm assignment, and which components drive the prediction.

```{r sens-sl-vimp, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
sl <- safe_readRDS(here::here("figure-data/SL_vim_plot_data.RDS"),
                   label = "SL_vim_plot_data.RDS")
if (!is.null(sl) && is.data.frame(sl)) {
  sl <- sl %>% dplyr::mutate(
    tooltip = paste0("Study: ", studyid,
                     "<br>Visit: ", visit,
                     "<br>Arm: ", arm,
                     "<br>Group: ", group,
                     "<br>N: ", N_obs,
                     "<br>CV-AUC: ", round(cvAUC, 3),
                     " (", round(ci.lb, 3), ", ", round(ci.ub, 3), ")")
  )
  p_sl <- ggplot2::ggplot(sl,
                          ggplot2::aes(x = cvAUC,
                                       y = stats::reorder(paste(studyid, visit_f, arm_f, sep=" | "), cvAUC),
                                       color = group, text = tooltip)) +
    ggplot2::geom_point(size = 2) +
    ggplot2::geom_errorbar(ggplot2::aes(xmin = ci.lb, xmax = ci.ub),
                           width = 0.2, orientation = "y") +
    ggplot2::geom_vline(xintercept = 0.5, linetype = "dashed", color = "grey50") +
    ggplot2::scale_color_manual(values = tableau10) +
    ggplot2::labs(x = "Cross-validated AUC",
                  y = NULL,
                  color = "Outcome group",
                  title = "SuperLearner multi-class treatment-arm prediction (CV-AUC)")
  out_sl <- if (knitr::is_html_output() && requireNamespace("plotly", quietly = TRUE)) {
    conv <- try(plotly::ggplotly(p_sl, tooltip = "text"), silent = TRUE)
    if (!inherits(conv, "try-error")) conv else p_sl
  } else p_sl
  htmltools::tagList(out_sl,
    nice_dt(sl %>% dplyr::select(studyid, visit, arm, group, N_obs, cvAUC, ci.lb, ci.ub) %>%
              dplyr::mutate(dplyr::across(where(is.numeric), ~ round(.x, 3))),
            caption = "SuperLearner CV-AUC for multi-class treatment-arm prediction."))
}
```

**Figure 2A.** Cross-validated AUC (CV-AUC) from SuperLearner ensembles trained to predict randomized treatment arm from each class of milk components (and from all classes combined), by study and milk-collection visit. CV-AUC ≈ 0.5 (dashed line) is chance-level; higher values indicate that the joint distribution of that component class carries intervention signal. B-vitamins and targeted metabolites give the strongest signals, while azithromycin-only and prenatal-BEP arms show none.

### Per-study variable importance

The pre-built per-study variable-importance plots, one per IMiC trial, are reproduced below for completeness.

```{r sens-sl-vimp-perstudy, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
sl_plots <- safe_readRDS(here::here("results/SLvim_lab_plots.RDS"),
                         label = "SLvim_lab_plots.RDS")
show_any(sl_plots)
```

## PCA-based composite outcome scores

ATEs were also estimated for the **first principal component** of each modality (macronutrients, B-vitamins, HMOs, bioactives, targeted metabolites). A PCA-based composite collapses correlated within-modality components into a single scalar capturing the dominant axis of variation, which makes it sensitive to coordinated multivariate shifts that may not survive per-component FDR. It offers a one-number check on whether each modality moved as a whole.

```{r sens-pca, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7}
pca <- safe_readRDS(here::here("figure-data/pca_intervention_effects_results.RDS"),
                    label = "pca_intervention_effects_results.RDS")
if (!is.null(pca) && is.data.frame(pca)) {
  pca <- pca %>% dplyr::filter(measure == "ATE" | is.na(measure)) %>%
    dplyr::mutate(
      sigFDR = factor(sigFDR, levels = c(0, 1)),
      tooltip = paste0(label_f,
                       "<br>", studytime, " | ", contrast,
                       "<br>ATE: ", round(est, 3),
                       " (", round(cil, 3), ", ", round(ciu, 3), ")",
                       "<br>q: ", signif(chi_pval_adj, 3))
    )
  p_pca <- ggplot2::ggplot(pca,
                           ggplot2::aes(y = stats::reorder(label_f, est), x = est,
                                        color = contrast, shape = sigFDR, text = tooltip)) +
    ggplot2::geom_point(position = ggplot2::position_dodge(width = 0.5), size = 2) +
    ggplot2::geom_errorbar(ggplot2::aes(xmin = cil, xmax = ciu), width = 0.2,
                           orientation = "y",
                           position = ggplot2::position_dodge(width = 0.5)) +
    ggplot2::geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
    ggplot2::scale_shape_manual(values = c(1, 19),
                                labels = c("Not significant","Significant (FDR)")) +
    ggplot2::scale_color_manual(values = tableau10) +
    ggplot2::facet_wrap(~ studytime) +
    ggplot2::labs(y = "Outcome-group PC1", x = "ATE on standardized PC1",
                  color = "Contrast", shape = NULL,
                  title = "Adjusted ATEs on outcome-group PC1 composites") +
    ggplot2::theme(legend.position = "bottom")
  htmltools::tagList(
    if (knitr::is_html_output() && requireNamespace("plotly", quietly = TRUE)) {
      conv <- try(plotly::ggplotly(p_pca, tooltip = "text"), silent = TRUE)
      if (!inherits(conv, "try-error")) conv else p_pca
    } else p_pca,
    nice_dt(pca %>% dplyr::select(studytime, contrast, outcome_group, label_f,
                                  est, cil, ciu, chi_pval, chi_pval_adj, sig, sigFDR) %>%
              dplyr::mutate(dplyr::across(where(is.numeric), ~ round(.x, 3))),
            caption = "PCA-based composite outcome ATEs (pooled intervention vs. control).")
  )
}
```

**Figure 2B.** Adjusted ATEs (95% CI) on the first principal component (PC1) of each milk-component class, pooled intervention vs. control, by study and visit. Effects are oriented positive for comparability across classes; filled symbols are significant after FDR correction. B-vitamin PC1 is increased across trials and timepoints, with no consistent effect on the other component classes.

### Arm-stratified PCA composites

```{r sens-pca-strat, echo=FALSE, message=FALSE, warning=FALSE, fig.height=7}
pca_strat <- safe_readRDS(here::here("results/pca_intervention_effects_results_arm_strat.RDS"),
                          label = "pca_intervention_effects_results_arm_strat.RDS")
if (!is.null(pca_strat) && is.data.frame(pca_strat)) {
  pca_strat <- pca_strat %>% dplyr::filter(measure == "ATE" | is.na(measure)) %>%
    dplyr::mutate(
      sigFDR  = factor(sigFDR, levels = c(0, 1)),
      tooltip = paste0(label_f,
                       "<br>", studytime, " | ", contrast,
                       "<br>ATE: ", round(est, 3),
                       " (", round(cil, 3), ", ", round(ciu, 3), ")",
                       "<br>q: ", signif(chi_pval_adj, 3))
    )
  p_pca_strat <- ggplot2::ggplot(pca_strat,
                                 ggplot2::aes(y = stats::reorder(label_f, est), x = est,
                                              color = contrast, shape = sigFDR, text = tooltip)) +
    ggplot2::geom_point(position = ggplot2::position_dodge(width = 0.5), size = 2) +
    ggplot2::geom_errorbar(ggplot2::aes(xmin = cil, xmax = ciu), width = 0.2,
                           orientation = "y",
                           position = ggplot2::position_dodge(width = 0.5)) +
    ggplot2::geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
    ggplot2::scale_shape_manual(values = c(1, 19),
                                labels = c("Not significant","Significant (FDR)")) +
    ggplot2::scale_color_manual(values = tableau10) +
    ggplot2::facet_wrap(~ studytime) +
    ggplot2::labs(y = "Outcome-group PC1", x = "ATE on standardized PC1",
                  color = "Contrast", shape = NULL,
                  title = "Arm-stratified PCA composite ATEs (per arm vs. control)") +
    ggplot2::theme(legend.position = "bottom")
  htmltools::tagList(
    if (knitr::is_html_output() && requireNamespace("plotly", quietly = TRUE)) {
      conv <- try(plotly::ggplotly(p_pca_strat, tooltip = "text"), silent = TRUE)
      if (!inherits(conv, "try-error")) conv else p_pca_strat
    } else p_pca_strat,
    nice_dt(pca_strat %>% dplyr::select(studytime, contrast, outcome_group, label_f,
                                        est, cil, ciu, chi_pval, chi_pval_adj, sig, sigFDR) %>%
              dplyr::mutate(dplyr::across(where(is.numeric), ~ round(.x, 3))),
            caption = "Arm-stratified PCA-based composite outcome ATEs.")
  )
}
```

## Pathway enrichment analyses

The pathway- and enrichment-level interpretation of the high-dimensional milk signals now lives in its own chapter, **[§9 *Reproducible pathway and enrichment analyses*](#reproducible-pathway-and-enrichment-analyses)**. That work covers MetaboAnalyst over-representation/MSEA of the primary, tertiary, and untargeted metabolome, Mummichog functional enrichment, and Gene Ontology enrichment of the milk proteome. It is built directly from the scripted MetaboAnalystR CSV outputs (`results/metaboanalyst/`, copied in by `port_results.R`) and covers manuscript **Tables S1–S6** and the enrichment panels of **Fig. 3B, Fig. 5B, and Fig. 6A–C**. It replaces an earlier placeholder here that loaded not-yet-produced `pathway_enrichment_*.RDS` objects.

The compositional fatty-acid enrichment of intervention-responsive triglycerides (manuscript **Table S3 / Fig. 5C**) is the compositional companion to the fat-adjusted triglyceride analysis in §8.3 above; its reproducible table is in [§9.5](#reproducible-pathway-and-enrichment-analyses).

### Network views of pathway and feature relationships

The network diagrams below summarise how intervention-responsive pathways and features relate to one another for the primary outcomes and the metabolome. Nodes are pathways or features, and edges denote shared members or co-occurrence, as defined in the upstream enrichment workflow.

```{r path-networks, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
net_primary <- safe_readRDS(here::here("figure-data/figures8_network_plots_primary.RDS"),
                            label = "figures8_network_plots_primary.RDS")
net_metab   <- safe_readRDS(here::here("figure-data/figures8_network_plots_metabolomics.RDS"),
                            label = "figures8_network_plots_metabolomics.RDS")
show_any(net_primary)
show_any(net_metab)
```

**Network views.** Intervention-responsive pathway/feature networks for the primary outcomes (top) and the metabolome (bottom). Nodes are pathways or features; edges denote shared members or co-occurrence.

## Infant growth analyses (Figure S1)

Here we report the effects of maternal nutritional supplementation on infant anthropometric outcomes within the IMiC subsamples of each trial: length-for-age z-score (LAZ), weight-for-length z-score (WLZ), weight-for-age z-score (WAZ), head-circumference-for-age z-score (HCAZ), and mid-upper-arm-circumference-for-age z-score (MUACZ). Estimates use the same covariate-adjusted TMLE framework as the primary milk-component analyses. These growth analyses act as within-trial sanity checks. They document whether the IMiC subsample is a plausible, representative window onto the parent trial, and they set the manuscript's HM-composition findings against the corresponding infant phenotype.

> **Negative-control note (ELICIT).** Infant azithromycin and infant nicotinamide were administered only after 6 months of age. Their effects on the 1-month and 5-month HM-collection visits are therefore null by design and act as a within-trial negative control on the analytic pipeline. Any non-null estimate in those contrasts should be read as noise.

```{r growth-forest, echo=FALSE, message=FALSE, warning=FALSE, fig.height=8}
g <- safe_readRDS(here::here("figure-data/primary_growth_plot.RDS"),
                  label = "primary_growth_plot.RDS")
if (!is.null(g) && is.data.frame(g)) {
  g <- g %>% dplyr::mutate(
    Measure = factor(Measure),
    contrast = factor(contrast),
    study   = factor(study, levels = c("Misame","MISAME-3","Mumta-LW","Vital","Elicit","ELICIT")) %>%
               droplevels(),
    sigFDR  = factor(sigFDR, levels = c(0, 1)),
    sig     = factor(sig, levels = c(0, 1)),
    tooltip = paste0(Measure, " @ ", age, " mo.",
                     "<br>Study: ", study,
                     "<br>Contrast: ", contrast,
                     "<br>ATE: ", round(est, 3),
                     " (", round(cil, 3), ", ", round(ciu, 3), ")",
                     "<br>q: ", signif(chi_pval_adj, 3))
  )
  p <- ggplot2::ggplot(g,
                       ggplot2::aes(x = factor(age), y = est, color = contrast,
                                    shape = sigFDR, text = tooltip)) +
    ggplot2::geom_point(position = ggplot2::position_dodge(width = 0.6), size = 2) +
    ggplot2::geom_errorbar(ggplot2::aes(ymin = cil, ymax = ciu),
                           width = 0.2,
                           position = ggplot2::position_dodge(width = 0.6)) +
    ggplot2::geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
    ggplot2::scale_shape_manual(values = c(1, 19),
                                labels = c("Not significant","Significant (FDR)")) +
    ggplot2::scale_color_manual(values = tableau10) +
    ggplot2::facet_grid(Measure ~ study, scales = "free") +
    ggplot2::labs(x = "Infant age at measurement (months)",
                  y = "Adjusted ATE on z-score",
                  color = "Contrast", shape = NULL,
                  title = "Intervention effects on infant anthropometry (IMiC subsamples)") +
    ggplot2::theme(legend.position = "bottom")
  out <- if (knitr::is_html_output() && requireNamespace("plotly", quietly = TRUE)) {
    conv <- try(plotly::ggplotly(p, tooltip = "text"), silent = TRUE)
    if (!inherits(conv, "try-error")) conv else p
  } else p
  out
}
```

**Figure S1.** Adjusted intervention effects (ATE on anthropometric z-score, with 95% CI) on infant length-for-age (LAZ), weight-for-length (WLZ), weight-for-age (WAZ), head-circumference-for-age (HCAZ), and MUAC-for-age (MUACZ), by study, randomized contrast, and infant age at measurement. Filled symbols are significant after FDR correction. In ELICIT, infant azithromycin and nicotinamide were administered after 6 months, so their effects at the 1- and 5-month visits are null by design and serve as a within-trial negative control.

```{r growth-table, echo=FALSE, message=FALSE, warning=FALSE}
if (!is.null(g) && is.data.frame(g)) {
  out <- g %>%
    dplyr::select(study, age, Measure, contrast, est, cil, ciu,
                  chi_pval, chi_pval_adj, sig, sigFDR) %>%
    dplyr::mutate(dplyr::across(where(is.numeric), ~ round(.x, 3)))
  nice_dt(out,
          caption = "Adjusted ATEs on infant anthropometric z-scores within the IMiC subsamples (Figure S1).")
}
```
