# Sensitivity analyses and additional outputs

```{r sens-setup, include=FALSE}
library(tidyverse)
library(here)
library(DT)
library(plotly)
source(here::here("functions.R"))
metadata_loaded <- safe_load(here::here("metadata/milk_component.Rdata"))

## One searchable table per outcome class of a results file.
class_tables <- function(res) {
  groups <- list(
    "Macronutrients"     = c(all_milk_components$macro, "total carbohydrate"),
    "Micronutrients"     = all_milk_components$micro,
    "B-vitamins"         = all_milk_components$bvit,
    "HMOs"               = all_milk_components$hmo,
    "Bioactive proteins" = all_milk_components$protein)
  out <- htmltools::tagList()
  for (g in names(groups)) {
    sub <- res %>% dplyr::filter(biomarker %in% groups[[g]])
    if (nrow(sub)) out <- htmltools::tagList(out, clean_tab(sub, caption = paste0(g, ".")))
  }
  out
}
```

## Unadjusted intervention effects

TMLE estimates without covariate adjustment, each randomized arm vs. control.

```{r sens-unadjusted, echo=FALSE, message=FALSE, warning=FALSE}
res_unadj <- safe_readRDS(here::here("results/unadjusted_intervention_effects_results_clean.RDS"),
                          label = "unadjusted_intervention_effects_results_clean.RDS")
if (!is.null(res_unadj) && isTRUE(metadata_loaded))
  class_tables(res_unadj %>% dplyr::filter(measure == "ATE") %>%
                 dplyr::mutate(biomarker = stringr::str_to_lower(biomarker)))
```

## Unscaled (natural-unit) intervention effects

Adjusted effects in each component's native units, for example g/100 mL for macronutrients, µg/L or mg/L for micronutrients, and nmol/mL for HMOs, pooled intervention vs. control.

```{r sens-unscaled, echo=FALSE, message=FALSE, warning=FALSE}
res_unscaled <- safe_readRDS(here::here("results/adjusted_combined_arms_intervention_effects_unscaled_results_clean.RDS"),
                             label = "adjusted_combined_arms_intervention_effects_unscaled_results_clean.RDS")
if (!is.null(res_unscaled) && isTRUE(metadata_loaded))
  class_tables(res_unscaled %>% dplyr::filter(measure == "ATE") %>%
                 dplyr::mutate(biomarker = stringr::str_to_lower(biomarker)))
```

### Targeted metabolites (tertiary, native units) {#native-units-tertiary}

For all 581 targeted metabolites, in µM: control and intervention means, the ATE and 95% CI, percent change, native-unit *P* and *Q*, and the standardized ATE and *Q* ([`targeted_metabolites_native_units.csv`](data-files/targeted_metabolites_native_units.csv)).

```{r sens-unscaled-tertiary, echo=FALSE, message=FALSE, warning=FALSE}
nu <- safe_read_csv(here::here("data-files/targeted_metabolites_native_units.csv"))
if (!is.null(nu) && nrow(nu)) {
  num <- vapply(nu, is.numeric, logical(1))
  nu[num] <- lapply(nu[num], function(x) signif(x, 3))
  nice_dt(nu, caption = "Targeted metabolites in native units.", page_length = 15)
}
```

### Table S1 {#native-units-table}

Table S1 in full: one row per trial, visit, and outcome, with each arm's sample size and raw mean, SD, median, and quartiles in native units side by side, and the native-unit ATE, 95% CI, *P*, and *Q* ([`table_s1_primary_secondary_native_units.csv`](data-files/table_s1_primary_secondary_native_units.csv)).

```{r sens-table-s1, echo=FALSE, message=FALSE, warning=FALSE}
tab_s1 <- safe_read_csv(here::here("data-files/table_s1_primary_secondary_native_units.csv"))
if (!is.null(tab_s1)) nice_dt(tab_s1, caption = "Table S1.")
```

## Treatment-arm prediction

Cross-validated AUC (CV-AUC) of SuperLearner models that predict randomized arm from each class of milk components, or from all classes together, by trial, arm, and visit; 0.5 is chance.

```{r sens-sl, echo=FALSE, message=FALSE, warning=FALSE}
sl <- safe_readRDS(here::here("figure-data/SL_vim_plot_data.RDS"), label = "SL_vim_plot_data.RDS")
if (!is.null(sl) && is.data.frame(sl)) {
  out <- sl %>%
    dplyr::filter(!is.na(cvAUC)) %>%
    dplyr::mutate(study = study_name(studyid),
                  arm = sub("Pre+Postnatal BEP", "Pre + Postnatal BEP",
                            gsub("\\s*\n\\s*", " ", as.character(arm_f)), fixed = TRUE)) %>%
    dplyr::select(study, visit, arm, predictors = group, N_obs, cvAUC, ci.lb, ci.ub) %>%
    dplyr::mutate(dplyr::across(where(is.numeric), ~ round(.x, 3)))
  nice_dt(out, caption = "CV-AUC for treatment-arm prediction.")
}
```

## Principal-component scores

Effects on the first principal component (PC1) of each class of milk components, pooled intervention vs. control, then by randomized arm.

```{r sens-pca, echo=FALSE, message=FALSE, warning=FALSE}
pca <- safe_readRDS(here::here("figure-data/pca_intervention_effects_results.RDS"),
                    label = "pca_intervention_effects_results.RDS")
if (!is.null(pca) && is.data.frame(pca)) {
  out <- pca %>% dplyr::filter(measure == "ATE" | is.na(measure)) %>%
    dplyr::mutate(panel = studytime_label(studytime)) %>%
    dplyr::select(panel, contrast, outcome_group, label_f, est, cil, ciu, chi_pval, chi_pval_adj, sig, sigFDR) %>%
    dplyr::mutate(dplyr::across(where(is.numeric), ~ round(.x, 3)))
  nice_dt(out, caption = "PC1 effects, pooled intervention vs. control.")
}
```

```{r sens-pca-strat, echo=FALSE, message=FALSE, warning=FALSE}
pca_strat <- safe_readRDS(here::here("results/pca_intervention_effects_results_arm_strat.RDS"),
                          label = "pca_intervention_effects_results_arm_strat.RDS")
if (!is.null(pca_strat) && is.data.frame(pca_strat)) {
  out <- pca_strat %>%
    dplyr::filter(measure == "ATE" | is.na(measure), contrast != "Control") %>%
    dplyr::mutate(panel = studytime_label(studytime)) %>%
    dplyr::select(panel, contrast, outcome_group, label_f, est, cil, ciu, chi_pval, chi_pval_adj, sig, sigFDR) %>%
    dplyr::mutate(dplyr::across(where(is.numeric), ~ round(.x, 3)))
  nice_dt(out, caption = "PC1 effects, each randomized arm vs. control.")
}
```

## Infant growth

Adjusted effects on infant length-for-age (LAZ) and weight-for-length (WLZ) z-scores in the IMiC subsample of each trial, each randomized arm vs. control, at birth (MISAME-III only) and at 3 and 6 months.

```{r growth-table, echo=FALSE, message=FALSE, warning=FALSE}
g <- safe_readRDS(here::here("figure-data/figureS1_growth_plot_data.RDS"),
                  label = "figureS1_growth_plot_data.RDS")
if (!is.null(g) && is.data.frame(g)) {
  out <- g %>%
    dplyr::mutate(contrast = gsub("\\s*\n\\s*", " ", as.character(contrast))) %>%
    dplyr::select(study, age, Measure, contrast, est, cil, ciu, pval, sig) %>%
    dplyr::mutate(dplyr::across(where(is.numeric), ~ round(.x, 3)))
  nice_dt(out, caption = "Effects on infant growth z-scores.")
}
```
