# Reproducible pathway and enrichment analyses

Pathway and enrichment analyses of the milk effects in Sections 4 and 5, run offline for each trial, visit, contrast, and direction of effect, with the features at *P* < 0.05 as input. Negative enrichment ratios mark down-regulated pathways.

```{r path-setup, include=FALSE}
library(tidyverse)
library(here)
library(DT)
library(plotly)
source(here::here("functions.R"))

## Read an enrichment CSV, degrading gracefully (also for an empty file).
read_enrich <- function(path) {
  full <- here::here(path)
  if (!file.exists(full)) {
    cat(sprintf("*Results file not found at `%s`. Run `port_results.R` to copy it from the analysis repository.*", path))
    return(invisible(NULL))
  }
  df <- tryCatch(utils::read.csv(full, check.names = FALSE), error = function(e) NULL)
  if (is.null(df) || ncol(df) == 0) return(data.frame())
  for (s in intersect(c("study", "Study"), names(df))) df[[s]] <- study_name(df[[s]])
  df
}

## Searchable table, with P and FDR columns in scientific notation so small
## values are not rounded to 0.
enrich_dt <- function(df, caption = NULL, page_length = 15) {
  if (is.null(df)) return(invisible(NULL))
  if (!nrow(df))  return(htmltools::tags$em("No rows."))
  sci <- intersect(c("raw_p", "fdr_native", "fdr", "p_value", "P-value", "FDR",
                     "pvalue", "p.adjust", "qvalue"), names(df))
  for (cc in sci) if (is.numeric(df[[cc]])) df[[cc]] <- formatC(df[[cc]], format = "e", digits = 2)
  nice_dt(df, caption = caption, page_length = page_length)
}
```

## Interactive pathway explorer {#pathway-explorer-app}

Every enrichment table in the supplementary materials (Tables S2–S7), including the arm-stratified runs, as a dot plot and table, filtered by analysis, trial, and visit. The app runs R in the browser, so the first load takes a moment. [Open it in a new tab](apps/pathway-explorer/).

```{r pathway-explorer-embed, echo=FALSE, results='asis'}
if (knitr::is_html_output()) {
  cat('<iframe src="apps/pathway-explorer/" title="IMiC Pathway-Enrichment Explorer" ',
      'width="100%" height="820" loading="lazy" ',
      'style="border:1px solid #ddd; border-radius:4px;"></iframe>\n', sep = "")
} else {
  cat("*The interactive pathway explorer is available in the web version of this online resource.*\n")
}
```

## Targeted-metabolome enrichment (Table S3) {#path-tertiary-msea}

Table S3 in full: over-representation of the SMPDB small-molecule and lipid metabolite sets among the targeted metabolites, computed in MetaboAnalystR against the whole measured human metabolome as background ([`table_s3_tertiary_msea_full.csv`](data-files/table_s3_tertiary_msea_full.csv)). `fdr_native` is the Benjamini–Hochberg FDR that MetaboAnalystR reports for each run.

```{r path-tertiary-msea-dual, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/tertiary_msea/tertiary_msea_dual.csv"),
          caption = "Table S3.")
```

## Pathways enriched on both metabolomics platforms {#path-cross-platform}

Pathways with *Q* < 0.05 in the targeted (Table S3) or untargeted (Table S5) enrichment and *P* < 0.05 in the other, listing every trial, visit, contrast, and direction with *P* < 0.05 on either platform.

```{r path-cross-platform, echo=FALSE, message=FALSE, warning=FALSE}
t3 <- read_enrich("results/metaboanalyst/tertiary_msea/tertiary_msea_dual.csv")
u5 <- read_enrich("results/metaboanalyst/untargeted_msea/untargeted_msea_combined.csv")
if (!is.null(t3) && !is.null(u5) && nrow(t3) && nrow(u5)) {
  key <- function(x) gsub("[^a-z0-9]", "", tolower(x))
  t3$key <- key(t3$pathway); u5$key <- key(u5$pathway)
  both <- union(intersect(t3$key[t3$fdr_native < 0.05], u5$key[u5$raw_p < 0.05]),
                intersect(u5$key[u5$fdr_native < 0.05], t3$key[t3$raw_p < 0.05]))
  xp <- dplyr::bind_rows(
    dplyr::mutate(t3[t3$key %in% both & t3$raw_p < 0.05, ], platform = "Targeted (Table S3)"),
    dplyr::mutate(u5[u5$key %in% both & u5$raw_p < 0.05, ], platform = "Untargeted (Table S5)")) %>%
    dplyr::mutate(pathway = stats::ave(pathway, key, FUN = function(p) p[1])) %>%
    dplyr::select(pathway, platform, study, timepoint, contrast, direction, raw_p, fdr_native) %>%
    dplyr::arrange(tolower(pathway), platform, study, timepoint)
  enrich_dt(xp, caption = "Pathways enriched on both platforms.", page_length = 25)
}
```

## Untargeted-metabolome Mummichog enrichment (Table S6) {#path-mummichog}

Table S6 in full: Mummichog pathway enrichment of the untargeted milk features, matched to the `human_mfn` network at 10 ppm, by trial, visit, contrast, direction, and ionization mode, for the pathways with *P* < 0.05 ([`table_s6_mummichog_full.csv`](data-files/table_s6_mummichog_full.csv)). `FDR` is Benjamini–Hochberg-adjusted within each run.

```{r path-mummichog-tab, echo=FALSE, message=FALSE, warning=FALSE}
mum <- read_enrich("results/metaboanalyst/mummichog_s5/milk_mummichog_tableS5.csv")
if (!is.null(mum) && nrow(mum))
  enrich_dt(mum %>% dplyr::filter(.data[["P-value"]] < 0.05) %>% dplyr::arrange(.data[["P-value"]]),
            caption = "Table S6.")
```

### Putative feature annotations {#path-mummichog-annotation}

Mass-based putative names (MSI level 3) of the FDR-significant untargeted milk features.

```{r path-mummichog-annot-full, echo=FALSE, message=FALSE, warning=FALSE}
annot <- read_enrich("results/metaboanalyst/mummichog_s5/milk_mummichog_annotation_finalized.csv")
if (!is.null(annot) && nrow(annot))
  enrich_dt(annot, caption = "Putative annotations of the FDR-significant untargeted features.", page_length = 10)
```

## Proteome Gene Ontology enrichment (Table S7) {#path-go}

Table S7 in full: over-representation of Gene Ontology Biological Process terms among the proteins with *P* < 0.05, using `clusterProfiler::enricher` with each measured UniProt protein counted once and all proteins measured in the trial and contrast as background, for the terms with *P* < 0.05 and Storey *q* < 0.20 ([`table_s7_proteomics_go_full.csv`](data-files/table_s7_proteomics_go_full.csv)). `fdr` is Benjamini–Hochberg-adjusted. The second table repeats the terms with *Q* < 0.05 with the contributing UniProt accessions.

```{r path-go-table-s7, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/proteomics_go/proteomics_go_tableS6.csv"),
          caption = "Table S7.")
```

```{r path-go-tab, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/proteomics_go/proteomics_go_uniprot_fdrsig.csv"),
          caption = "Terms with Q < 0.05, with gene ratio, background ratio, rich factor, z-score, Storey q, and UniProt accessions.")
```
