# Reproducible pathway and enrichment analyses

This chapter reports the **pathway- and enrichment-level interpretation** of the per-component milk signals shown in §4 and §5, converting large tables of individual features into a smaller set of biologically interpretable pathways. It replaces the earlier manual `metaboanalyst.ca` point-and-click workflow with a fully scripted **MetaboAnalystR** pipeline (`src/metaboanalyst/` in the upstream analysis repository), so every table below can be regenerated deterministically from the intervention-effect estimates.

Each analysis is reported twice: for the **combined-arm** contrast (all maternal nutritional-intervention arms pooled vs. control, the framing used in the printed manuscript figures) and for the **arm-stratified** contrasts (each randomised arm vs. control, the supplement view). Enrichment tests use a permissive foreground (features at unadjusted *p* < 0.05, split by direction of effect) evaluated against the assay-specific background, then Benjamini–Hochberg FDR correction of the pathway *p*-values within each study × timepoint × contrast × direction stratum (and, for Mummichog, ionisation mode). A pathway is flagged **significant** at *Q* < 0.05 (`fdr_native` / `fdr` / `FDR` column, depending on the table).

> **Interpretation note.** These enrichment results summarise and contextualise the prespecified per-component analyses; they are not themselves a separate confirmatory test. Over-representation depends on the choice of background metabolite set (see §9.2), and pathway membership in reference databases is incomplete, so absence of a pathway is not evidence of absence of biology. Read the results as a hypothesis-generating map of where the intervention signal concentrates.

> **Coverage: all six manuscript enrichment tables (S1–S6) are reproduced here** as searchable tables, in manuscript order: primary ORA (**Table S1**, §9.1), the ORA-vs-impact comparison (§9.2), the empty secondary result (§9.3), tertiary MSEA (**Table S2**, §9.4), triglyceride fatty-acid composition (**Table S3**, §9.5), untargeted-metabolome MSEA (**Table S4**, §9.6), milk untargeted Mummichog (**Table S5**, §9.7), and milk-proteome GO (**Table S6**, §9.8).

```{r path-setup, include=FALSE}
library(tidyverse)
library(here)
library(DT)
library(plotly)
source(here::here("functions.R"))
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 14))
knitr::opts_chunk$set(fig.width = 12, fig.height = 8, dpi = 110)

## Read an enrichment CSV, degrading gracefully. Unlike safe_read_csv this also
## survives an empty (0-byte / header-only) file, since some cells of the enrichment
## grid legitimately return no pathways (e.g. HMOs/bioactives, §9.3).
read_enrich <- function(path) {
  full <- here::here(path)
  if (!file.exists(full)) {
    cat(sprintf("*Results file not found at `%s`. Run `port_results.R` to copy it from the upstream analysis repo.*", path))
    return(invisible(NULL))
  }
  df <- tryCatch(utils::read.csv(full, check.names = FALSE), error = function(e) NULL)
  if (is.null(df) || ncol(df) == 0) return(data.frame())
  df
}

## Present p-value / FDR columns in scientific notation so very small enrichment
## p-values (e.g. 3.7e-11) are not rounded to 0.000 by nice_dt().
p_cols <- c("raw_p", "fdr_native", "fdr_bh_pooled", "fdr", "impact",
            "p_value", "P-value", "FDR", "Enrichment Ratio", "enrichment_ratio",
            "fold_enrichment")
enrich_dt <- function(df, caption = NULL, page_length = 15) {
  if (is.null(df)) return(invisible(NULL))
  if (!nrow(df))  return(htmltools::tags$em("No enriched pathways in this stratum."))
  sci <- intersect(c("raw_p","fdr_native","fdr_bh_pooled","fdr","p_value","P-value","FDR"), names(df))
  for (cc in sci) df[[cc]] <- formatC(df[[cc]], format = "e", digits = 2)
  nice_dt(df, caption = caption, page_length = page_length)
}
```

## Interactive pathway explorer {#pathway-explorer-app}

The app below spans all six enrichment analyses in this chapter (manuscript Tables S1–S6). Pick an analysis, restrict to FDR-significant pathways or show everything, and filter by study and timepoint; the dot plot ranks pathways by −log10(FDR) and the table gives the full numbers. Like the effect explorer in §5, it runs entirely in your browser (WebAssembly R), so the first load fetches the runtime and is then instant. [Open it full-screen in a new tab](apps/pathway-explorer/).

```{r pathway-explorer-embed, echo=FALSE, results='asis'}
if (knitr::is_html_output()) {
  cat('<iframe src="apps/pathway-explorer/" title="IMiC Pathway-Enrichment Explorer" ',
      'width="100%" height="820" loading="lazy" ',
      'style="border:1px solid #ddd; border-radius:4px;"></iframe>\n', sep = "")
} else {
  cat("*The interactive pathway explorer is available in the HTML build of this appendix.*\n")
}
```

## Primary outcomes: MetaboAnalyst enrichment of micronutrient / B-vitamin features {#path-primary}

These are the pathway summaries of the **primary-outcome** signals (macronutrients, micronutrients, B-vitamins). They correspond to manuscript **Table S1** and the enrichment panels of manuscript **Fig. 3B**. Over-representation analysis (ORA) is run per study × timepoint × contrast × direction against the targeted-panel background; the `significant` column flags pathways with `fdr_native` < 0.05.

The clearest result is **nicotinate and nicotinamide metabolism**, driven by the ELICIT nicotinamide-during-lactation arm (up-regulated at both the 1-month and 5-month visits; 5/15 pathway metabolites hit, hypergeometric *p* ≈ 3.7 × 10⁻¹¹, *Q* ≈ 3.0 × 10⁻⁹), and also FDR-significant in the BEP trials (Mumta-LW, MISAME-III). Riboflavin, vitamin B~6~, and thiamine metabolism appear as secondary FDR-significant pathways across BEP trials and timepoints.

```{r path-primary-fig-combined, echo=FALSE, message=FALSE, warning=FALSE}
safe_img("figures/figure4_combined_arms.png", label = "figure4_combined_arms.png")
```

***Figure (manuscript Fig. 3B, combined arms).** Volcano and pathway-enrichment composite for the primary micronutrient / B-vitamin outcomes, all maternal-intervention arms pooled vs. control. Points are individual components; enriched pathways are annotated. Nicotinate/nicotinamide metabolism (ELICIT) and riboflavin / vitamin B~6~ metabolism (BEP trials) are the leading FDR-significant pathways.*

```{r path-primary-tab-combined, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/primary_combined/primary_combined_supplementary_table.csv"),
          caption = "Primary-outcome MetaboAnalyst ORA, combined intervention arms vs. control (manuscript Table S1). significant = fdr_native < 0.05.")
```

### Arm-stratified

```{r path-primary-fig-strat, echo=FALSE, message=FALSE, warning=FALSE}
safe_img("figures/figure4_stratified_supplement.png", label = "figure4_stratified_supplement.png")
```

***Figure (manuscript Fig. 3B, arm-stratified supplement).** As above, with each randomised arm contrasted separately against its trial control.*

```{r path-primary-tab-strat, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/primary_stratified/primary_stratified_supplementary_table.csv"),
          caption = "Primary-outcome MetaboAnalyst ORA, arm-stratified contrasts vs. control.")
```

### Direction-split ORA (up- vs down-regulated; Fig. S9A)

These standalone tables run the over-representation separately on the up- and down-regulated primary feature lists (combined arms). The §9.1 ORA already splits the *foreground* by direction internally; here that split is reported directly, recovering **nicotinate and nicotinamide metabolism** among the up-regulated features (ELICIT) and **lactose synthesis and superoxide degradation** among the down-regulated features (MISAME-III).

```{r path-primary-dirsplit-fig, echo=FALSE, message=FALSE, warning=FALSE}
safe_img("figures/figureS_primary_ora_by_direction.png", label = "figureS_primary_ora_by_direction.png")
```

***Figure S9A.** Primary-outcome ORA split by direction of effect, combined arms. Enriched pathways among up- vs. down-regulated primary features. Nicotinate/nicotinamide metabolism (ELICIT) dominates the up-regulated side; lactose synthesis and superoxide degradation (MISAME-III) appear among the down-regulated side. Points are coloured by study, non-significant pathways are grey, and enrichment was tested per study, time point, and direction with Benjamini–Hochberg correction of the pathway-level p-values. (Numbered **Fig. S9A** in the submitted supplement. Dropped from the print supplement in the 2026-08-26 revision, where the searchable tables below became the primary record.)*

```{r path-primary-ora-up, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/primary_combined/primary_ora_upregulated.csv"),
          caption = "Primary-outcome ORA, up-regulated features (combined arms).")
```

```{r path-primary-ora-down, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/primary_combined/primary_ora_downregulated.csv"),
          caption = "Primary-outcome ORA, down-regulated features (combined arms).")
```

## Primary outcomes: over-representation vs. quantitative pathway/impact analysis {#path-ora-vs-pathway}

MetaboAnalyst supports two related enrichment styles, and comparing them makes the **background dependence** of enrichment explicit:

- **Over-representation analysis (ORA)**, the §9.1 tables, tests whether the significant features are unexpectedly concentrated in a pathway, using the **restricted targeted-panel** as the background universe.
- **Pathway / impact analysis**, the table below, additionally weights each pathway by the topological position of the hit metabolites and evaluates it against the **full SMPDB** background.

Both recover **nicotinate and nicotinamide metabolism** as the leading primary-outcome pathway. The impact analysis additionally flags **vitamin B~6~ metabolism**; because that difference tracks the larger background set rather than a distinct data signal, we show the two side by side so the dependence stays visible.

```{r path-ora-pathway-tab, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/primary_combined_pathway/primary_combined_pathway_all_cells.csv"),
          caption = "Primary-outcome quantitative pathway/impact analysis (full SMPDB background), combined arms. impact = pathway topology score; fdr = BH-adjusted pathway p-value.")
```

## Secondary outcomes: HMOs and bioactive proteins {#path-secondary}

Enrichment was also attempted for the secondary outcomes (human milk oligosaccharides and bioactive proteins). These returned **no enriched metabolite-set pathways**: HMOs and milk bioactive proteins are largely absent from the SMPDB metabolite-set libraries, so the mapping step leaves too few annotated features to test. We report the empty result here for completeness; it is not a null biological finding.

```{r path-secondary-tab, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/secondary_combined/secondary_combined_all_cells.csv"),
          caption = "Secondary-outcome enrichment, combined arms (empty; HMOs/bioactives are not represented in SMPDB metabolite sets).")
enrich_dt(read_enrich("results/metaboanalyst/secondary_stratified/secondary_stratified_all_cells.csv"),
          caption = "Secondary-outcome enrichment, arm-stratified contrasts (also empty, for the same reason).")
```

## Tertiary outcomes: targeted-metabolome MSEA {#path-tertiary}

Metabolite-set enrichment analysis (MSEA) of the tertiary **Biocrates MxP Quant 500** targeted panel summarises the per-metabolite effects from §5 at the pathway level. It corresponds to manuscript **Table S2** and **Fig. 5B**.

Under BEP supplementation, breastmilk metabolism undergoes **time-dependent reprogramming**. In the early postpartum period the enrichment is dominated by broad **down-regulation of amino-acid and nitrogen-handling pathways** (arginine–proline, glycine–serine, methionine, and glutamate metabolism, and ammonia recycling) together with **mitochondrial energy metabolism** (fatty-acid β-oxidation, the citric-acid cycle, the electron-transport chain, ketone-body metabolism, and carnitine synthesis). By 3–4 months several of these pathways reverse, showing enrichment of mitochondrial respiration, lipid oxidation, and ketone metabolism. (The distinct down-regulation of *de novo fatty-acid biosynthesis* is an **untargeted**-metabolome finding reported in §9.6, not a targeted-panel MSEA result.)

### MSEA: manuscript Table S2 (dual-library run)

The table below **is** manuscript Table S2: the dual-library MSEA that combines the SMPDB
small-molecule and lipid metabolite-set libraries, reported for every pathway × study × visit ×
contrast × direction cell, unfiltered and sorted by raw *p*. All 463 rows, 52 of them FDR-significant.
The print supplement externalises the same 463 rows as a machine-readable file; both it and the
table below are written from `results/metaboanalyst/tertiary_msea/tertiary_msea_dual.csv`, so they
carry identical numbers. **Download:** [`table_s2_tertiary_msea_full.csv`](appendix-data/table_s2_tertiary_msea_full.csv).

```{r path-tertiary-msea-dual, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/tertiary_msea/tertiary_msea_dual.csv"),
          caption = "Tertiary targeted-metabolome MSEA, dual-library run across all arm contrasts: manuscript Table S2 (463 rows, 52 FDR-significant). klass = which metabolite-set library the pathway comes from; fdr_native = BH-adjusted enrichment p-value.")
```

### MSEA: single-library pooled-arm and per-arm views

These two tables are **separate, single-library runs** of the same targeted panel, kept here as the
pooled-arm and per-arm views. They are *not* alternative renderings of Table S2 and their numbers do
not match it: they use a different metabolite-set background (the `total` column differs row for row),
so pathway *p*- and *q*-values differ throughout. Cite Table S2 above, not these, when reconciling
against the printed manuscript or Fig. 5B.

```{r path-tertiary-msea-combined, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/tertiary_msea/tertiary_msea_combined.csv"),
          caption = "Tertiary targeted-metabolome MSEA, single-library run, combined intervention arms vs. control (supplementary view, NOT manuscript Table S2). fdr_native = BH-adjusted enrichment p-value.")
```

```{r path-tertiary-msea-strat, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/tertiary_msea/tertiary_msea_stratified.csv"),
          caption = "Tertiary targeted-metabolome MSEA, single-library run, arm-stratified contrasts vs. control (supplementary view, NOT manuscript Table S2).")
```

### ORA counterpart

The over-representation view of the same tertiary features, with the same column structure as §9.1:

```{r path-tertiary-ora-combined, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/tertiary_combined/tertiary_combined_supplementary_table.csv"),
          caption = "Tertiary targeted-metabolome over-representation analysis, combined arms.")
```

```{r path-tertiary-ora-strat, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/tertiary_stratified/tertiary_stratified_supplementary_table.csv"),
          caption = "Tertiary targeted-metabolome over-representation analysis, arm-stratified contrasts.")
```

### Direction-split ORA (up- vs down-regulated; Fig. S9B)

As for the primary outcomes, these tables run the over-representation separately on the up- and down-regulated tertiary (targeted-metabolite) feature lists (combined arms): the per-direction companion to the MSEA and ORA views above.

```{r path-tertiary-dirsplit-fig, echo=FALSE, message=FALSE, warning=FALSE}
safe_img("figures/figureS_tertiary_ora_by_direction.png", label = "figureS_tertiary_ora_by_direction.png")
```

***Figure S9B.** Tertiary targeted-metabolome ORA split by direction of effect, combined arms. Enriched pathways among up- vs. down-regulated tertiary features, with the same study colouring, grey non-significant points, and per-study/time-point/direction BH correction as the primary-outcome panel above. (Numbered **Fig. S9B** in the submitted supplement. Dropped from the print supplement in the 2026-08-26 revision; see the note on that panel.)*

```{r path-tertiary-ora-up, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/tertiary_combined/tertiary_ora_upregulated.csv"),
          caption = "Tertiary-outcome ORA, up-regulated features (combined arms).")
```

```{r path-tertiary-ora-down, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/tertiary_combined/tertiary_ora_downregulated.csv"),
          caption = "Tertiary-outcome ORA, down-regulated features (combined arms).")
```

## Triglyceride fatty-acid composition {#path-tg-fa}

The compositional companion to the fat-adjusted triglyceride analysis in §8.3, corresponding to manuscript **Table S3** and **Fig. 5C**. For each study × timepoint × contrast × direction stratum, the fatty-acyl moieties carried by the intervention-responsive triglycerides are compared against the background TG pool (a Wilcoxon test on within-TG proportional representation across the sn-1/sn-2/sn-3 positions). This identifies fatty acids **compositionally enriched or depleted** among the BEP-responsive triglycerides, *not* absolute fatty-acid concentrations. The consistent signal is enrichment of **polyunsaturated fatty acids**, including α-linolenic (18:3n-3), γ-linolenic, and related PUFAs, among the triglycerides increased under BEP across MISAME-III contrasts and timepoints. The signal resolves in the **arm-stratified** analysis (the combined-arm table is empty).

```{r path-tg-fa-tab, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/triglyceride_fa/triglyceride_fa_composition_stratified.csv"),
          caption = "Compositional fatty-acid enrichment of intervention-responsive triglycerides, arm-stratified contrasts (manuscript Table S3 / Fig. 5C). fdr_native = BH-adjusted p; positive enrichment_ratio = fatty acid over-represented among up-regulated TGs.")
```

## Untargeted-metabolome MSEA {#path-untargeted-msea}

Metabolite-set enrichment of the **untargeted** milk-metabolome feature lists, corresponding to manuscript **Table S4** and **Fig. 6A**. This is the named-metabolite-set counterpart to the Mummichog view in §9.7; both operate on the same untargeted features, but this enrichment tests curated metabolite sets on name-mapped compounds whereas Mummichog infers pathways directly from *m/z*. The table is read **verbatim from the 14 MetaboAnalyst enrichment exports** (one per study × timepoint × direction cell) downloaded for the manuscript, rather than regenerated by a script, and reproduces the **submitted signature**: the recurring pathways are **Fatty Acid Biosynthesis, Warburg Effect, Urea Cycle, Citric Acid Cycle, and Mitochondrial Electron Transport Chain**, with points from **both MISAME-III and Mumta-LW** (not MISAME-III alone). Backgrounds are **per-study name-mapped reference metabolomes** (246 compounds for MISAME-III, 234 for the ELICIT/Mumta-LW–Vital platform). Labels and thresholds are on the **raw p-value** (grey at *P* < 0.05, green at *P* < 0.01); **no pathway clears FDR** under this reproduction (the smallest `fdr_native` is ≈ 0.15). The strongest raw-p signals are a co-ordinated *up*-regulation of Warburg Effect / Citric Acid Cycle / Urea Cycle / mitochondrial ETC in Mumta-LW at 1.5 mo. and a *down*-regulation of Fatty Acid Biosynthesis in MISAME-III.

> **Reproducibility note.** This enrichment is **compound-name-matching- and background-dependent** and is *not* scriptably reproducible from the raw features (it reads MetaboAnalyst web-tool exports, and the mapped-name set and reference background are fixed at export time). The identification-robust untargeted view is the Mummichog analysis in §9.7, which works directly from *m/z*.

```{r path-untargeted-msea-tab, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/untargeted_msea/untargeted_msea_combined.csv"),
          caption = "Untargeted-metabolome enrichment across the 14 MetaboAnalyst export cells (manuscript Table S4 / Fig. 6A). study/timepoint/direction identify the cell; raw_p is the reported p-value (label/threshold basis); fdr_native = BH-adjusted p (none < 0.05); enrichment_ratio = hits / expected.")
```

## Exploratory untargeted milk metabolome: Mummichog functional enrichment {#path-mummichog}

Mummichog infers enriched pathways directly from mass-spectral features, without prior compound identification, by matching *m/z* to putative compounds within each ionisation mode. Annotation uses a **10 ppm** database tolerance against the `human_mfn` reference network, and pathways are tested at *p* < 0.05. This is the exploratory **untargeted-metabolome** enrichment. It corresponds to manuscript **Table S5** and **Fig. 6B**, and is reported by study × timepoint × contrast × direction × ionisation mode.

> **Two tolerances, one number apart.** The **10 ppm** here is a *compound-annotation* tolerance (matching a feature's *m/z* to a database compound). It is a different quantity from the **±25 ppm** *cross-platform feature-matching* tolerance used in §10.2 (matching a feature in one dataset to a feature in another). A reviewer comparing the two chapters should not read the difference as an inconsistency.

The dominant signature is **down-regulation of polyunsaturated-fatty-acid and lipid pathways** (de novo fatty-acid biosynthesis, fatty-acid activation and metabolism, arachidonic-acid and linoleate metabolism, and leukotriene / prostaglandin formation), concentrated in the MISAME-III BEP contrasts. The full grid is large (9,520 pathway × stratum rows), so the table below is filtered exactly as manuscript **Table S5** is: to the strata significant at **raw *P* < 0.05**, sorted by *P* ascending, giving 505 rows, the set underlying Fig. 6B. The `FDR` column is retained so the more stringent *Q* < 0.05 frontier (20 strata) can be read off directly. **Download:** [`table_s5_mummichog_full.csv`](appendix-data/table_s5_mummichog_full.csv).

```{r path-mummichog-note, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
mum <- read_enrich("results/metaboanalyst/mummichog_s5/milk_mummichog_tableS5.csv")
## Match the printed Table S5 exactly: raw P < 0.05, sorted by P ascending.
## (Filtering to FDR < 0.05 instead would show only 20 of these 505 rows and
## would not correspond to any table in the manuscript.)
mum_sig <- if (!is.null(mum) && nrow(mum))
  mum %>% dplyr::filter(.data[["P-value"]] < 0.05) %>% dplyr::arrange(.data[["P-value"]]) else NULL
if (!is.null(mum_sig) && nrow(mum_sig))
  cat(sprintf("\n*Showing %s pathway strata significant at raw P < 0.05, of %s total; manuscript Table S5. Of these, %s also clear FDR Q < 0.05.*\n\n",
              format(nrow(mum_sig), big.mark = ","), format(nrow(mum), big.mark = ","),
              format(sum(mum_sig[["FDR"]] < 0.05), big.mark = ",")))
```

```{r path-mummichog-tab, echo=FALSE, message=FALSE, warning=FALSE}
## The widget must be the chunk's last auto-printed value: wrapping it in
## print() inside a block sends it to the viewer instead of the document, which
## is why this table silently failed to appear in earlier builds.
if (!is.null(mum_sig) && nrow(mum_sig))
  enrich_dt(mum_sig,
            caption = "Milk untargeted-metabolome Mummichog enrichment, strata significant at raw P < 0.05 (manuscript Table S5 / Fig. 6B; 505 rows). FDR = BH-adjusted enrichment p-value.")
```

### Feature-level putative annotation {#path-mummichog-annotation}

The finalized per-feature annotation of the same untargeted milk features complements the pathway view above. Across **1,034 FDR-significant** features, organised by direction × ionisation mode, **243 carry a name** (48 curated + 195 Mummichog-putative) and the remainder are unresolved; the strongest named up-regulated signal is a vitamin/cofactor cluster (pantothenate/B5, nicotinamide, ascorbate). Every name is **putative**, mass-based at MSI level ~3, not a confirmed identification.

```{r path-mummichog-annot-summary, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/mummichog_s5/milk_mummichog_annotation_summary.csv"),
          caption = "FDR-significant milk features by direction and ionization mode: named vs unresolved.")
```

```{r path-mummichog-annot-full, echo=FALSE, message=FALSE, warning=FALSE}
annot <- read_enrich("results/metaboanalyst/mummichog_s5/milk_mummichog_annotation_finalized.csv")
if (!is.null(annot) && nrow(annot)) {
  DT::datatable(annot, filter = "top", rownames = FALSE,
                options = list(pageLength = 10, scrollX = TRUE),
                caption = "Finalized putative annotation of FDR-significant untargeted milk features.")
}
```

## Exploratory untargeted milk proteome: GO functional enrichment {#path-go}

Gene Ontology (Biological Process) over-representation of the intervention-associated **untargeted milk proteomic** features corresponds to manuscript **Table S6** and **Fig. 6C**. Enrichment uses a **UniProt-native** over-representation test: `clusterProfiler::enricher` run against a custom UniProt→GO-BP annotation map, with each measured protein counted **once** and a **per-cell measured-proteome background** (the set of proteins actually quantified in that study × timepoint × contrast). This supersedes the earlier UniProt→Entrez→`enrichGO` gene-level workflow, whose identifier-mapping step distorted the background and produced a now-retracted "three-domain" result.

The corrected signature is a **predominantly down-regulated shift**; no GO-BP term is FDR-significantly *up*-regulated in the primary BEP contrasts:

- **MISAME-III (1–2 mo., BEP):** **27 FDR-significant down-regulated** GO-BP terms, dominated by proteostasis, translation, protein folding, and intracellular trafficking. The telomere-maintenance and DNA-templated terms that head the list are **driven by molecular chaperones**, the CCT/TCP1 chaperonin subunits together with HSP90AA1/AB1 and the co-chaperone PTGES3, appearing under those GO labels because chaperones annotate to telomere/DNA-repair complexes; this is a **chaperone/proteostasis** signal, **not** a telomere-specific effect.
- **Mumta-LW (2 mo., BEP+ExBf):** **58 FDR-significant down-regulated** terms centred on **energy metabolism**: glycolysis, pyruvate metabolism, cellular respiration, ATP generation, and nucleotide / NAD (pyridine-nucleotide) metabolism, driven by glycolytic and redox enzymes (PKM, GAPDH, ALDOA/ALDOC, PGAM1, GPI, LDHA, G6PD, IDH2, NAMPT).

> **Provenance of the 27/58 counts.** An earlier frozen spreadsheet (`Manuscript/Table_S6_proteomics_GO.xlsx`, approved run of 2026-08-10) reported **31** and **65** FDR-significant terms. That file is superseded: the underlying proteomic effect estimates were regenerated on 2026-08-12, growing the measured-proteome background from ~1,196/1,555 to **1,320/1,681 quantified proteins**, the counts used for the individual-protein analyses elsewhere in the manuscript. The manuscript's own Table S6 and Fig. 6C now report **27** and **58**, which is what the current pipeline produces and what the tables below show. There is no live discrepancy to reconcile; the 31/65 figures should not be cited.

> **Proteomic-interpretation caveat (GO BP).** Milk proteomic features draw from a heterogeneous mixture of cellular debris (mammary epithelial cells, leukocytes, exosomes) and secreted proteins. Enrichment of terms such as "telomere maintenance" or "glycolytic process" therefore indicates a signature **consistent with** changes in cellular composition or intracellular protein content in milk, not that the named process is being executed within secreting cells. Read these as hypothesis-generating; mechanistic claims require functional assays that were not part of the IMiC protocol.

The first table below **is** manuscript Table S6: the terms enriched at nominal *P* < 0.05 **and**
Storey *q* < 0.20, the inclusion threshold of the originally submitted table, giving 182 rows. Its `fdr`
column marks the *Q* < 0.05 frontier shown in Fig. 6C (27 MISAME-III + 58 Mumta-LW = 85 terms, all
down-regulated; no up-regulated term survives FDR correction). The print supplement externalises the same 182 rows as a machine-readable file; both it and the
table below come from `results/metaboanalyst/proteomics_go/proteomics_go_tableS6.csv`, so the
numbers are identical. **Download:** [`table_s6_proteomics_go_full.csv`](appendix-data/table_s6_proteomics_go_full.csv).

```{r path-go-tableS6-note, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
s6 <- read_enrich("results/metaboanalyst/proteomics_go/proteomics_go_tableS6.csv")
if (!is.null(s6) && nrow(s6)) {
  cat(sprintf("\n*Showing all %d terms at nominal P < 0.05 and Storey q < 0.20; manuscript Table S6. Of these, %d clear FDR Q < 0.05 (%d MISAME-III, %d Mumta-LW), all down-regulated.*\n\n",
              nrow(s6), sum(s6$fdr < 0.05, na.rm = TRUE),
              sum(s6$study == "MISAME-III" & s6$fdr < 0.05, na.rm = TRUE),
              sum(s6$study == "Mumta-LW"   & s6$fdr < 0.05, na.rm = TRUE)))
  ## Format only after the counts above are taken; formatC turns these numeric.
  for (cc in intersect(c("p_value", "fdr"), names(s6)))
    s6[[cc]] <- formatC(s6[[cc]], format = "e", digits = 2)
}
```

```{r path-go-tableS6, echo=FALSE, message=FALSE, warning=FALSE}
if (!is.null(s6) && nrow(s6))
  nice_dt(s6, caption = "Untargeted milk-proteome GO Biological Process enrichment: manuscript Table S6 (182 rows, nominal P < 0.05 and Storey q < 0.20). UniProt-native over-representation (clusterProfiler::enricher, per-cell measured-proteome background); regulation = direction of effect; fdr = BH-adjusted enrichment p-value.")
```

The second table is the **FDR-significant subset only** (the 85 rows with *Q* < 0.05), carrying the
fuller per-term detail (gene ratio, background ratio, rich factor, z-score, Storey *q*, and the
contributing UniProt accessions) that the printed table has no room for. Its 85 terms are exactly
the *Q* < 0.05 rows of Table S6 above; the two are the same run, one filtered further.

```{r path-go-tab-note, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
go <- read_enrich("results/metaboanalyst/proteomics_go/proteomics_go_uniprot_fdrsig.csv")
if (!is.null(go) && nrow(go)) {
  ## Scientific notation for the small enrichment p/q columns (nice_dt would
  ## otherwise round e.g. 1.2e-06 to 0.000).
  for (cc in intersect(c("pvalue", "p.adjust", "qvalue"), names(go)))
    go[[cc]] <- formatC(go[[cc]], format = "e", digits = 2)
  cat(sprintf("\n*Showing all %d FDR-significant (Q<0.05) GO-BP terms: %d MISAME-III (BEP, 1–2 mo.) and %d Mumta-LW (BEP+ExBf, 2 mo.), all down-regulated.*\n\n",
              nrow(go), sum(go$study == "MISAME-III", na.rm = TRUE),
              sum(go$study == "Mumta-LW", na.rm = TRUE)))
}
```

```{r path-go-tab, echo=FALSE, message=FALSE, warning=FALSE}
if (!is.null(go) && nrow(go)) {
  nice_dt(go, caption = "Untargeted milk-proteome GO Biological Process enrichment, FDR-significant subset of Table S6 with extended per-term detail (the Q < 0.05 frontier shown in Fig. 6C). UniProt-native over-representation (clusterProfiler::enricher, per-cell measured-proteome background); direction = sign of effect; p.adjust = BH-adjusted enrichment p-value; qvalue = Storey q.")
}
```
