# Reproducible pathway and enrichment analyses

This chapter reports the **pathway- and enrichment-level interpretation** of the per-component milk signals shown in §4 and §5, converting large tables of individual features into a smaller set of biologically interpretable pathways. It replaces the earlier manual `metaboanalyst.ca` point-and-click workflow with a fully scripted **MetaboAnalystR** pipeline (`src/metaboanalyst/` in the upstream analysis repository), so that every table below can be regenerated deterministically from the intervention-effect estimates.

Each analysis is reported twice: for the **combined-arm** contrast (all maternal nutritional-intervention arms pooled vs. control — the framing used in the printed manuscript figures) and for the **arm-stratified** contrasts (each randomised arm vs. control — the supplement view). Enrichment tests use a permissive foreground (features at unadjusted *p* < 0.05, split by direction of effect) evaluated against the assay-specific background, then Benjamini–Hochberg FDR correction of the pathway *p*-values within each study × timepoint × contrast × direction stratum (and, for Mummichog, ionisation mode). A pathway is flagged **significant** at *Q* < 0.05 (`fdr_native` / `fdr` / `FDR` column, depending on the table).

> **Interpretation note.** These enrichment results summarise and contextualise the prespecified per-component analyses; they are not themselves a separate confirmatory test. Over-representation depends on the choice of background metabolite set (see §9.2), and pathway membership in reference databases is incomplete, so absence of a pathway is not evidence of absence of biology. Read the results as a structured, hypothesis-generating map of where the intervention signal concentrates.

> **Coverage — all six manuscript enrichment tables (S1–S6) are reproduced here** as searchable tables, in manuscript order: primary ORA (**Table S1**, §9.1), the ORA-vs-impact comparison (§9.2), the empty secondary result (§9.3), tertiary MSEA (**Table S2**, §9.4), triglyceride fatty-acid composition (**Table S3**, §9.5), untargeted-metabolome MSEA (**Table S4**, §9.6), milk untargeted Mummichog (**Table S5**, §9.7), and milk-proteome GO (**Table S6**, §9.8).

```{r path-setup, include=FALSE}
library(tidyverse)
library(here)
library(DT)
library(plotly)
source(here::here("functions.R"))
ggplot2::theme_set(ggplot2::theme_minimal(base_size = 14))
knitr::opts_chunk$set(fig.width = 12, fig.height = 8, dpi = 110)

## Read an enrichment CSV, degrading gracefully. Unlike safe_read_csv this also
## survives an empty (0-byte / header-only) file — some cells of the enrichment
## grid legitimately return no pathways (e.g. HMOs/bioactives, §9.3).
read_enrich <- function(path) {
  full <- here::here(path)
  if (!file.exists(full)) {
    cat(sprintf("*Results file not found at `%s`. Run `port_results.R` to copy it from the upstream analysis repo.*", path))
    return(invisible(NULL))
  }
  df <- tryCatch(utils::read.csv(full, check.names = FALSE), error = function(e) NULL)
  if (is.null(df) || ncol(df) == 0) return(data.frame())
  df
}

## Present p-value / FDR columns in scientific notation so very small enrichment
## p-values (e.g. 3.7e-11) are not rounded to 0.000 by nice_dt().
p_cols <- c("raw_p", "fdr_native", "fdr_bh_pooled", "fdr", "impact",
            "p_value", "P-value", "FDR", "Enrichment Ratio", "enrichment_ratio",
            "fold_enrichment")
enrich_dt <- function(df, caption = NULL, page_length = 15) {
  if (is.null(df)) return(invisible(NULL))
  if (!nrow(df))  return(htmltools::tags$em("No enriched pathways in this stratum."))
  sci <- intersect(c("raw_p","fdr_native","fdr_bh_pooled","fdr","p_value","P-value","FDR"), names(df))
  for (cc in sci) df[[cc]] <- formatC(df[[cc]], format = "e", digits = 2)
  nice_dt(df, caption = caption, page_length = page_length)
}
```

## Interactive pathway explorer {#pathway-explorer-app}

The app below spans all six enrichment analyses in this chapter (manuscript Tables S1–S6). Pick an analysis, restrict to FDR-significant pathways or show everything, and filter by study and timepoint; the dot plot ranks pathways by −log10(FDR) and the table gives the full numbers. Like the effect explorer in §5, it runs entirely in your browser (WebAssembly R), so the first load fetches the runtime and is then instant. [Open it full-screen in a new tab](apps/pathway-explorer/).

```{r pathway-explorer-embed, echo=FALSE, results='asis'}
if (knitr::is_html_output()) {
  cat('<iframe src="apps/pathway-explorer/" title="IMiC Pathway-Enrichment Explorer" ',
      'width="100%" height="820" loading="lazy" ',
      'style="border:1px solid #ddd; border-radius:4px;"></iframe>\n', sep = "")
} else {
  cat("*The interactive pathway explorer is available in the HTML build of this appendix.*\n")
}
```

## Primary outcomes — MetaboAnalyst enrichment of micronutrient / B-vitamin features {#path-primary}

These are the pathway summaries of the **primary-outcome** signals (macronutrients, micronutrients, B-vitamins). They correspond to manuscript **Table S1** and the enrichment panels of manuscript **Fig. 3B**. Over-representation analysis (ORA) is run per study × timepoint × contrast × direction against the targeted-panel background; the `significant` column flags pathways with `fdr_native` < 0.05.

The clearest result is **nicotinate and nicotinamide metabolism**, driven by the ELICIT nicotinamide-during-lactation arm (up-regulated at both the 1-month and 5-month visits; 5/15 pathway metabolites hit, hypergeometric *p* ≈ 3.7 × 10⁻¹¹, *Q* ≈ 3.0 × 10⁻⁹), and also FDR-significant in the BEP trials (Mumta-LW, MISAME-III). Riboflavin, vitamin B~6~, and thiamine metabolism appear as secondary FDR-significant pathways across BEP trials and timepoints.

```{r path-primary-fig-combined, echo=FALSE, message=FALSE, warning=FALSE}
safe_img("figures/figure4_combined_arms.png", label = "figure4_combined_arms.png")
```

***Figure (manuscript Fig. 3B, combined arms).** Volcano and pathway-enrichment composite for the primary micronutrient / B-vitamin outcomes, all maternal-intervention arms pooled vs. control. Points are individual components; enriched pathways are annotated. Nicotinate/nicotinamide metabolism (ELICIT) and riboflavin / vitamin B~6~ metabolism (BEP trials) are the leading FDR-significant pathways.*

```{r path-primary-tab-combined, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/primary_combined/primary_combined_supplementary_table.csv"),
          caption = "Primary-outcome MetaboAnalyst ORA, combined intervention arms vs. control (manuscript Table S1). significant = fdr_native < 0.05.")
```

### Arm-stratified

```{r path-primary-fig-strat, echo=FALSE, message=FALSE, warning=FALSE}
safe_img("figures/figure4_stratified_supplement.png", label = "figure4_stratified_supplement.png")
```

***Figure (manuscript Fig. 3B, arm-stratified supplement).** As above, with each randomised arm contrasted separately against its trial control.*

```{r path-primary-tab-strat, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/primary_stratified/primary_stratified_supplementary_table.csv"),
          caption = "Primary-outcome MetaboAnalyst ORA, arm-stratified contrasts vs. control.")
```

### Direction-split ORA (up- vs down-regulated)

These standalone tables run the over-representation separately on the up- and down-regulated primary feature lists (combined arms): the §9.1 ORA already splits the *foreground* by direction internally when it builds the enrichment, and here that split is reported directly, recovering **nicotinate and nicotinamide metabolism** among the up-regulated features (ELICIT) and **lactose synthesis and superoxide degradation** among the down-regulated features (MISAME-III).

```{r path-primary-dirsplit-fig, echo=FALSE, message=FALSE, warning=FALSE}
safe_img("figures/figureS_primary_ora_by_direction.png", label = "figureS_primary_ora_by_direction.png")
```

***Figure. Primary-outcome ORA split by direction of effect (combined arms).** Enriched pathways among up- vs. down-regulated primary features. Nicotinate/nicotinamide metabolism (ELICIT) dominates the up-regulated side; lactose synthesis and superoxide degradation (MISAME-III) appear among the down-regulated side.*

```{r path-primary-ora-up, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/primary_combined/primary_ora_upregulated.csv"),
          caption = "Primary-outcome ORA, up-regulated features (combined arms).")
```

```{r path-primary-ora-down, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/primary_combined/primary_ora_downregulated.csv"),
          caption = "Primary-outcome ORA, down-regulated features (combined arms).")
```

## Primary outcomes — over-representation vs. quantitative pathway/impact analysis {#path-ora-vs-pathway}

MetaboAnalyst supports two related enrichment styles that answer subtly different questions, and comparing them makes the **background dependence** of enrichment explicit:

- **Over-representation analysis (ORA)** — the §9.1 tables — tests whether the significant features are unexpectedly concentrated in a pathway, using the **restricted targeted-panel** as the background universe.
- **Pathway / impact analysis** — the table below — additionally weights each pathway by the topological position of the hit metabolites and evaluates it against the **full SMPDB** background.

Both recover **nicotinate and nicotinamide metabolism** as the leading primary-outcome pathway. The impact analysis additionally flags **vitamin B~6~ metabolism**; because that difference tracks the larger background set rather than a distinct data signal, we show the two side by side so the dependence stays visible.

```{r path-ora-pathway-tab, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/primary_combined_pathway/primary_combined_pathway_all_cells.csv"),
          caption = "Primary-outcome quantitative pathway/impact analysis (full SMPDB background), combined arms. impact = pathway topology score; fdr = BH-adjusted pathway p-value.")
```

## Secondary outcomes — HMOs and bioactive proteins {#path-secondary}

Enrichment was also attempted for the secondary outcomes (human milk oligosaccharides and bioactive proteins). These returned **no enriched metabolite-set pathways**: HMOs and milk bioactive proteins are largely absent from the SMPDB metabolite-set libraries, so the mapping step leaves too few annotated features to test. We report the empty result here for completeness; it is not a null biological finding.

```{r path-secondary-tab, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/secondary_combined/secondary_combined_all_cells.csv"),
          caption = "Secondary-outcome enrichment, combined arms (empty — HMOs/bioactives are not represented in SMPDB metabolite sets).")
enrich_dt(read_enrich("results/metaboanalyst/secondary_stratified/secondary_stratified_all_cells.csv"),
          caption = "Secondary-outcome enrichment, arm-stratified contrasts (also empty, for the same reason).")
```

## Tertiary outcomes — targeted-metabolome MSEA {#path-tertiary}

Metabolite-set enrichment analysis (MSEA) of the tertiary **Biocrates MxP Quant 500** targeted panel summarises the per-metabolite effects from §5 at the pathway level. It corresponds to manuscript **Table S2** and **Fig. 5B**.

Under BEP supplementation, breastmilk metabolism undergoes **time-dependent reprogramming**. In the early postpartum period the enrichment is dominated by broad **down-regulation of amino-acid and nitrogen-handling pathways** (arginine–proline, glycine–serine, methionine, and glutamate metabolism, and ammonia recycling) together with **mitochondrial energy metabolism** (fatty-acid β-oxidation, the citric-acid cycle, the electron-transport chain, ketone-body metabolism, and carnitine synthesis). By 3–4 months several of these pathways reverse, showing enrichment of mitochondrial respiration, lipid oxidation, and ketone metabolism. (The distinct down-regulation of *de novo fatty-acid biosynthesis* is an **untargeted**-metabolome finding — see §9.6 — not a targeted-panel MSEA result.)

### MSEA — combined and arm-stratified

```{r path-tertiary-msea-combined, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/tertiary_msea/tertiary_msea_combined.csv"),
          caption = "Tertiary targeted-metabolome MSEA, combined intervention arms vs. control (manuscript Table S2 / Fig. 5B). fdr_native = BH-adjusted enrichment p-value.")
```

```{r path-tertiary-msea-strat, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/tertiary_msea/tertiary_msea_stratified.csv"),
          caption = "Tertiary targeted-metabolome MSEA, arm-stratified contrasts vs. control.")
```

### ORA counterpart

For readers who prefer the over-representation view of the same tertiary features (consistent column structure with §9.1):

```{r path-tertiary-ora-combined, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/tertiary_combined/tertiary_combined_supplementary_table.csv"),
          caption = "Tertiary targeted-metabolome over-representation analysis, combined arms.")
```

```{r path-tertiary-ora-strat, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/tertiary_stratified/tertiary_stratified_supplementary_table.csv"),
          caption = "Tertiary targeted-metabolome over-representation analysis, arm-stratified contrasts.")
```

### Direction-split ORA (up- vs down-regulated)

As for the primary outcomes, these standalone tables run the over-representation separately on the up- and down-regulated tertiary (targeted-metabolite) feature lists (combined arms) — the per-direction companion to the MSEA and ORA views above, with the composite figure alongside.

```{r path-tertiary-dirsplit-fig, echo=FALSE, message=FALSE, warning=FALSE}
safe_img("figures/figureS_tertiary_ora_by_direction.png", label = "figureS_tertiary_ora_by_direction.png")
```

***Figure. Tertiary targeted-metabolome ORA split by direction of effect (combined arms).** Enriched pathways among up- vs. down-regulated tertiary features.*

```{r path-tertiary-ora-up, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/tertiary_combined/tertiary_ora_upregulated.csv"),
          caption = "Tertiary-outcome ORA, up-regulated features (combined arms).")
```

```{r path-tertiary-ora-down, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/tertiary_combined/tertiary_ora_downregulated.csv"),
          caption = "Tertiary-outcome ORA, down-regulated features (combined arms).")
```

## Triglyceride fatty-acid composition {#path-tg-fa}

The compositional companion to the fat-adjusted triglyceride analysis in §8.3, corresponding to manuscript **Table S3** and **Fig. 5C**. For each study × timepoint × contrast × direction stratum, the fatty-acyl moieties carried by the intervention-responsive triglycerides are compared against the background TG pool (a Wilcoxon test on within-TG proportional representation across the sn-1/sn-2/sn-3 positions). This identifies fatty acids **compositionally enriched or depleted** among the BEP-responsive triglycerides — *not* absolute fatty-acid concentrations. The consistent signal is enrichment of **polyunsaturated fatty acids** — α-linolenic (18:3n-3), γ-linolenic, and related PUFAs — among the triglycerides increased under BEP across MISAME-III contrasts and timepoints. The signal resolves in the **arm-stratified** analysis (the combined-arm table is empty).

```{r path-tg-fa-tab, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/triglyceride_fa/triglyceride_fa_composition_stratified.csv"),
          caption = "Compositional fatty-acid enrichment of intervention-responsive triglycerides, arm-stratified contrasts (manuscript Table S3 / Fig. 5C). fdr_native = BH-adjusted p; positive enrichment_ratio = fatty acid over-represented among up-regulated TGs.")
```

## Untargeted-metabolome MSEA {#path-untargeted-msea}

Metabolite-set enrichment analysis (MSEA) of the **untargeted** milk-metabolome feature lists, corresponding to manuscript **Table S4** and **Fig. 6A**. This is the named-metabolite-set counterpart to the Mummichog view in §9.7 — both operate on the same untargeted features, but MSEA tests curated metabolite sets whereas Mummichog infers pathways directly from *m/z*. The signature is a broad **down-regulation** under BEP, spanning polyamine biosynthesis (spermidine / spermine), amino-acid metabolism (glycine–serine), branched-chain and de novo fatty-acid pathways, and related lipid metabolism, concentrated in MISAME-III. The signal is carried by the **combined-arm** contrast (the arm-stratified table is empty).

```{r path-untargeted-msea-tab, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/untargeted_msea/untargeted_msea_combined.csv"),
          caption = "Untargeted-metabolome MSEA, combined intervention arms vs. control (manuscript Table S4 / Fig. 6A). fdr_native = BH-adjusted enrichment p-value; negative enrichment_ratio = enrichment among down-regulated features.")
```

## Exploratory untargeted milk metabolome — Mummichog functional enrichment {#path-mummichog}

Mummichog infers enriched pathways directly from mass-spectral features (without prior compound identification), matching *m/z* to putative compounds within each ionisation mode. Annotation uses a **10 ppm** database tolerance against the `human_mfn` reference network, with pathways tested at *p* < 0.05. This is the exploratory **untargeted-metabolome** enrichment, corresponding to manuscript **Table S5** and **Fig. 6B**, reported by study × timepoint × contrast × direction × ionisation mode.

> **Two tolerances, one number apart.** The **10 ppm** here is a *compound-annotation* tolerance (matching a feature's *m/z* to a database compound). It is a different quantity from the **±25 ppm** *cross-platform feature-matching* tolerance used in §10.2 (matching a feature in one dataset to a feature in another). A reviewer comparing the two chapters should not read the difference as an inconsistency.

The dominant signature is **down-regulation of polyunsaturated-fatty-acid and lipid pathways** — de novo fatty-acid biosynthesis, fatty-acid activation and metabolism, arachidonic-acid and linoleate metabolism, and leukotriene / prostaglandin formation — concentrated in the MISAME-III BEP contrasts. The full grid is large (≈9,500 pathway × stratum rows), so the table below is filtered to the FDR-significant rows.

```{r path-mummichog-tab, echo=FALSE, message=FALSE, warning=FALSE}
mum <- read_enrich("results/metaboanalyst/mummichog_s5/milk_mummichog_tableS5.csv")
if (!is.null(mum) && nrow(mum)) {
  mum_sig <- mum %>% dplyr::filter(.data[["FDR"]] < 0.05) %>% dplyr::arrange(.data[["FDR"]])
  cat(sprintf("\n*Showing %s FDR-significant (Q<0.05) pathway strata of %s total.*\n\n",
              format(nrow(mum_sig), big.mark = ","), format(nrow(mum), big.mark = ",")))
  print(enrich_dt(mum_sig,
        caption = "Milk untargeted-metabolome Mummichog enrichment, FDR-significant strata (manuscript Table S5 / Fig. 6B)."))
}
```

### Feature-level putative annotation {#path-mummichog-annotation}

The pathway view above is complemented by the finalized per-feature annotation of the same untargeted milk features. Across **1,034 FDR-significant** features, organised by direction × ionisation mode, **243 carry a name** (48 curated + 195 Mummichog-putative) and the remainder are unresolved; the strongest named up-regulated signal is a vitamin/cofactor cluster (pantothenate/B5, nicotinamide, ascorbate). Every name is **putative** — mass-based, MSI ~level 3 — not a confirmed identification.

```{r path-mummichog-annot-summary, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/mummichog_s5/milk_mummichog_annotation_summary.csv"),
          caption = "FDR-significant milk features by direction and ionization mode: named vs unresolved.")
```

```{r path-mummichog-annot-full, echo=FALSE, message=FALSE, warning=FALSE}
annot <- read_enrich("results/metaboanalyst/mummichog_s5/milk_mummichog_annotation_finalized.csv")
if (!is.null(annot) && nrow(annot)) {
  DT::datatable(annot, filter = "top", rownames = FALSE,
                options = list(pageLength = 10, scrollX = TRUE),
                caption = "Finalized putative annotation of FDR-significant untargeted milk features.")
}
```

## Exploratory untargeted milk proteome — GO functional enrichment {#path-go}

Gene Ontology (Biological Process) over-representation of the intervention-associated **untargeted milk proteomic** features (UniProt → Entrez → `clusterProfiler::enrichGO`, against the measured-proteome background) corresponds to manuscript **Table S6** and **Fig. 6C**. The signal spans three domains: down-regulated cellular/housekeeping processes (chromatin organisation, telomere maintenance, translation), up-regulated negative regulation of endopeptidase activity (SERPINs, TIMPs), and humoral/complement immune terms.

> **Proteomic-interpretation caveat (GO BP).** Milk proteomic features draw from a heterogeneous mixture of cellular debris (mammary epithelial cells, leukocytes, exosomes) and secreted proteins. Enrichment of terms such as "telomere maintenance" or "negative regulation of peptidase activity" therefore indicates a signature **consistent with** changes in cellular composition or proteolytic balance in milk, not that the named process is being executed within secreting cells. Read these as hypothesis-generating; mechanistic claims require functional assays that were not part of the IMiC protocol.

```{r path-go-tab, echo=FALSE, message=FALSE, warning=FALSE}
enrich_dt(read_enrich("results/metaboanalyst/proteomics_go/proteomics_go_tableS6.csv"),
          caption = "Untargeted milk-proteome GO Biological Process enrichment (manuscript Table S6 / Fig. 6C). fdr = BH-adjusted enrichment p-value.")
```
