---
title: "IMiC supplementary material"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
description: |
  Online supplementary appendix for "Nutritional interventions' impacts on human milk: three trial analyses from low-resource settings" (Dailey-Chwalibóg, Mertens et al., 2025), reporting the full analytical outputs of the International Milk Composition (IMiC) Consortium intervention-effects analysis.
link-citations: yes
github-repo: amertens/IMiC-intervention-effects-supplement
url: https://amertens.github.io/IMiC-intervention-effects-supplement/
---

# Overview {-}

This online appendix is the **complete, searchable companion** to the manuscript:

> **Nutritional interventions' impacts on human milk: three trial analyses from low-resource settings** (Dailey-Chwalibóg, Mertens et al., 2025). International Milk Composition (IMiC) Consortium.

It provides every analytical output that supports the main manuscript, in more detail than the printed article PDF can hold: interactive forest and volcano plots, searchable and filterable estimate tables, longitudinal trajectory visualisations, MILQ-anchored deficiency analyses, BMI-stratified subgroup analyses, reproducible pathway-enrichment results, and the cross-compartment maternal/infant blood analyses.

The appendix is organised into **ten numbered sections**, preceded by this unnumbered overview and followed by a reproducibility appendix. The paper cites Sections 1–8 by name in its Methods and Results; **§9 (reproducible pathway and enrichment analyses)** and **§10 (cross-compartment analyses in maternal and infant blood)** were added in this revision and are cited without a section number, as the "Online Appendix" behind the manuscript's pathway tables (Tables S1–S6) and its cross-compartment tables (Tables S8–S11, main-text Fig. 6D).

**Figure and table numbering.** Cross-references throughout this appendix follow the **submitted supplement** (Figs. S1–S11, Tables S1–S11), so a number cited in the submitted PDF resolves to the same exhibit here. Two of the submitted figures have no counterpart in this appendix: **Fig. S1** (study design and sampling timeline) and **Fig. S7** (the cross-compartment transfer composite), neither of which is regenerated from the analysis pipeline.

| Submitted supplement | Subject | Reported here in |
|---|---|---|
| **Fig. S1** | Study design, component classes, sampling timeline | not reproduced online |
| **Fig. S2** | Infant growth (LAZ / WLZ / WAZ / HCAZ / MUACZ) | §8.7 |
| **Fig. S3** | MILQ-referenced micronutrient deficiency (relative risks) | §6 |
| **Fig. S4** | HMOs and bioactive proteins (secondary outcomes) | §5 |
| **Fig. S5** | Milk triglycerides, with and without fat adjustment | §8.3 |
| **Fig. S6** | Milk-microbiome α-diversity | §4 |
| **Fig. S7** | Cross-compartment transfer of BEP-responsive metabolites | not reproduced online |
| **Fig. S8** | Chemical-class enrichment of blood metabolite effects | §10.6 |
| **Fig. S9** | Direction-split ORA, (A) primary and (B) tertiary outcomes | §9.1 (A), §9.4 (B) |
| **Fig. S10** | Cross-compartment feature tracking and supplement origin | §10.5 |
| **Fig. S11** | Cross-compartment volcanoes across compartments and visits | §10.5 |

Figures in this appendix that carry no submitted-supplement number are labelled *online-only* where they appear. Main-text figures are cited as **Fig. 1–6** and match the article: machine-learning / PCA composites are **Fig. 1** (§8.4–§8.5), the primary-outcome forest plots **Fig. 2** (§5), the primary-outcome volcano and pathway panels **Fig. 3A–B** (§5, §9.1), the MILQ-referenced distributions **Fig. 4** (§2), the targeted-metabolite panels **Fig. 5A–C** (§5, §9.4–§9.5), and the untargeted metabolome/proteome panels **Fig. 6A–D** (§9.6–§9.8, §10.5).

## How this appendix maps to the manuscript

| Manuscript reference | Section here |
|---|---|
| Methods: "baseline participant characteristics … (Section 1)" | [§1 Baseline characteristics by study and intervention arm](#baseline-characteristics-by-study-and-intervention-arm) |
| Methods: "outcomes list … (Section 2)"; manuscript **Fig. 4** (MILQ-referenced distributions) | [§2 Outcomes and human milk component distributions](#outcomes-and-human-milk-component-distributions) |
| Results: "Online Appendix 3" (effect modification by maternal BMI) | [§3 Subgroup analyses stratified by maternal BMI](#subgroup-analyses-stratified-by-maternal-bmi) |
| Results: "Online Appendix 4" (individual microbial taxa); Fig. S6 (α-diversity) | [§4 Microbiome and exploratory high-dimensional outcomes](#exploratory-high-dimensional-outcomes-microbiome-untargeted-proteomics-untargeted-metabolomics) |
| manuscript Fig. 2 (primary-outcome forest plots), Fig. S4 (secondary outcomes), the per-component estimates underlying Fig. 5A (tertiary metabolites), and Fig. 3A's "metadata in Online Appendix 5" pointer (§5.1) | [§5 Adjusted intervention effects on individual milk components](#adjusted-intervention-effects-on-individual-milk-components) |
| Fig. 5A: "metadata in Online Appendix 6" (targeted metabolites); manuscript Fig. S3 (MILQ deficiency) | [§6 Targeted-metabolite metadata and reductions in HM nutrient deficiency](#targeted-metabolite-metadata-and-reductions-in-hm-nutrient-deficiency) |
| Longitudinal change in HM composition | [§7 Trajectory analyses of HM component change across lactation](#trajectory-analyses-of-human-milk-component-change-across-lactation) |
| manuscript Fig. 1 (treatment-arm prediction & PCA composites); Fig. S2 (growth); Fig. S5 (TG fat adjustment); **Table S7** (native-unit concentrations); sensitivity analyses | [§8 Sensitivity analyses and additional outputs](#sensitivity-analyses-and-additional-outputs) |
| manuscript Fig. 3B (primary MSEA), Fig. 5B–C and Fig. 6A–C (tertiary/untargeted enrichment); **Tables S1–S6** (pathway enrichment); direction-split ORA panels (**Fig. S9A–B**) | [§9 Reproducible pathway and enrichment analyses](#reproducible-pathway-and-enrichment-analyses) |
| Results: "intervention effects are mirrored in maternal and infant circulation"; **Table S8** (cross-compartment features), **Table S9** (chemical-class enrichment), **Table S10** (temporal persistence), and **Table S11**/main-text **Fig. 6D** (cross-compartment feature tracking and pathway enrichment). No printed-supplement figure; all figures here are online-only | [§10 Cross-compartment analyses in maternal and infant blood](#cross-compartment-analyses-in-maternal-and-infant-blood) |

The manuscript's section-numbered citations resolve as follows: *"Section 1"* → §1; *"Section 2"* (outcomes list) → §2; *"Online Appendix 3"* → §3; *"Online Appendix 4"* → §4; Fig. 3A's *"Online Appendix 5"* (primary/secondary component metadata) → §5.1; and Fig. 5A's *"Online Appendix 6"* (targeted-metabolite metadata) → §6.1. The abstract-level pointer to *"Sections 1–8"* covers the outputs carried over from the original submission; §9 and §10, added in this revision, are the "Online Appendix" cited without a number elsewhere in the Results.

## Methods recap

We conducted a harmonised individual-participant data meta-analysis of human milk from 640 mother–infant dyads drawn from three randomised controlled trials in one low- and two lower-middle-income country settings: **MISAME-III** (Burkina Faso; prenatal and/or postnatal balanced energy–protein supplementation), **Mumta-LW** (Pakistan; postnatal BEP and BEP+infant azithromycin), and **ELICIT** (Tanzania; maternal nicotinamide during lactation with independent infant antimicrobial arms). Each trial contributed its own intent-to-treat population, restricted to dyads with at least one HM measurement contributing to IMiC.

Primary outcomes are macronutrients and micronutrients; secondary are HMOs and bioactive proteins; tertiary are targeted metabolites (Biocrates MxP Quant 500); exploratory are untargeted metabolomics, untargeted proteomics, and microbiome composition. All components were expressed in native assay units, then standardised to study × visit z-scores before TMLE. Effect estimates from the primary chapter therefore represent **mean differences in SD units**, with FDR correction applied within each outcome group × study × visit. The complete methods are in the manuscript and its Methods supplement.

Two different mass tolerances appear in the enrichment and cross-compartment chapters and are **not** interchangeable: pathway/compound annotation in §9 (Mummichog) uses a **10 ppm** database tolerance against the `human_mfn` library, whereas cross-platform feature matching in §10 uses a **±25 ppm** experimental-over-experimental tolerance (Kim's formula, `ppm = |m₁ − m₂| / m × 10⁶ ≤ 25`). The 10 ppm annotation tolerance and the ±25 ppm feature-matching tolerance are distinct quantities.

The full outcomes list (every milk component analysed in this supplement, grouped by prespecified outcome class with per-study availability) now lives in [§2 *Outcomes and human milk component distributions*](#outcomes-and-human-milk-component-distributions). The manuscript Methods reference *"A list of all outcomes and descriptions is available in the Online Appendix (Section 2)"* therefore resolves to a numbered section rather than an unnumbered overview.

## How to read interactive figures

In the HTML build, forest and volcano plots are rendered as `plotly` widgets. Hover any point to reveal the component name, study, visit, ATE, 95% CI, and FDR-adjusted q-value, and use the mode bar at the top right of each plot to zoom, pan, lasso-select, or export a PNG. Every plot is paired with a `DT::datatable` carrying the same underlying data, with column filters, free-text search, and sortable columns.

Where a section relies on an upstream data artefact that is not present in the current build, the section prints an inline notice naming the missing file rather than failing. Re-run the upstream analysis pipeline (see *Reproducing the supplement* in the repository README) to regenerate those artefacts.

## Machine-readable data files {#data-files}

Four supplementary tables were too large to print in full, so the printed supplement gives a short caption and points here for the complete data. Each is a plain CSV, downloadable directly:

```{r appendix-data-files, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
## Publish appendix-data/ alongside the book so the filenames the printed
## supplement names are actually downloadable from the rendered site.
src <- here::here("appendix-data")
dst <- here::here("docs", "appendix-data")
if (dir.exists(src)) {
  dir.create(dst, recursive = TRUE, showWarnings = FALSE)
  file.copy(list.files(src, full.names = TRUE), dst, overwrite = TRUE)
}
files <- list(
  c("table_s2_tertiary_msea_full.csv",
    "Table S2: tertiary targeted-metabolome MSEA, every pathway x study x visit x contrast x direction row (see §9.4)"),
  c("table_s5_mummichog_full.csv",
    "Table S5: milk untargeted-metabolome Mummichog enrichment, pathways at raw P < 0.05 (see §9.7)"),
  c("table_s6_proteomics_go_full.csv",
    "Table S6: milk-proteome GO Biological Process enrichment, terms at nominal P < 0.05 and Storey q < 0.20 (see §9.8)"),
  c("table_s7_primary_secondary_native_units.csv",
    "Table S7: native-unit concentrations and intervention effects for all primary and secondary outcomes, long format (see §8.4)"),
  c("table_s7_primary_secondary_native_units_wide.csv",
    "Table S7: the same data in wide format, one row per component x study x visit with the arms side by side")
)
for (f in files) {
  p <- file.path(src, f[1])
  n <- if (file.exists(p)) max(0, length(readLines(p, warn = FALSE)) - 1L) else NA_integer_
  cat(sprintf("- [`%s`](appendix-data/%s): %s%s\n", f[1], f[1], f[2],
              if (is.na(n)) " *(not present in this build)*" else ""))
}
```

Every one of these is also rendered as a searchable table in the section named beside it; the CSVs are for readers who want the numbers in a spreadsheet or script.

## Citation and reproducibility

- Source for this appendix: <https://github.com/amertens/IMiC-intervention-effects-supplement>
- Upstream analysis code (TMLE pipelines, SuperLearner, enrichment workflows): <https://github.com/amertens/imic_intervention_effects>
- Trial registration: ClinicalTrials.gov **NCT05119166**
- DOI (Zenodo archive): *pending*

```{r write-bib, include=FALSE}
knitr::write_bib(c(.packages(), "bookdown","knitr","rmarkdown","tidyverse","here",
                   "DT","plotly","table1"),
                 "packages.bib")
```
