---
title: "IMiC intervention effects on human milk: online resource"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
description: |
  Online resource for "Nutritional interventions' impacts on human milk: three trials in low-resource settings" (Dailey-Chwalibóg, Mertens et al., 2026), reporting the full analytical outputs of the International Milk Composition (IMiC) Consortium intervention-effects analysis.
link-citations: yes
github-repo: amertens/IMiC-intervention-effects-supplement
url: https://amertens.github.io/IMiC-intervention-effects-supplement/
---

# Overview {-}

This online resource accompanies the manuscript:

> **Nutritional interventions' impacts on human milk: three trials in low-resource settings** (Dailey-Chwalibóg, Mertens et al., 2026). International Milk Composition (IMiC) Consortium.

It reports the analytical outputs behind the manuscript in more detail than the printed article can hold: interactive forest and volcano plots, searchable and filterable estimate tables, longitudinal trajectory plots, MILQ-referenced deficiency analyses, BMI-stratified subgroup analyses, reproducible pathway-enrichment results, and the cross-compartment analyses in maternal and infant blood.

The online resource has ten numbered sections, preceded by this unnumbered overview and followed by a page recording the R session that produced it. The paper cites Sections 1–6 by number in its Methods, Results, and figure legends, and cites the rest of the online resource without a section number. §9 (reproducible pathway and enrichment analyses) holds the full results behind the manuscript's pathway tables (Tables S2–S7), and §10 (cross-compartment analyses in maternal and infant blood) those behind its cross-compartment tables (Tables S8–S11) and main-text Fig. 6D; the Results cite both as "the online resource".

Table and figure numbers throughout this online resource match the manuscript's supplementary materials (Tables S1–S11 and Figs. S1–S5). The supplementary figures appear here as follows:

| Supplementary figure | Subject | Reported here in |
|---|---|---|
| **Fig. S1** | Infant growth (LAZ and WLZ) | §8.7 |
| **Fig. S2** | MILQ-referenced micronutrient deficiency (relative risks) | §6 |
| **Fig. S3** | HMOs and bioactive proteins (secondary outcomes) | §5 |
| **Fig. S4** | Milk triglycerides, with and without fat adjustment | §8.3 |
| **Fig. S5** | Milk-microbiome α-diversity | §4 |

Figures with no supplementary-figure counterpart (the direction-split pathway ORA panels in §9 and every figure in §10) are labeled *online-only*. Main-text figures are cited as Fig. 1–6, numbered as in the article: the machine-learning and PCA composites are Fig. 1 (§8.4–§8.5), the primary-outcome forest plots Fig. 2 (§5), the primary-outcome volcano and pathway panels Fig. 3A–B (§5, §\@ref(path-primary)), the MILQ-referenced distributions Fig. 4 (§2), the targeted-metabolite panels Fig. 5A–C (§5, §\@ref(path-tertiary)–§\@ref(path-tg-fa)), and the untargeted metabolome and proteome panels Fig. 6A–D (§\@ref(path-untargeted-msea)–§\@ref(path-go), §\@ref(cc-tracking)).

## How this online resource maps to the manuscript

| Manuscript reference | Section here |
|---|---|
| Methods: "Baseline participant characteristics … are available in Section 1 of the online resource" | [§1 Baseline characteristics by study and intervention arm](#baseline-characteristics-by-study-and-intervention-arm) |
| Methods: "A list of all outcomes and descriptions is available in Section 2 of the online resource"; manuscript **Fig. 4** (MILQ-referenced distributions) | [§2 Outcomes and human milk component distributions](#outcomes-and-human-milk-component-distributions) |
| Results: effect modification by maternal BMI, "as shown in Section 3 of the online resource" | [§3 Subgroup analyses stratified by maternal BMI](#subgroup-analyses-stratified-by-maternal-bmi) |
| Results: individual microbial taxa, "as reported in Section 4 of the online resource"; Fig. S5 (α-diversity) | [§4 Microbiome and exploratory high-dimensional outcomes](#exploratory-high-dimensional-outcomes-microbiome-untargeted-proteomics-untargeted-metabolomics) |
| manuscript Fig. 2 (primary-outcome forest plots), Fig. S3 (secondary outcomes), the per-component estimates underlying Fig. 5A (tertiary metabolites), and the Fig. 3A legend's "metadata in Section 5 of the online resource" (§5.1) | [§5 Adjusted intervention effects on individual milk components](#adjusted-intervention-effects-on-individual-milk-components) |
| Fig. 5A legend: "metadata are in Section 6 of the online resource" (targeted metabolites, §6.1); manuscript Fig. S2 (MILQ deficiency) | [§6 Targeted-metabolite metadata and reductions in HM nutrient deficiency](#targeted-metabolite-metadata-and-reductions-in-hm-nutrient-deficiency) |
| Longitudinal change in HM composition | [§7 Trajectory analyses of HM component change across lactation](#trajectory-analyses-of-human-milk-component-change-across-lactation) |
| manuscript Fig. 1 (treatment-arm prediction & PCA composites); Fig. S1 (growth); Fig. S4 (TG fat adjustment); **Table S1** (native-unit concentrations, §\@ref(native-units-table)) and the native-unit estimates for every targeted metabolite (§\@ref(native-units-tertiary)); sensitivity analyses | [§8 Sensitivity analyses and additional outputs](#sensitivity-analyses-and-additional-outputs) |
| manuscript Fig. 3B (primary pathway analysis), Fig. 5B–C and Fig. 6A–C (tertiary/untargeted enrichment); **Tables S2–S7** (pathway enrichment); Results: "up- and down-regulated pathways are reported separately in the online resource" (§\@ref(path-tertiary)) and "Amino-acid and central-carbon pathways recurred across the targeted and untargeted metabolome … as detailed in the online resource" (§\@ref(path-tertiary), §\@ref(path-untargeted-msea)); direction-split ORA panels (online-only) | [§9 Reproducible pathway and enrichment analyses](#reproducible-pathway-and-enrichment-analyses) |
| Results subsection "BEP-associated changes in maternal blood, milk, and infant blood (MISAME-III)"; **Table S8** (cross-compartment features), **Table S9** (chemical-class enrichment), **Table S10** (temporal persistence), and **Table S11**/main-text **Fig. 6D** (features increased in two or more compartments, and their pathway enrichment). No supplementary figure; all figures here are online-only | [§10 Cross-compartment analyses in maternal and infant blood](#cross-compartment-analyses-in-maternal-and-infant-blood) |

Each "Section *N* of the online resource" in the manuscript corresponds to §*N* here.

## Methods recap

We conducted a harmonized cross-trial analysis of individual participant human-milk data from 640 mother–infant dyads drawn from three randomized controlled trials in one low-income and two lower-middle-income countries: MISAME-III (Burkina Faso; prenatal and/or postnatal balanced energy–protein supplementation), Mumta-LW (Pakistan; postnatal BEP and BEP+infant azithromycin), and ELICIT (Tanzania; maternal nicotinamide during lactation with independent infant antimicrobial arms). Each trial contributed its own intent-to-treat population, restricted to dyads with at least one HM measurement contributing to IMiC.

Primary outcomes are macronutrients and micronutrients; secondary are HMOs and bioactive proteins; tertiary are targeted metabolites (Biocrates MxP Quant 500); exploratory are untargeted metabolomics, untargeted proteomics, and microbiome composition. All components were expressed in native assay units, then standardized to study × visit z-scores before TMLE, so effect estimates on the z-score scale are mean differences in SD units. FDR correction was applied within each outcome group × study × visit. Complete methods are in the manuscript's Materials and methods.

The enrichment and cross-compartment sections use two different mass tolerances, which are not interchangeable. Pathway and compound annotation in §9 (Mummichog) uses a 10 ppm database tolerance against the `human_mfn` library, whereas cross-platform feature matching in §10 uses a ±25 ppm tolerance between two measured masses (`ppm = |m₁ − m₂| / m × 10⁶ ≤ 25`).


## How to read interactive figures

In the web version, forest and volcano plots are interactive `plotly` widgets. Hovering over a point shows the component name, study, visit, ATE, 95% CI, and FDR-adjusted *Q*-value, and the mode bar at the top right of each plot zooms, pans, lasso-selects, or exports a PNG. Each plot is paired with a searchable table of the same data, with column filters and sortable columns.

Plots drawn for this online resource follow the conventions of the printed figures, and panels are titled by study and visit, as in "MISAME-III (1-2 mo.)":

- **Forest plots** (as in Fig. 2 and Figs. S2–S3) mark each estimate as not significant (gray open circle), nominally significant at *P* < 0.05 (blue open circle), or significant after FDR correction at *Q* < 0.05 (orange filled circle). Where randomized arms are shown separately, each contrast has its own color and CI line type, and the symbol alone gives the significance (open circle, open triangle, filled circle).
- **Volcano plots** (as in Figs. 3A and 5A) draw estimates that are not FDR-significant in black (squares if *P* < 0.05, circles otherwise) and give FDR-significant estimates the color and symbol of their component category. The gray line marks *P* = 0.05 and the green line the FDR threshold of each panel.
- **Studies** are colored and shaped as in the pathway figures: ELICIT blue circles, MISAME-III orange triangles, and Mumta-LW reddish-purple squares.

Every color code has a second cue (symbol shape, fill, or line type), so the plots can be read with a color-vision deficiency or in grayscale.

If a data file that a section depends on is missing, the section prints a notice naming the file in place of its output. Rerunning the analysis code (see *Reproducing the online resource* in the repository README) regenerates these files.

## Machine-readable data files {#data-files}

Four supplementary tables (S1, S3, S6, and S7) were too large to print in full, so the supplementary materials give a short caption for each and point here for the complete data; Table S1 comes in wide and long formats. A sixth file holds the native-unit intervention effects for every targeted metabolite. Each is a plain CSV that can be downloaded directly:

```{r data-files-list, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
## Publish data-files/ alongside the book so the files the supplementary
## materials name are downloadable from the rendered site. The published copy
## is rebuilt on every render, so a file removed from data-files/ also leaves
## docs/; docs/appendix-data/ is the folder's name before the 2026-09 rename.
src <- here::here("data-files")
dst <- here::here("docs", "data-files")
unlink(here::here("docs", "appendix-data"), recursive = TRUE)
if (dir.exists(src)) {
  unlink(dst, recursive = TRUE)
  dir.create(dst, recursive = TRUE, showWarnings = FALSE)
  file.copy(list.files(src, full.names = TRUE), dst, overwrite = TRUE)
}
## file name | section label it is rendered in | description
files <- list(
  c("table_s1_primary_secondary_native_units_wide.csv", "native-units-table",
    "Table S1: native-unit concentrations and intervention effects for all primary and secondary outcomes, one row per trial × visit × outcome, with the per-arm statistics side by side and the ATE, 95% CI, *P*, and *Q*"),
  c("table_s1_primary_secondary_native_units.csv", "native-units-table",
    "Table S1 in long format: per-arm sample sizes, adjusted means, and raw distributional summaries, one row per trial × visit × outcome × arm (per-arm statistics only)"),
  c("targeted_metabolites_native_units.csv", "native-units-tertiary",
    "native-unit intervention effects for all 581 targeted metabolites, including every triglyceride species, one row per trial × visit × metabolite"),
  c("table_s3_tertiary_msea_full.csv", "path-tertiary-msea",
    "Table S3: tertiary targeted-metabolome MSEA, every pathway × study × visit × contrast × direction row"),
  c("table_s6_mummichog_full.csv", "path-mummichog",
    "Table S6: milk untargeted-metabolome Mummichog enrichment, pathways at raw *P* < 0.05"),
  c("table_s7_proteomics_go_full.csv", "path-go",
    "Table S7: milk-proteome GO Biological Process enrichment, terms at nominal *P* < 0.05 and Storey *q* < 0.20")
)
for (f in files) {
  p <- file.path(src, f[1])
  n <- if (file.exists(p)) max(0, length(readLines(p, warn = FALSE)) - 1L) else NA_integer_
  cat(sprintf("- [`%s`](data-files/%s): %s (%s; see §\\@ref(%s))\n", f[1], f[1], f[3],
              if (is.na(n)) "*not present in this build*" else paste(format(n, big.mark = ","), "rows"),
              f[2]))
}
```

Each file is also shown as a searchable table in the section listed beside it.

## How to cite this online resource {#how-to-cite}

Please cite the article together with this online resource:

> Dailey-Chwalibóg T, Mertens A, *et al.* (2026). Nutritional interventions' impacts on human milk: three trials in low-resource settings. International Milk Composition (IMiC) Consortium.
>
> Mertens A (2026). IMiC intervention effects on human milk: online resource. https://amertens.github.io/IMiC-intervention-effects-supplement/. Archived on Zenodo with the analysis code, DOI: 10.5281/zenodo.22104233 (https://doi.org/10.5281/zenodo.22104233).

- Online resource: <https://amertens.github.io/IMiC-intervention-effects-supplement/>
- DOI (Zenodo archive of the analysis code and this online resource): 10.5281/zenodo.22104233, <https://doi.org/10.5281/zenodo.22104233>
- Source for this online resource: <https://github.com/amertens/IMiC-intervention-effects-supplement>
- Analysis code (TMLE pipelines, SuperLearner, enrichment workflows): <https://github.com/amertens/IMiC-intervention-effects-public>
- Registration: IMiC, ClinicalTrials.gov NCT05119166; constituent trials NCT03533712 (MISAME-III), NCT03564652 (Mumta-LW), and NCT03268902 (ELICIT)

```{r write-bib, include=FALSE}
knitr::write_bib(c(.packages(), "bookdown","knitr","rmarkdown","tidyverse","here",
                   "DT","plotly","table1"),
                 "packages.bib")
```
