---
title: "IMiC intervention effects on human milk: online resource"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
description: |
  Online resource for "Nutritional interventions' impacts on human milk: three trials in low-resource settings" (Dailey-Chwalibóg, Mertens et al., 2026).
link-citations: yes
github-repo: amertens/IMiC-intervention-effects-supplement
url: https://amertens.github.io/IMiC-intervention-effects-supplement/
---

# Overview {-}

This online resource accompanies Dailey-Chwalibóg, Mertens *et al.* (2026), "Nutritional interventions' impacts on human milk: three trials in low-resource settings", from the International Milk Composition (IMiC) Consortium. It holds the full results tables, interactive views of the estimates, and analyses that are not in the article.

| Section | Contents |
|---|---|
| [1](#baseline-characteristics-by-study-and-intervention-arm) | Baseline characteristics by trial and arm |
| [2](#outcomes-and-human-milk-component-distributions) | All milk components analyzed; their distributions by trial, visit, and arm |
| [3](#subgroup-analyses-stratified-by-maternal-bmi) | Effects by maternal BMI |
| [4](#exploratory-high-dimensional-outcomes-microbiome-untargeted-proteomics-untargeted-metabolomics) | Effects on the microbiome, untargeted proteome, and untargeted metabolome |
| [5](#adjusted-intervention-effects-on-individual-milk-components) | Component metadata; effects on each targeted milk component, with an interactive explorer |
| [6](#effects-on-human-milk-nutrient-deficiency) | Effects on nutrient deficiency |
| [7](#trajectory-analyses-of-human-milk-component-change-across-lactation) | Effects on change between visits |
| [8](#sensitivity-analyses-and-additional-outputs) | Unadjusted and native-unit effects (Table S1), treatment-arm prediction, principal-component scores, infant growth |
| [9](#reproducible-pathway-and-enrichment-analyses) | Pathway and enrichment analyses (Tables S2–S7), with an interactive explorer |
| [10](#cross-compartment-analyses-in-maternal-and-infant-blood) | Maternal and infant blood in MISAME-III (Tables S8–S11) |

## Estimates and plots {-}

Effects are average treatment effects (ATEs) estimated by targeted maximum likelihood estimation (TMLE) with adjustment for baseline covariates, in standard deviations of each component's study- and visit-specific z-score unless a table gives native units. *Q* is the Benjamini–Hochberg FDR-adjusted *P* value, computed within outcome group, study, and visit.

In volcano plots, points with *Q* ≥ 0.05 are black (squares if *P* < 0.05), and points with *Q* < 0.05 take the color and symbol of their component category; the gray line marks *P* = 0.05 and the green line the FDR threshold of each panel. In forest plots of randomized arms, each arm has its own color and the symbol marks significance: open circle *P* ≥ 0.05, open triangle *P* < 0.05, filled circle *Q* < 0.05. Hover over a point for its values. Tables can be searched, sorted, and filtered by column.

## Data files {#data-files .unnumbered}

The complete versions of the supplementary tables printed in part, as CSV files:

```{r data-files-list, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
## Publish data-files/ alongside the book so the files the supplementary
## materials name are downloadable from the rendered site. The published copy
## is rebuilt on every render, so a file removed from data-files/ also leaves
## docs/; docs/appendix-data/ is the folder's name before the 2026-09 rename.
src <- here::here("data-files")
dst <- here::here("docs", "data-files")
unlink(here::here("docs", "appendix-data"), recursive = TRUE)
if (dir.exists(src)) {
  unlink(dst, recursive = TRUE)
  dir.create(dst, recursive = TRUE, showWarnings = FALSE)
  invisible(file.copy(list.files(src, full.names = TRUE), dst, overwrite = TRUE))
}
## file name | section label it is rendered in | description
files <- list(
  c("table_s1_primary_secondary_native_units.csv", "native-units-table",
    "Table S1, one row per trial, visit, and outcome: per-arm sample size and raw distribution side by side, with the native-unit ATE, 95% CI, *P*, and *Q*"),
  c("targeted_metabolites_native_units.csv", "native-units-tertiary",
    "native-unit arm means and effects for the 581 targeted metabolites, one row per trial, visit, and metabolite"),
  c("table_s3_tertiary_msea_full.csv", "path-tertiary-msea",
    "Table S3, targeted-metabolome MSEA, all rows"),
  c("table_s6_mummichog_full.csv", "path-mummichog",
    "Table S6, untargeted-metabolome Mummichog enrichment, pathways with *P* < 0.05"),
  c("table_s7_proteomics_go_full.csv", "path-go",
    "Table S7, proteome GO Biological Process enrichment, terms with *P* < 0.05 and Storey *q* < 0.20")
)
for (f in files) {
  p <- file.path(src, f[1])
  n <- if (file.exists(p)) max(0, length(readLines(p, warn = FALSE)) - 1L) else NA_integer_
  cat(sprintf("- [`%s`](data-files/%s): %s (%s; Section \\@ref(%s))\n", f[1], f[1], f[3],
              if (is.na(n)) "*not present in this build*" else paste(format(n, big.mark = ","), "rows"),
              f[2]))
}
```

## How to cite {#how-to-cite .unnumbered}

Cite the article, and cite this online resource by the DOI of its archive, [10.5281/zenodo.22104233](https://doi.org/10.5281/zenodo.22104233), which also holds the analysis code.

- Source for this site: <https://github.com/amertens/IMiC-intervention-effects-supplement>
- Analysis code: <https://github.com/amertens/IMiC-intervention-effects-public>
- Trial registrations (ClinicalTrials.gov): IMiC NCT05119166; MISAME-III NCT03533712; Mumta-LW NCT03564652; ELICIT NCT03268902

```{r write-bib, include=FALSE}
knitr::write_bib(c(.packages(), "bookdown","knitr","rmarkdown","tidyverse","here",
                   "DT","plotly","table1"),
                 "packages.bib")
```
